### 2015-02-17 ProjectProposal/ProjectProposal.tex
Adding one page project proposal

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}
\noindent Kelsey Maass \\
Brisa Davis \\
AMATH 574\\
\vspace{0.2in}

We are planning on studying different second order traffic flow models. Currently, we plan to study a second order models that failed to accurately capture the behavior seen in reality. We will examine the Payne and Whitham (PW) model, and explore the logical flaws in the arguments that have been used to derive this higher order model \cite{Daganzo1995}. With this study we will illustrate the importance of validation in modeling, and highlight the fact that simply having a higher order code does not equate to more accurately modeling a given physical system.

We also examine the more accurate Aw Rascle model presented in \cite{AwRascle2000}, which uses a convective derivative to resolve inconsistencies present in older models. The main critique this newer model has of the PW model is that the PW model uses fluid flow behavior to generate their model, while car behavior is in reality not accurately captured by fluid flow equations. As an example, fluid particles receive and respond to stimuli from both the front and the back whereas a car mostly responds to frontal stimuli.

We will compare both models to show that the PW model gives nonphysical results while Aw Rascle model more accurately reflects reality.

\bibliography{ProjectProposal}{}
\bibliographystyle{plain}
\end{document}

### 2015-03-03 FirstDraftMaterials/TrafficModels.tex
Added folders Papers and FirstDraftMaterials

\documentclass{article}

\usepackage{amsmath}
\usepackage{mathtools}
\setlength\parindent{0pt}

\begin{document}

\section{Traffic Flow}

To model cars moving on a one-way highway without exits, entrances, or passing, we can use the fact that the quantity (or density) of cars must be conserved:

\[ \frac{d}{dt} \int_{x_1}^{x_2} \rho(x,t) dt = -\Bigg[ f\Big(\rho(x,t)\Big) \Bigg]_{x_1}^{x_2} \] \\

In differential form, this becomes:

\[ \rho_t + f \Big( \rho(x,t) \Big)_x = 0 \] \\

In light traffic conditions velocity might only depend upon $x$.  Therefore the flux function becomes $f(\rho(x,t)) = \rho(x,t) u(x)$:

\[ \rho_t + (\rho u)_x = 0 \] \\

In more congested traffic, velocity may depend more upon density than $x$, so that $u = U(\rho)$.  The flux function then becomes $f(\rho(x,t)) = \rho(x,t) U(\rho(x,t))$:

\[ \rho_t + \Big( \rho U(\rho) \Big)_x = 0 \] \\

This is the LWR model, proposed by Lighthill and Whitham (1955) and Richards (1956).  Here $U(\rho)$ is the preferred velocity, a given non increasing function of $\rho$, nonnegative for $\rho$ between 0 and some positive maximum density $\rho_m$. (WOULD IT MAKES SENSE FOR IT TO BE 1? CHECK THE TEXTBOOK...). \\

This model behaves well macroscopically, but does have some flaws.  For example, it does not describe what happens to cars as they travel through shocks.  Realistically, we know that cars cannot change their velocities instantaneously, so one way to try to improve the model would be to eliminate the presence of shocks.  This can be done by adding adding a diffusion term:

\[ \rho_t + \Big( \rho U(\rho) \Big)_x = \mu \rho_{xx} \] \\

where $\mu$ is a positive constant.  This term is supposed to account for the driver's awareness of the road ahead.  (Though as Daganzo points out, it also implies that drivers are aware of conditions behind, which is a problem).  (REALISTICALLY DRIVERS ARE AWARE OF WHAT IS GOING ON BEHIND THEM, THOUGH IT MAY NOT HAVE AS STRONG OF AN AFFECT ON HOW THEY DRIVE).  This added term causes problems as it predicts negative velocities in some instances, which is obviously wrong.

\section{Compressible Gas Dynamics}

Another attempt to improve the LWR model is to try to capture more of they dynamics present by adding another equation.  Treating traffic as a compressible gas, we can use equations for the conservation of mass and conservation of momentum used to model gas dynamics.  For conservation of mass, we can use one of the equation we derived above, where the flux function is $f = \rho u$:

\[ \rho_t + (\rho u)_x = 0 \] \\

For conservation of momentum, the flux function includes the contribution of the flux of momentum $\rho u^2$ and a pressure term $p$, where pressure is a function of density.  (DESCRIBE HOW THIS FUNCTION BEHAVES).  The integral form of this conservation law is:

\[ \frac{d}{dt} \int_{x_1}^{x_2} \rho(x,t) u(x,t) dx = -\Big[ \rho u^2 + p \Big]_{x_1}^{x_2} \] \\

In differential form this equation becomes:

\[ (\rho u)_t + (\rho u^2 + p)_x = 0 \] \\

Combining the differential forms of these two equations, we can rewrite the second equation in terms of velocity instead of momentum:

\begin{align*}
0 &= (\rho u)_t + (\rho u^2 + p)_x \\[1ex]
&= \rho_t u + \rho u_t + \rho_x u^2 + 2 \rho u u_x + p_x \\[1ex]
&= \left( -\rho_x u - \rho u_x \right) u + \rho u_t + \rho_x u^2 + 2 \rho u u_x + p_x \\[1ex]
&= \rho u_t + \rho u u_x + p_x \\[1ex]
&= u_t + u u_x + \frac{p_x}{\rho} 
\end{align*}

\section{PW Model}

By modifying these equations for conservation of mass and momentum in gas dynamics, both Payne (1971) and Whitham (1974) attempted to improve upon the LWR model with the 2-equation model:

\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho u)_x = 0 \\[2ex] & u_t + u u_x + \dfrac{p'(\rho)}{\rho} \rho_x = \dfrac{U(\rho) - u}{
\tau} + \mu u_{xx} \end{matrix*} \right. \] \\

Here we have the two conservation laws with two additional terms in the second equation.  The first represents a relaxation towards a preferred velocity (EQUILIBRIUM VELOCITY) $U(\rho)$, where $\tau$ is (I DON'T REMEMBER, THE REACTION TIME?) and the second is a diffusion term, as discussed previously (though here it is applied to velocity instead of density).  (ONE OF THE PAPERS I READ MENTIONED THAT IN THE LIMIT THAT TAU GOES TO 0, THESE EQUATIONS REDUCE TO THE VISCUOUS LWR MODEL MENTIONED ABOVE.) \\

At first look, this seems like a reasonable model, since macroscopically cars can behave like a compressible gas.  However, there are some problematic assumptions made by the model which can produce unrealistic results, such as negative velocities.  WHAT ARE ALL OF THE PROBLEMS?  (CARS AND GAS MOLECULES ARE FUNDAMENTALLY DIFFERENT, CARS DON'T GO BACKWARDS ON THE FREEWAY, THEY ARE MOSTLY INFLUENCED BY WHAT IS IN FRONT OF THEM AND NOT BEHIND, CARS (DRIVERS) CAN HAVE DIFFERENT PERSONALITIES, PLATOONS...)

\[ \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_t + \begin{pmatrix} u & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & u \end{pmatrix} \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] \dfrac{U(\rho) - u}{\tau} + \mu u_{xx} \end{pmatrix} \] \\

\[ \lambda_1 = u - c,  \, r_1 = \begin{pmatrix} \rho \\ -c \end{pmatrix} \hspace{1em} \lambda_2 = u + c, \, r_2 = \begin{pmatrix} \rho \\ c \end{pmatrix} \] \\

where $c = \sqrt{p'(\rho)}$, the traffic sound speed.  We see from the second eigenvalue that some information is moving faster than the velocity of the cars, which implies that drivers are influenced by what is going on behind them.

\section{Improved Model}

As A. Aw and M. Rascle point out in their paper, the anticipation factor (TERM WITH PRESSURE), involves the derivative of pressure with respect to $x$.  However, we should instead be concerned with the perspective of the driver, which is relative to a moving timeframe.  (THOUGHT EXPERIMENT IN PAPER).  In this case, we should really be using a convective derivative $\partial_t + u \partial_x$ of the pressure.  In this case, we can consider the model:

\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho u)_x = 0 \\[1ex] (u + p)_t + u (u + p)_x = 0 \end{matrix*} \right. \] \\

SHOULD I ALWAYS WRITE P AS A FUNCTION OF RHO, OR IS IT OKAY TO LEAVE IT OUT??? ALSO NEED TO MENTION THAT WE ARE ASSUMING THESE FUNCTIONS ARE SMOOTH/DIFFERENTIABLE

This can also be rewritten as a system of two equations, one for density and the other for velocity:

\begin{align*}
0 &= (u + p)_t + v (v + p)_x \\[1ex]
&= u_t + p'(\rho) \rho_t + u u_x + u p'(\rho) \rho_x \\[1ex]
&= u_t + p'(\rho) (-\rho_x u - \rho u_x) + u u_x + u p'(\rho) \rho_x \\[1ex]
&= u_t + (u - p'(\rho) \rho) u_x 
\end{align*}

Second-order model:

\[ \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_t + \begin{pmatrix} u & \rho \\[1ex] 0 & u - \rho p'(\rho) \end{pmatrix} \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_x = 0 \] \\

This model conserves density and "momentum": $y = \rho u + \rho p(\rho)$ (WE CAN REWRITE THE EQUATIONS IN CONSERVATION FORM...) In this case, $y$ doesn't seem to have an important physical meaning... \\

Where the PW model failed, this model performs well, though it still doesn't perfectly capture what happens in light traffic conditions.  We can show with the eigenvalues that no information travels faster than the speed of the cars.  We can demonstrate through an example where the PW predicts negative velocities that this model still produces reasonable results.  \\

One moral of the story: Adding complexity does not guarantee improved accuracy.  The assumptions you make when you design your model must agree with the dynamics you are trying to capture.  In this case, assumptions that can be made about gas particles cannot be applied to traffic conditions.  This is one reason why the PW model fails to improve the LWR model.  ONE PAPER HAD CONDITIONS FOR WHAT CONSTITUTES A GENUINE IMPROVEMENT.  However, the authors of the second paper laid out what conditions must be satisfied for modeling traffic, and they tested their model against these conditions to show that their assumptions were indeed valid.

\end{document}

### 2015-03-03 FirstDraftMaterials/Outline.tex
adding outline

--- 2015-03-03 FirstDraftMaterials/TrafficModels.tex
@@
 
-\section{Traffic Flow}
-
-To model cars moving on a one-way highway without exits, entrances, or passing, we can use the fact that the quantity (or density) of cars must be conserved:
-
-\[ \frac{d}{dt} \int_{x_1}^{x_2} \rho(x,t) dt = -\Bigg[ f\Big(\rho(x,t)\Big) \Bigg]_{x_1}^{x_2} \] \\
-
-In differential form, this becomes:
-
-\[ \rho_t + f \Big( \rho(x,t) \Big)_x = 0 \] \\
-
-In light traffic conditions velocity might only depend upon $x$.  Therefore the flux function becomes $f(\rho(x,t)) = \rho(x,t) u(x)$:
-
-\[ \rho_t + (\rho u)_x = 0 \] \\
-
-In more congested traffic, velocity may depend more upon density than $x$, so that $u = U(\rho)$.  The flux function then becomes $f(\rho(x,t)) = \rho(x,t) U(\rho(x,t))$:
-
-\[ \rho_t + \Big( \rho U(\rho) \Big)_x = 0 \] \\
-
-This is the LWR model, proposed by Lighthill and Whitham (1955) and Richards (1956).
-Here $U(\rho)$ is the preferred velocity, a given non increasing function of $\rho$, nonnegative for $\rho$ between 0 and some positive maximum density $\rho_m$.
-(WOULD IT MAKES SENSE FOR IT TO BE 1?
-CHECK THE TEXTBOOK...).
-\\
-
-This model behaves well macroscopically, but does have some flaws.
-For example, it does not describe what happens to cars as they travel through shocks.
-Realistically, we know that cars cannot change their velocities instantaneously, so one way to try to improve the model would be to eliminate the presence of shocks.
-This can be done by adding adding a diffusion term:
-
-\[ \rho_t + \Big( \rho U(\rho) \Big)_x = \mu \rho_{xx} \] \\
-
-where $\mu$ is a positive constant.
-This term is supposed to account for the driver's awareness of the road ahead.
-(Though as Daganzo points out, it also implies that drivers are aware of conditions behind, which is a problem).
-(REALISTICALLY DRIVERS ARE AWARE OF WHAT IS GOING ON BEHIND THEM, THOUGH IT MAY NOT HAVE AS STRONG OF AN AFFECT ON HOW THEY DRIVE).
-This added term causes problems as it predicts negative velocities in some instances, which is obviously wrong.
-
-\section{Compressible Gas Dynamics}
-
-Another attempt to improve the LWR model is to try to capture more of they dynamics present by adding another equation.
-Treating traffic as a compressible gas, we can use equations for the conservation of mass and conservation of momentum used to model gas dynamics.
-For conservation of mass, we can use one of the equation we derived above, where the flux function is $f = \rho u$:
-
-\[ \rho_t + (\rho u)_x = 0 \] \\
-
-For conservation of momentum, the flux function includes the contribution of the flux of momentum $\rho u^2$ and a pressure term $p$, where pressure is a function of density.
-(DESCRIBE HOW THIS FUNCTION BEHAVES).
-The integral form of this conservation law is:
-
-\[ \frac{d}{dt} \int_{x_1}^{x_2} \rho(x,t) u(x,t) dx = -\Big[ \rho u^2 + p \Big]_{x_1}^{x_2} \] \\
-
-In differential form this equation becomes:
-
-\[ (\rho u)_t + (\rho u^2 + p)_x = 0 \] \\
-
-Combining the differential forms of these two equations, we can rewrite the second equation in terms of velocity instead of momentum:
-
-\begin{align*}
-0 &= (\rho u)_t + (\rho u^2 + p)_x \\[1ex]
-&= \rho_t u + \rho u_t + \rho_x u^2 + 2 \rho u u_x + p_x \\[1ex]
-&= \left( -\rho_x u - \rho u_x \right) u + \rho u_t + \rho_x u^2 + 2 \rho u u_x + p_x \\[1ex]
-&= \rho u_t + \rho u u_x + p_x \\[1ex]
-&= u_t + u u_x + \frac{p_x}{\rho} 
-\end{align*}
-
+\section{Introduction}
+Kelsey
+\section{LWR Model}
+Kelsey
 \section{PW Model}
 
-By modifying these equations for conservation of mass and momentum in gas dynamics, both Payne (1971) and Whitham (1974) attempted to improve upon the LWR model with the 2-equation model:
+\subsection{Theory}
+Kelsey
+\subsection{Hugoniot Loci and Integral Curves}
+Brisa
+\section{AR Model}
 
-\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho u)_x = 0 \\[2ex] & u_t + u u_x + \dfrac{p'(\rho)}{\rho} \rho_x = \dfrac{U(\rho) - u}{
-\tau} + \mu u_{xx} \end{matrix*} \right. \] \\
+\subsection{Theory}
+Kelsey
+\subsection{Hugoniot Loci and Integral Curves}
+Brisa
 
-Here we have the two conservation laws with two additional terms in the second equation.
-The first represents a relaxation towards a preferred velocity (EQUILIBRIUM VELOCITY) $U(\rho)$, where $\tau$ is (I DON'T REMEMBER, THE REACTION TIME?) and the second is a diffusion term, as discussed previously (though here it is applied to velocity instead of density).
-(ONE OF THE PAPERS I READ MENTIONED THAT IN THE LIMIT THAT TAU GOES TO 0, THESE EQUATIONS REDUCE TO THE VISCUOUS LWR MODEL MENTIONED ABOVE.) \\
-
-At first look, this seems like a reasonable model, since macroscopically cars can behave like a compressible gas.
-However, there are some problematic assumptions made by the model which can produce unrealistic results, such as negative velocities.
-WHAT ARE ALL OF THE PROBLEMS?
-(CARS AND GAS MOLECULES ARE FUNDAMENTALLY DIFFERENT, CARS DON'T GO BACKWARDS ON THE FREEWAY, THEY ARE MOSTLY INFLUENCED BY WHAT IS IN FRONT OF THEM AND NOT BEHIND, CARS (DRIVERS) CAN HAVE DIFFERENT PERSONALITIES, PLATOONS...)
-
-\[ \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_t + \begin{pmatrix} u & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & u \end{pmatrix} \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] \dfrac{U(\rho) - u}{\tau} + \mu u_{xx} \end{pmatrix} \] \\
-
-\[ \lambda_1 = u - c,  \, r_1 = \begin{pmatrix} \rho \\ -c \end{pmatrix} \hspace{1em} \lambda_2 = u + c, \, r_2 = \begin{pmatrix} \rho \\ c \end{pmatrix} \] \\
-
-where $c = \sqrt{p'(\rho)}$, the traffic sound speed.
-We see from the second eigenvalue that some information is moving faster than the velocity of the cars, which implies that drivers are influenced by what is going on behind them.
-
-\section{Improved Model}
-
-As A.
-Aw and M.
-Rascle point out in their paper, the anticipation factor (TERM WITH PRESSURE), involves the derivative of pressure with respect to $x$.
-However, we should instead be concerned with the perspective of the driver, which is relative to a moving timeframe.
-(THOUGHT EXPERIMENT IN PAPER).
-In this case, we should really be using a convective derivative $\partial_t + u \partial_x$ of the pressure.
-In this case, we can consider the model:
-
-\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho u)_x = 0 \\[1ex] (u + p)_t + u (u + p)_x = 0 \end{matrix*} \right. \] \\
-
-SHOULD I ALWAYS WRITE P AS A FUNCTION OF RHO, OR IS IT OKAY TO LEAVE IT OUT??? ALSO NEED TO MENTION THAT WE ARE ASSUMING THESE FUNCTIONS ARE SMOOTH/DIFFERENTIABLE
-
-This can also be rewritten as a system of two equations, one for density and the other for velocity:
-
-\begin{align*}
-0 &= (u + p)_t + v (v + p)_x \\[1ex]
-&= u_t + p'(\rho) \rho_t + u u_x + u p'(\rho) \rho_x \\[1ex]
-&= u_t + p'(\rho) (-\rho_x u - \rho u_x) + u u_x + u p'(\rho) \rho_x \\[1ex]
-&= u_t + (u - p'(\rho) \rho) u_x 
-\end{align*}
-
-Second-order model:
-
-\[ \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_t + \begin{pmatrix} u & \rho \\[1ex] 0 & u - \rho p'(\rho) \end{pmatrix} \begin{pmatrix} \rho \\[1ex] u \end{pmatrix}_x = 0 \] \\
-
-This model conserves density and "momentum": $y = \rho u + \rho p(\rho)$ (WE CAN REWRITE THE EQUATIONS IN CONSERVATION FORM...) In this case, $y$ doesn't seem to have an important physical meaning...
-\\
-
-Where the PW model failed, this model performs well, though it still doesn't perfectly capture what happens in light traffic conditions.
-We can show with the eigenvalues that no information travels faster than the speed of the cars.
-We can demonstrate through an example where the PW predicts negative velocities that this model still produces reasonable results.
-\\
-
-One moral of the story: Adding complexity does not guarantee improved accuracy.
-The assumptions you make when you design your model must agree with the dynamics you are trying to capture.
-In this case, assumptions that can be made about gas particles cannot be applied to traffic conditions.
-This is one reason why the PW model fails to improve the LWR model.
-ONE PAPER HAD CONDITIONS FOR WHAT CONSTITUTES A GENUINE IMPROVEMENT.
-However, the authors of the second paper laid out what conditions must be satisfied for modeling traffic, and they tested their model against these conditions to show that their assumptions were indeed valid.
+\section{Examples}
+Brisa
 

### 2015-03-04 MatlabCode/variableSetup.m
Adding Matlab code and generated images for AW model

@@ -1,0 +1,12 @@
+function y = variableSetup
+
+global v_star rho_star p pprime rho
+
+% Declaring various starting values
+v_star = [0.2 .5 .8];
+rho_star = [0.2 .06 .1];
+gamma = 0.2;
+
+p = @(x) x^gamma;
+pprime = @(x) gamma*x^(gamma-1);
+rho = 0:0.001:5;

### 2015-03-04 FirstDraftMaterials/Draft.tex
Adding draft document
Adding initial filling out of HL and IC for AW model, and bibliography.

--- 2015-03-03 FirstDraftMaterials/Outline.tex
@@
 \usepackage{mathtools}
+\usepackage{subfigure}
 \setlength\parindent{0pt}
+
+\setlength{\textwidth}{6.2in}
+\setlength{\oddsidemargin}{0.0in}
+\setlength{\evensidemargin}{0in}
+\setlength{\textheight}{8.9in}
+\setlength{\voffset}{-1in}
+\setlength{\headsep}{26pt}
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{5pt}
 
@@
 \subsection{Hugoniot Loci and Integral Curves}
-Brisa
+The AR model starts from the system
+\begin{align}
+&\partial_t\rho + \partial_x(\rho v) = 0, \label{AW:eq1}\\
+&\partial_t \left(v + p(\rho )\right) + v\partial_x \left( v + p(\rho )\right) = 0.
+\end{align}
+
+To find the Hugoniot loci and the integral curves, \cite{AwRascle2000} utilizes two different forms of this system. 
+We will examine each of them in turn. What we will find is that the integral curves and the Hugoniot loci coincide, 
+and furthermore that one of the waves is a contact discontinuity. Let us first consider the Hugoniot loci.
+
+\subsubsection{Hugoniot Loci}
+In order to consider the Hugoniot loci for this system we must first write the system in conversation form. 
+To accomplish this, note that multiplying the first equation by $(v + p(\rho ))$ and the second equation 
+by $\rho$ gives us the two equations 
+\begin{align*}
+&(v + p(\rho ))\partial_t\rho + (v + p(\rho ))\partial_x(\rho v) = 0,  \\
+&\rho\partial_t \left(v + p(\rho )\right) + \rho v\partial_x \left( v + p(\rho )\right) = 0.
+\end{align*}
+Adding these two equations gives
+\begin{align}\label{AW:eq2}
+\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
+\end{align}
+Then from (\ref{AW:eq1}) and (\ref{AW:eq2}) we have the system 
+\begin{align*}
+&\partial_t\rho + \partial_x(\rho v) = 0, \\
+&\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
+\end{align*}
+This can be rewritten as $q_t + f(q)_x = $ by defining
+\begin{align*}
+q = \left[ \begin{matrix}
+\rho \\ y
+\end{matrix}\right], \hspace{0.3in}
+f(q) = \left[ \begin{matrix}
+v\rho \\
+vy
+\end{matrix}\right].
+\end{align*}
+where $y = \rho\left(v + p(\rho )\right)$.
+The Rankine-Hugoniot condition tells us that 
+\begin{align*}
+s(q_*- q) = f(q_*) - f(q).
+\end{align*}
+Therefore, from this condition we get the two equations
+\begin{align}
+s(\rho_* - \rho) = v_*\rho_* - v\rho, \label{AW:eq3}\\
+s(y_* - y) = v_*y_* - vy.\label{AW:eq4}
+\end{align}
+With a little rearranging and by plugging in for $y$, these two equations can be rewritten as
+\begin{align}
+&\rho_*(s - v_*) = \rho (s - v), \label{AW:eq5}\\
+&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).
+\end{align}
+From (\ref{AW:eq5}), we know that there are two cases:
+
+(a) $\rho_*(s - v_*) = \rho (s - v) = 0$. In this case we have that
+\begin{align*}
+s - v_* = s - v 
+\end{align*}
+except if one of the two densities, $\rho_*$, $\rho$ is zero. WHY DOES THIS NOT MATTER? Therefore, we can say that
+\begin{align*}
+v_* = v.
+\end{align*}
+Since the velocity does not change, this is a contact discontinuity. Using the fact that $v = v_*$ in (\ref{AW:eq4}) gives
+\begin{align*}
+y_*(s - v) = y(s - v).
+\end{align*}
+So, either $y = y_*$ or $s = v$. If $y = y_*$ we have a stationary point in the $Y = (\rho, y)$ plane, which does not 
+describe a wave. Therefore, $s = v$. Note that we have no restrictions on $y$ other than its definition. 
+Therefore this loci in the $Y$ plane is given by 
+\begin{align*}
+y = \rho ( v_* + p(\rho )).
+\end{align*}
+(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, solving (\ref{AW:eq3}) for $s$ gives
+\begin{align*}
+s = \frac{\rho_*v_* - \rho v}{\rho_* - \rho}.
+\end{align*}
+
+\begin{figure}[h!]
+ \centering
+ \subfigure[$\lambda_1$-curves in the U plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U_lamb1.jpg}
+   \label{fig:subfig1}
+   }
+ \subfigure[$\lambda_2$-curves in the U plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U_lamb2.jpg}
+   \label{fig:subfig2}
+   }
+ \subfigure[$\lambda_1$-curve and $\lambda_2$-curve in the U plane passing through
+  the point $(\rho_*, v_*)$ denoted by the red point.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U.jpg}
+   \label{fig:subfig3}
+   }
+    \subfigure[$\lambda_1$-curves in the M plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb1.jpg}
+   \label{fig:subfig4}
+   }
+ \subfigure[$\lambda_2$-curves in the M plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb2.jpg}
+   \label{fig:subfig5}
+   }
+ \subfigure[$\lambda_1$-curve and $\lambda_2$-curve in the M plane passing through 
+ the point $(\rho_*,m_*)$ denoted by the red point.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M.jpg}
+   \label{fig:subfig6}
+   }
+       \subfigure[$\lambda_1$-curves in the Y plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y_lamb1.jpg}
+   \label{fig:subfig7}
+   }
+ \subfigure[$\lambda_2$-curves in the Y plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y_lamb2.jpg}
+   \label{fig:subfig8}
+   }
+ \subfigure[$\lambda_1$-curve and $\lambda_2$-curve in the Y plane passing through 
+ the point $(\rho_*, y_*)$ denoted by the red point.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y.jpg}
+   \label{fig:subfig9}
+   }
+ \label{fig:11_8c}
+ \caption[Optional caption for list of figures]
+ {Figures of the Hugonio loci and integral curves for $\lambda_1$ and $\lambda_2$ 
+ in each of the different planes. Since the loci and integral curves coincide, they are both
+  simply called the ``curves'' in the captions of the above figures. Figures (c), (f), (i) show
+   both curves for a given point in each plane.}
+\end{figure}
 
@@
 
+\bibliography{Draft}{}
+\bibliographystyle{plain}
 \end{document}

### 2015-03-04 FirstDraftMaterials/Outline.tex
Added Kelsey's sections in outline

@@
 \usepackage{mathtools}
+\usepackage{enumerate}
+\usepackage[margin=1in]{geometry}
 \setlength\parindent{0pt}
@@
 \section{Introduction}
-Kelsey
+When a model is called ``higher order", we usually associate it with being more accurate than a first order model.
+This is not always the case.
+Adding complexity to a model does not necessarily improve it, as is demonstrated in the class of higher order traffic models developed with the intention of improving the first order traffic model proposed by Lighthill and Whitham (1955) and Richards (1956).
+As Daganzo observes in his paper, \\
+
+``A new theory (e.g.
+relativity) is deemed successful if it explains previously unexplained phenomena, plus everything else that was correctly explained with the theory it intends to replace (e.g.
+Newtonian physics).
+In the case of traffic flow, the new theories (high-order fluid models) fail to explain the behavior of traffic at the end of a queue (unlike the simpler, older LWR model; which explains it perfectly)." \\
+
+In this paper we examine the second order traffic model developed by Payne (1971) and Whitham (1974) which was meant to improve how the LWR model explains traffic behavior near shocks.
+Their model takes its form from equations used to describe compressible fluid dynamics.
+Unfortunately, cars and fluid particles have some fundamental differences which make the model actually perform worse than the LWR model in certain cases.
+For example, by smoothing out discontinuities, the PW model predicts negative velocities, which is clearly a problem.
+\\
+
+In addition to considering the PW model, we also take a look a more successful model proposed by Aw and Rascle (2000).
+These authors challenge some of the assumptions of the PW model and suggest the use of a convective derivative to fix the problem of negative velocities.
+In order to verify that their model is valid, they test their new model against criteria they claim any reasonable model will satisfy.
+
+% LWR Model
 \section{LWR Model}
-Kelsey
+We can start by considering a first order (one equation) model of traffic flow on a one-way highway without any exits, entrances, or passing.
+In this case, we can expect that the number of cars will be conserved, so we can use the equation for conservation of mass (with density $\rho$ and velocity $v$):
+
+\begin{enumerate}[(a)]
+\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho \, dt = -\Big[ f( \rho ) \Big]_{x_1}^{x_2}$
+\item Differential form of equation: $\rho_t + f( \rho )_x = 0$
+\begin{enumerate}[i]
+\item Light traffic: $\rho_t + (\rho v)_x = 0$
+\item Congested traffic: $\rho_t + \Big( \rho V(\rho) \Big)_x = 0$ (LWR model)
+\end{enumerate}
+\end{enumerate}
+
+The LWR behaves well macroscopically, but it does have some flaws.
+For example, it assumes cars change their velocity instantaneously as they travel through shocks.
+One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big( \rho V(\rho) \Big)_x = \mu \rho_{xx}$.
+This term is meant to account for the drivers awareness of the road ahead.
+However, this model can be shown to predict negative velocities in some instances, which is obviously wrong.
+
+% PW Model
 \section{PW Model}
+\subsection{Theory}
+Another approach to improving the LWR model would be to use a second order (two equation) model.
+Since cars can be thought of as a compressible gas, we can use equations for conservation of mass (above) and conservation of momentum, $\rho v$, used to model compressible gas flow:
 
-\subsection{Theory}
-Kelsey
+\begin{enumerate}[(a)]
+\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho v \, dt = -\Big[ f( \rho v ) \Big]_{x_1}^{x_2}$
+\item Differential form of equation: $(\rho v)_t + f( \rho v )_x = (\rho v)_t + \Big( \rho v^2 + \rho \Big)_x = 0$
+\end{enumerate}
+
+We can rewrite the second equation in terms of velocity instead of momentum: $v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = 0$. \\
+
+The PW model is a modification of these equations for conservation of mass and conservation of momentum in gas dynamics, with added relaxation and diffusion terms:
+
+\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho v)_x = 0 \\[1ex] & v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = \dfrac{V(\rho) - v}{\tau} + \mu v_{xx} \end{matrix*} \right. \]
+
+At first glance, this seems like a reasonable model.
+However, there a couple of fundamental differences between gas particles and cars that could be problematic.
+For example, as pointed out by Daganzo, ``a fluid particle responds to stimuli from the front and from behind, but a car is an anisotropic particle that mostly responds to frontal stimuli."  If we consider the eigenvalues of the Jacobian matrix of the linearized system, we see that one of the eigenvalues, $\lambda_2 = v + \sqrt{p'(\rho)}$, represents information traveling faster than the velocity of cars in the system.
+This implies that traffic conditions are partially determined by what is going on behind.
+
 \subsection{Hugoniot Loci and Integral Curves}
 Brisa
+
+% AR Model
 \section{AR Model}
+\subsection{Theory}
+Another issues with the PW model, as Aw and Rascle point out in their paper, is that there is no conservation of momentum in car traffic.
+Instead, the term involving pressure is mean to represent an anticipation factor, or how a driver would react to a variation in the concentration of cars with respect to space.
+However, what we should really be concerned with is the perspective of the driver, which is relative to a moving timeframe.
+In this case, we should be using a convective derivative, $\partial_t + v \partial_x$, of the pressure.
+The new model becomes:
 
-\subsection{Theory}
-Kelsey
+\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho v)_x = 0 \\[1ex] & (v + p)_t + v (v + p)_x = 0 \end{matrix*} \right. \]
+
+The second equation can also be rewritten as an equation for velocity: $v_t + \Big(v - \rho p'(\rho)\Big)v_x = 0$. \\
+
+This updated model no longer conserves momentum.
+Instead, it conserves the quantity $y = v + p(\rho)$, which we can see if we rearrange the second equation to a conservation form: $(\rho (v + p(\rho)))_t + (\rho v(v + p(\rho)))_x = 0$.
+While the authors admit that there is no obvious physical interpretation of this quantity, they claim that it is the most natural pair of conservative variables.
+\\
+
+The authors also claim that their improved model satisfies the following important conditions:
+\begin{enumerate}
+\item The system is hyperbolic.
+\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity must remain nonnegative and bounded from above.
+\item When solving the Riemann problem, no waves connecting any state to its left (behind it) can have a propagation speed greater than the velocity $v$.
+\item The solution to the Riemann problem must agree with qualitative properties that drivers actually experience: braking produces shock waves and acceleration produces rarefaction waves.
+\item Near a vacuum, the solution to the Riemann problem must be sensitive to the data.
+\end{enumerate}
+
 \subsection{Hugoniot Loci and Integral Curves}

### 2015-03-04 FirstDraftMaterials/Draft.tex
Moved Kelsey's stuff from outline to draft

@@
 \usepackage{subfigure}
+\usepackage{enumerate}
 \setlength\parindent{0pt}
@@
 
+% Introduction
 \section{Introduction}
-Kelsey
+When a model is called ``higher order", we usually associate it with being more accurate than a first order model.
+This is not always the case.
+Adding complexity to a model does not necessarily improve it, as is demonstrated in the class of higher order traffic models developed with the intention of improving the first order traffic model proposed by Lighthill and Whitham (1955) and Richards (1956).
+As Daganzo observes in his paper, \\
+
+``A new theory (e.g.
+relativity) is deemed successful if it explains previously unexplained phenomena, plus everything else that was correctly explained with the theory it intends to replace (e.g.
+Newtonian physics).
+In the case of traffic flow, the new theories (high-order fluid models) fail to explain the behavior of traffic at the end of a queue (unlike the simpler, older LWR model; which explains it perfectly)." \\
+
+In this paper we examine the second order traffic model developed by Payne (1971) and Whitham (1974) which was meant to improve how the LWR model explains traffic behavior near shocks.
+Their model takes its form from equations used to describe compressible fluid dynamics.
+Unfortunately, cars and fluid particles have some fundamental differences which make the model actually perform worse than the LWR model in certain cases.
+For example, by smoothing out discontinuities, the PW model predicts negative velocities, which is clearly a problem.
+\\
+
+In addition to considering the PW model, we also take a look a more successful model proposed by Aw and Rascle (2000).
+These authors challenge some of the assumptions of the PW model and suggest the use of a convective derivative to fix the problem of negative velocities.
+In order to verify that their model is valid, they test their new model against criteria they claim any reasonable model will satisfy.
+
+% LWR Model
 \section{LWR Model}
-Kelsey
+We can start by considering a first order (one equation) model of traffic flow on a one-way highway without any exits, entrances, or passing.
+In this case, we can expect that the number of cars will be conserved, so we can use the equation for conservation of mass (with density $\rho$ and velocity $v$):
+
+\begin{enumerate}[(a)]
+\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho \, dt = -\Big[ f( \rho ) \Big]_{x_1}^{x_2}$
+\item Differential form of equation: $\rho_t + f( \rho )_x = 0$
+\begin{enumerate}[i]
+\item Light traffic: $\rho_t + (\rho v)_x = 0$
+\item Congested traffic: $\rho_t + \Big( \rho V(\rho) \Big)_x = 0$ (LWR model)
+\end{enumerate}
+\end{enumerate}
+
+The LWR behaves well macroscopically, but it does have some flaws.
+For example, it assumes cars change their velocity instantaneously as they travel through shocks.
+One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big( \rho V(\rho) \Big)_x = \mu \rho_{xx}$.
+This term is meant to account for the drivers awareness of the road ahead.
+However, this model can be shown to predict negative velocities in some instances, which is obviously wrong.
+
+% PW Model
 \section{PW Model}
 
+% PW Model: Theory
 \subsection{Theory}
-Kelsey
+Another approach to improving the LWR model would be to use a second order (two equation) model.
+Since cars can be thought of as a compressible gas, we can use equations for conservation of mass (above) and conservation of momentum, $\rho v$, used to model compressible gas flow:
+
+\begin{enumerate}[(a)]
+\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho v \, dt = -\Big[ f( \rho v ) \Big]_{x_1}^{x_2}$
+\item Differential form of equation: $(\rho v)_t + f( \rho v )_x = (\rho v)_t + \Big( \rho v^2 + \rho \Big)_x = 0$
+\end{enumerate}
+
+We can rewrite the second equation in terms of velocity instead of momentum: $v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = 0$. \\
+
+The PW model is a modification of these equations for conservation of mass and conservation of momentum in gas dynamics, with added relaxation and diffusion terms:
+
+\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho v)_x = 0 \\[1ex] & v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = \dfrac{V(\rho) - v}{\tau} + \mu v_{xx} \end{matrix*} \right. \]
+
+At first glance, this seems like a reasonable model.
+However, there a couple of fundamental differences between gas particles and cars that could be problematic.
+For example, as pointed out by Daganzo, ``a fluid particle responds to stimuli from the front and from behind, but a car is an anisotropic particle that mostly responds to frontal stimuli."  If we consider the eigenvalues of the Jacobian matrix of the linearized system, we see that one of the eigenvalues, $\lambda_2 = v + \sqrt{p'(\rho)}$, represents information traveling faster than the velocity of cars in the system.
+This implies that traffic conditions are partially determined by what is going on behind.
+
+% PW Model: Hugoniot Loci and Integral Curves
 \subsection{Hugoniot Loci and Integral Curves}
 Brisa
+
+% AR Model
 \section{AR Model}
 
+% AR Model: Theory
 \subsection{Theory}
-Kelsey
+Another issues with the PW model, as Aw and Rascle point out in their paper, is that there is no conservation of momentum in car traffic.
+Instead, the term involving pressure is mean to represent an anticipation factor, or how a driver would react to a variation in the concentration of cars with respect to space.
+However, what we should really be concerned with is the perspective of the driver, which is relative to a moving timeframe.
+In this case, we should be using a convective derivative, $\partial_t + v \partial_x$, of the pressure.
+The new model becomes:
+
+\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho v)_x = 0 \\[1ex] & (v + p)_t + v (v + p)_x = 0 \end{matrix*} \right. \]
+
+The second equation can also be rewritten as an equation for velocity: $v_t + \Big(v - \rho p'(\rho)\Big)v_x = 0$. \\
+
+This updated model no longer conserves momentum.
+Instead, it conserves the quantity $y = v + p(\rho)$, which we can see if we rearrange the second equation to a conservation form: $(\rho (v + p(\rho)))_t + (\rho v(v + p(\rho)))_x = 0$.
+While the authors admit that there is no obvious physical interpretation of this quantity, they claim that it is the most natural pair of conservative variables.
+\\
+
+The authors also claim that their improved model satisfies the following important conditions:
+\begin{enumerate}
+\item The system is hyperbolic.
+\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity must remain nonnegative and bounded from above.
+\item When solving the Riemann problem, no waves connecting any state to its left (behind it) can have a propagation speed greater than the velocity $v$.
+\item The solution to the Riemann problem must agree with qualitative properties that drivers actually experience: braking produces shock waves and acceleration produces rarefaction waves.
+\item Near a vacuum, the solution to the Riemann problem must be sensitive to the data.
+\end{enumerate}
+
+% AR Model: Hugoniot Loci and Integral Curves
 \subsection{Hugoniot Loci and Integral Curves}
@@
 
+% Examples
 \section{Examples}

### 2015-03-04 MatlabCode/AW_CurvesAndLamdbaContours.m
Adding updated Matlab code, and new images.

@@ -1,0 +1,80 @@
+% This function generates plots for the Hugoniot loci and Integral Curves
+% for the Aw Rascle model, and overlays contour lines for lambda1 and 
+% lambda2.
+
+clear all
+close all
+clc
+
+global size_font
+run variableSetup
+
+size_font = 25;
+validity = true;
+
+%% Setting up curves in the U = (rho, v) plane
+
+% Calling function to generate figures
+Ufigs = UPlot(1,validity);
+
+% Exporting figures
+figure(Ufigs(1))
+cd export_fig_code
+export_fig( gcf, ...      % figure handle
+    '../Images/Validity_U',... % name of output file without extension
+    '-painters', ...      % renderer
+    '-jpg', ...           % file format
+    '-r72' );             % resolution in dpi
+cd ..
+
+figure(Ufigs(2))
+cd export_fig_code
+export_fig( gcf, ...      % figure handle
+    '../Images/Validity_U_ql',... % name of output file without extension
+    '-painters', ...      % renderer
+    '-jpg', ...           % file format
+    '-r72' );             % resolution in dpi
+cd ..
+
+figure(Ufigs(3))
+cd export_fig_code
+export_fig( gcf, ...      % figure handle
+    '../Images/Validity_U_qr',... % name of output file without extension
+    '-painters', ...      % renderer
+    '-jpg', ...           % file format
+    '-r72' );             % resolution in dpi
+cd ..
+
+%% Setting up curves in the M = (rho,rho*v) plane
+
+% Calling function to generate figures
+Mfigs = MPlot(1,validity);
+
+% Exporting figues
+figure(Mfigs(1))
+cd export_fig_code
+export_fig( gcf, ...      % figure handle
+    '../Images/Validity_M',... % name of output file without extension
+    '-painters', ...      % renderer
+    '-jpg', ...           % file format
+    '-r72' );             % resolution in dpi
+cd ..
+
+figure(Mfigs(2))
+cd export_fig_code
+export_fig( gcf, ...      % figure handle
+    '../Images/Validity_M_ql',... % name of output file without extension
+    '-painters', ...      % renderer
+    '-jpg', ...           % file format
+    '-r72' );             % resolution in dpi
+cd ..
+
+figure(Mfigs(3))
+cd export_fig_code
+export_fig( gcf, ...      % figure handle
+    '../Images/Validity_M_qr',... % name of output file without extension
+    '-painters', ...      % renderer
[…]

### 2015-03-04 MatlabCode/AwRascle_LociAndCurves.m

@@ -1,54 +1,25 @@
-% This function plots the Hugoniot loci and Integral Curves
+% This function generates plots for the Hugoniot loci and Integral Curves
 % for the Aw Rascle model.
 % Note that the loci and the integral curves coincide for this model,
 % so only two equations are needed.
 
 clear all
 close all
-clc 
+clc
 
-global v_star rho_star p pprime rho
-variableSetup();
+global size_font
+run variableSetup
 
-plotLambdaContours = 0;
-fontSize = 25;
+size_font = 25;
+validity = false;
 
 %% Setting up curves in the U = (rho, v) plane
-for h = 1:length(v_star)
-    for k = 1:length(rho)
-        v1(h,k) = v_star(h) + p(rho_star(h)) - p(rho(k));
-        v2(h,k) = v_star(h);
-    end
-end
 
-% Creating contours for lambda1 = v - p'(rho)*rho
-lambda = [.1 .6 1.1];
-for h = 1:length(lambda)
-    for k = 1:length(rho)
-        contour2(h,k) = lambda(h) + pprime(rho(k))*rho(k);
-    end
-end
+% Calling function to generate figures
+Ufigs = UPlot(1, validity);
 
-% Plotting 1-curves
-figure(1)
-plot(rho,v1,'k')
-if (plotLambdaContours)
-    hold on
-    plot(rho,contour2(1,:),'g')
-    plot(rho,contour2(2,:),'r')
-    plot(rho,contour2(3,:),'b')
-end
-
-axis([0 1 0 1])
-set(gca,'XTick',[0 1])
-set(gca,'YTick',[0 1])
-
-hXLabel = xlabel('\rho');
-hYLabel = ylabel('v');
-set([hXLabel, hYLabel], 'FontSize', fontSize);
-set(gca,'FontSize',fontSize)
-set(gcf, 'Color', 'w');
-
+% Exporting figures
+figure(Ufigs(1))
 cd export_fig_code
 export_fig( gcf, ...      % figure handle
     '../Images/HLIC_U_lamb1',... % name of output file without extension
@@ -57,20 +28,7 @@ export_fig( gcf, ...      % figure handle
     '-r72' );             % resolution in dpi
 cd ..
 
-% Plotting 2-curves 
-figure(2)
[…]

### 2015-03-04 MatlabCode/MPlot.m

@@ -1,0 +1,123 @@
+% Function to generate plots in the M plane for the AR and PW model.
+% Inputs: model - 1 means AR model, 2 means PW model.
+%         validity - boolean indicating whether or not validity is being
+%                    considered
+% Generates 3 plots: curves for lambda1, curves for lambda2, and both
+% curves for a given point.
+% Note: expects that the function "variableSetup" has been run.
+
+function Mfigs = MPlot(model,validity)
+
+global v_star rho_star p pprime rho size_font
+
+if (model == 1)
+    
+    % Setting up equations
+    for h = 1:length(v_star)
+        for k = 1:length(rho)
+            rv1(h,k) = rho(k)*(v_star(h) + p(rho_star(h)) - p(rho(k)));
+            rv2(h,k) = v_star(h)*rho(k);
+        end
+    end
+    
+    % Plotting 1-curves
+    Mfigs(1) = figure;
+    plot(rho,rv1,'k')
+    
+    axis([0 5 0 1])
+    set(gca,'XTick',[0 5])
+    set(gca,'YTick',[0 1])
+    
+    state = 1;
+    if (~validity)
+        % Plotting 2-curves
+        Mfigs(2) = figure;
+        plot(rho,rv2,'b')
+        
+        axis([0 5 0 1])
+        set(gca,'XTick',[0 5])
+        set(gca,'YTick',[0 1])
+        
+        % Plotting both curves for a given initial state
+        Mfigs(3) = figure;
+        plot(rho,rv1(state,:),'k')
+        hold on
+        plot(rho,rv2(state,:),'b')
+        plot(rho_star(state),...
+            rho_star(state)*v_star(state),'ro', 'LineWidth', 2)
+        
+        axis([0 .7 0 .1])
+        set(gca,'XTick',[0 .69])
+        set(gca,'XTickLabel',{'0' '.7'})
+        set(gca,'YTick',[0 .1])
+        set(gca,'YTickLabel',{'0' '.1'})
+    else
+        % Creating contours for lambda1 = v - p'(rho)*rho
+        % Solving for v gives v = lambda1 + p'(rho)*rho
+        % So, rho*v = rho*(lambda1 + p'(rho)*rho)
+        lambda = [0 0.3 0.6];
+        for h = 1:length(lambda)
+            for k = 1:length(rho)
+                contourM(h,k) = rho(k)*(lambda(h) + pprime(rho(k))*rho(k));
+            end
+        end
+        
+        hold on
[…]

### 2015-03-04 MatlabCode/UPlot.m

@@ -1,0 +1,116 @@
+% Function to generate plots in the U plane for the AR and PW model. 
+% Inputs: model - 1 means AR model, 2 means PW model.
+%         validity - boolean indicating whether or not validity is being
+%                    considered
+% Generates 3 plots: curves for lambda1, curves for lambda2, and both
+% curves for a given point.
+% Note: expects that the function "variableSetup" has been run.
+
+function Ufigs = UPlot(model,validity)
+
+global v_star rho_star p pprime rho size_font 
+
+if (model == 1)
+    
+    % Setting up equations
+    for h = 1:length(v_star)
+        for k = 1:length(rho)
+            v1(h,k) = v_star(h) + p(rho_star(h)) - p(rho(k));
+            v2(h,k) = v_star(h);
+        end
+    end
+    
+    % Plotting 1-curves
+    Ufigs(1) = figure;
+    plot(rho,v1,'k')
+    
+    axis([0 1 0 1])
+    set(gca,'XTick',[0 1])
+    set(gca,'YTick',[0 1])
+    state = 1;
+
+    if (~validity)
+        % Plotting 2-curves
+        Ufigs(2) = figure;
+        plot(rho,v2,'b')
+        
+        axis([0 1 0 1])
+        set(gca,'XTick',[0 1])
+        set(gca,'YTick',[0 1])
+        
+        % Plotting both curves for a given initial state
+        Ufigs(3) = figure;
+        plot(rho,v1(state,:),'k')
+        hold on
+        plot(rho,v2(state,:),'b')
+        plot(rho_star(state),v_star(state),'ro', 'LineWidth', 2)
+    
+        axis([0 1 0 1])
+        set(gca,'XTick',[0 1])
+        set(gca,'YTick',[0 1])
+    else
+        % Creating contours for lambda1 = v - p'(rho)*rho
+        % Solving for v gives v = lambda1 + p'(rho)*rho
+        lambda = [0 0.3 0.6];
+        for h = 1:length(lambda)
+            for k = 1:length(rho)
+                contourU(h,k) = lambda(h) + pprime(rho(k))*rho(k);
+            end
+        end
+        
+        hold on
+        plot(rho,contourU,'m')
+        text(0.8,0.25,'\lambda_1 = 0','FontSize',size_font)
+        text(0.8,0.55,'\lambda_1 = 0.3','FontSize',size_font)
+        text(0.8,0.85,'\lambda_1 = 0.6','FontSize',size_font)
+
+        % Finding index for given point 
[…]

### 2015-03-04 MatlabCode/YPlot.m

@@ -1,0 +1,72 @@
+% Function to generate plots in the Y plane for the AR and PW model.
+% Input: 1 = AR model, 2 = PW model.
+% Generates 3 plots: curves for lambda1, curves for lambda2, and both
+% curves for a given point.
+% Note: expects that the function "variableSetup" has been run.
+
+function Yfigs = YPlot(model)
+
+global v_star rho_star p rho size_font
+
+if (model == 1)
+    
+    % Setting up equations
+    for h = 1:length(v_star)
+        for k = 1:length(rho)
+            y1(h,k) = rho(k)*v_star(h) + rho(k)*p(rho_star(h));
+            y2(h,k) = rho(k)*v_star(h) + rho(k)*p(rho(k));
+        end
+    end
+    
+    % Plotting 1-curves
+    Yfigs(1) = figure;
+    plot(rho,y1,'k')
+    
+    axis([0 0.1 0 0.1])
+    set(gca,'XTick',[0 0.1])
+    set(gca,'XTickLabel',{'0' '.1'})
+    set(gca,'YTick',[0 0.1])
+    set(gca,'YTickLabel',{'0' '.1'})
+    
+    % Plotting 2-curves
+    Yfigs(2) = figure;
+    plot(rho,y2,'b')
+    
+    axis([0 0.1 0 0.1])
+    set(gca,'XTick',[0 0.1])
+    set(gca,'XTickLabel',{'0' '.1'})
+    set(gca,'YTick',[0 0.1])
+    set(gca,'YTickLabel',{'0' '.1'})
+    
+    % Plotting both curves for a given initial state
+    Yfigs(3) = figure;
+    state = 1;
+    plot(rho,y1(state,:),'k')
+    hold on
+    plot(rho,y2(state,:),'b')
+    y_star = rho_star(state)*v_star(state)...
+        + rho_star(state)*p(rho_star(state));
+    plot(rho_star(state),y_star,'ro', 'LineWidth', 2)
+    
+    axis([0 0.4 0 0.3])
+    set(gca,'XTick',[0 0.4])
+    set(gca,'XTickLabel',{'0' '.4'})
+    set(gca,'YTick',[0 0.3])
+    set(gca,'YTickLabel',{'0' '.3'})
+    
+elseif (model == 2)
+    % TODO: Add code for PW model
+else
+    display('Model number must be either 1 or 2')
+end
+
+
+% Formatting figures
[…]

### 2015-03-04 MatlabCode/variableSetup.m

@@ -1,5 +1,3 @@
-function y = variableSetup
-
 global v_star rho_star p pprime rho
 
 % Declaring various starting values

### 2015-03-04 FirstDraftMaterials/Draft.tex
Finished AR loci and integral curves section.

@@
 &\partial_t\rho + \partial_x(\rho v) = 0, \label{AW:eq1}\\
-&\partial_t \left(v + p(\rho )\right) + v\partial_x \left( v + p(\rho )\right) = 0.
+&\partial_t \left(v + p(\rho )\right) + v\partial_x \left( v + p(\rho )\right) = 0\label{AW:eq1.5}
 \end{align}
+with $p(\rho) = \rho^\gamma$.
 
@@
 &\rho_*(s - v_*) = \rho (s - v), \label{AW:eq5}\\
-&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).
+&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).\label{AW:eq6}
 \end{align}
@@
 \end{align*}
-(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, solving (\ref{AW:eq3}) for $s$ gives
+(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AW:eq6}) gives
+\begin{align*}
+\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
+\rho(s - v)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
+v_* + p(\rho_* )&= v + p(\rho ).
+\end{align*}
+Therefore, the Hugoniot loci for this case in the Y plane are given by
+\begin{align*}
+y = v_* + p(\rho_* ),
+\end{align*}
+since $y = v + p(\rho )$ by definition. Note that solving (\ref{AW:eq3}) for $s$ gives
 \begin{align*}
@@
 \end{align*}
+Using the fact that $v_* = v + p(\rho) - p(\rho_*)$, we get that
+\begin{align*}
+s = \frac{\rho_*\left( v + p(\rho) - p(\rho_*)\right)- \rho v}{\rho_* - \rho}
+= v - \rho_*\frac{p(\rho ) - p(\rho_*)}{\rho - \rho_*}
+\approx v - \rho p'(\rho)
+\end{align*}
+for $\rho_* \approx \rho$.
+
+In the $U = (\rho, v)$ plane these loci become (a) $v = v_*$, $\rho$ free, and (b) $v = v_* + p(\rho_* ) - p(\rho )$.
+ In the $M = (\rho, \rho v)$ plane these loci become (a) $\rho v = \rho_* v_*$, $\rho$ free, 
+ and (b) $\rho v = \rho \left( v_* + p(\rho_*)\right) - \rho p(\rho))$. Figure \ref{fig:AR_curves} shows the Hugoniot loci 
+ in the different planes. TODO: ADD PARAMETERS USED FOR FIGURES? OR, WILL INCLUDING THE CODE MAKE THAT REDUNDANT? 
+ IF WE ADD PARAMETERS, MAYBE THE BEST PRESENTATION WOULD BE IN A TABLE?
 
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U_lamb1.jpg}
-   \label{fig:subfig1}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U_lamb2.jpg}
-   \label{fig:subfig2}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U.jpg}
-   \label{fig:subfig3}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb1.jpg}
-   \label{fig:subfig4}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb2.jpg}
-   \label{fig:subfig5}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M.jpg}
-   \label{fig:subfig6}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y_lamb1.jpg}
-   \label{fig:subfig7}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y_lamb2.jpg}
-   \label{fig:subfig8}
    }
@@
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y.jpg}
-   \label{fig:subfig9}
    }
- \label{fig:11_8c}
+ \label{fig:AR_curves}
  \caption[Optional caption for list of figures]
  {Figures of the Hugonio loci and integral curves for $\lambda_1$ and $\lambda_2$ 
- in each of the different planes. Since the loci and integral curves coincide, they are both
-  simply called the ``curves'' in the captions of the above figures. Figures (c), (f), (i) show
-   both curves for a given point in each plane.}
+ in each of the different planes. Since the loci and integral curves coincide, they are 
+ both simply called the ``curves'' in the captions of the above figures. 
+ Figures (c), (f), (i) show both curves for a given point in each plane.}
 \end{figure}
+
+\subsubsection{Integral Curves}
+In order to consider the integral curves for this system \cite{AwRascle2000} 
+works from the system
+\begin{align*}
+&\partial_t \rho + \partial_x (\rho v) = 0 \\ 
+&\partial_t v + \left(v - p'(\rho)\rho\right)\partial_x v = 0,
+\end{align*}
+where the second equation is found by multiplying (\ref{AW:eq1}) 
+by $p'(\rho)$ and subtracting it from (\ref{AW:eq1.5}). 
+This can be rewritten as $q_t + Aq_t = 0$ by defining
+\begin{align*}
+q = \left[ \begin{matrix}
+\rho \\ v
+\end{matrix}\right] , \hspace{0.3in}
+A = \left[ \begin{matrix}
+v & \rho \\
+0 & v - p'(\rho ) \rho
+\end{matrix}\right].
+\end{align*}
+The eigenvalues of $A$ are $\lambda_1 = v - p'(\rho ) \rho$ and 
+$\lambda_2 = v$. Note that since $p(\rho ) = \rho^{\gamma}$ 
+and $\rho \geq 0$, $\lambda_1 < \lambda_2$ so long as $\rho \neq 0$. 
+Also note that if $\rho = 0$ then $A$ is no longer 
+hyperbolic. The eigenvectors of $A$ are
+\begin{align*}
+r_1 = \left[ \begin{matrix}
+1 \\ - p'(\rho )
+\end{matrix}\right], \hspace{0.3in}
+r_2 = \left[ \begin{matrix}
+1 \\ 0
+\end{matrix}\right].
+\end{align*}
+Using equation (13.24) from \cite{LeVeque2002} with $\alpha = 1$ tells us 
+that the integral curves can be found by considering $q'(\xi ) = r_p$. 
+For $p = 1$ this gives us the two equations
+\begin{align*}
+&\rho'(\xi ) = 1 \\
+&v'(\xi ) = - p'(\rho ),
+\end{align*}
+which can be solved to find that we have parameratized our curve by $\rho$, 
+and $v = - p'(\rho ) + c_1$ while $\rho$ is free, where $c_1$ is a constant. 
+Since for a given point $(\rho_*, v_*)$ when $\rho = \rho_*$ we want 
+$v = v_*$, we get that $c_1 = v_* + p'(\rho_*)$. Thus, the integral curve 
+is given by 
+\begin{align*}
+v = v_* + p'(\rho_*) - p'(\rho ).
+\end{align*}
+This coincides with the Hugoniot loci from case (b). For $p = 2$, we get the 
+two equations
+\begin{align*}
+&\rho'(\xi ) = 1 \\
+&v'(\xi ) = 0,
+\end{align*}
+which can be solved to find that we have parameratized our curve by $\rho$, 
+and that $v$ is a constant. Therefore, $v = v_*$ and $\rho$ is free. 
+This coincides with the Hugoniot loci from case (a).
+
+\subsubsection{Regions of Validity}
+Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point.
+When traveling across a shock from a left state, $q_l$, to a right state $q_r$, $\lambda$ must decrease.
+When traveling across a rarefaction wave from a left state, $q_l$, to a right state $q_r$, $\lambda$ must increase.
+
+\begin{figure}[h!]
+ \centering
+ \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the U plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U.jpg}
+   }
+    \subfigure[Valid waves for the point $q_l$ in the U plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_ql.jpg}
+   }
+    \subfigure[Valid waves for the point $q_r$ in the U plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_qr.jpg}
+   }
+    \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the M plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M.jpg}
+   }
+ \subfigure[Valid waves for the point $q_l$ in the M plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_ql.jpg}
+   }
+ \subfigure[Valid waves for the point $q_r$ in the M plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_qr.jpg}
+   }
+ \label{fig:AW_validity}
+ \caption[Optional caption for list of figures]
+ {Figures showing the calculation of the valid waves for a given point. Figures (a)
+  and (d) show the $\lambda_1$ curves in black with the $\lambda_1$ contours 
+  in magenta. The other figures show valid shocks in red and valid rarefaction 
+  waves in green. The contact discontinuity is shown in blue.}
+\end{figure}
+
+Therefore, it is important to take into account the contour lines of $\lambda$ when 
+looking at the loci and integral curves. Figures \ref{fig:AW_validity}(a) and 
+\ref{fig:AW_validity}(d) show the $\lambda_1$-curves in the $U$ and $M$ planes,
+with the contour lines of $\lambda_1$ shown in magenta. This information is used
+to determine which areas of the $\lambda_1$ curves are valid shock waves and which
+are valid rarefaction waves. The rest of the subfigures show the regions of validity for 
+specific points $q_*$, depending on whether $q_* = q_l$ or $q_* = q_r$. For these plots,
+$q_* = (\rho_*,v_*) = (0.2, 0.2)$. Note that the contact discontinuity is valid
+anywhere in the plane.
 

### 2015-03-04 MatlabCode/MPlot.m
Fixing comments

@@ -3,7 +3,9 @@
 %         validity - boolean indicating whether or not validity is being
 %                    considered
 % Generates 3 plots: curves for lambda1, curves for lambda2, and both
-% curves for a given point.
+% curves for a given point if validity = false. curves for lambda1 with
+% contour lines overlayed, and valid shocks and rarefactions for ql and qr
+% if validity = true.
 % Note: expects that the function "variableSetup" has been run.
 
 function Mfigs = MPlot(model,validity)

### 2015-03-04 MatlabCode/UPlot.m

@@ -3,7 +3,9 @@
 %         validity - boolean indicating whether or not validity is being
 %                    considered
 % Generates 3 plots: curves for lambda1, curves for lambda2, and both
-% curves for a given point.
+% curves for a given point if validity = false. curves for lambda1 with
+% contour lines overlayed, and valid shocks and rarefactions for ql and qr
+% if validity = true.
 % Note: expects that the function "variableSetup" has been run.
 
 function Ufigs = UPlot(model,validity)

### 2015-03-05 MatlabCode/AR_examples.m
Updating Matlab code and adding AR examples.

@@ -1,0 +1,118 @@
+% Function to generate examples at the end of the Aw Rascle paper.
+
+clear all
+close all
+clc
+
+% Declaring various starting values
+gamma = 0.2;
+p = @(x) x^gamma;
+pprime = @(x) gamma*x^(gamma-1);
+rho = 0:0.001:5;
+size_font = 20;
+
+% Loop plots both examples
+for example = 1:2
+    figure
+    % Setting up data for the two points ql and qr
+    if (example == 1)
+        v_star = [0.2 .3];
+        rho_star = [0.3 .4];
+    else
+        v_star = [0 .3];
+        rho_star = [0 .4];
+    end
+    
+    % Setting up equations
+    for h = 1:length(v_star)
+        for k = 1:length(rho)
+            rv1(h,k) = rho(k)*(v_star(h) + p(rho_star(h)) - p(rho(k)));
+            rv2(h,k) = v_star(h)*rho(k);
+        end
+    end
+    
+    % Loop plots each of the two points ql and qr
+    for k = 1:length(rho_star)
+        state = k;
+        
+        % Finding index for given point
+        switch_index = find(rho==rho_star(state));
+        
+        x = rho_star(state) + 0.009;
+        y = rho_star(state)*v_star(state)-0.007;
+        if (k == 1)
+            plot(rho(1:switch_index),rv1(state,1:switch_index),'r')
+            hold on
+            plot(rho(switch_index:end),rv1(state,switch_index:end),'g')
+            text(x,y,'$m_l$','Interpreter', 'latex','FontSize',size_font)
+        else
+            plot(rho(1:switch_index),rv1(state,1:switch_index),'g')
+            hold on
+            plot(rho(switch_index:end),rv1(state,switch_index:end),'r')
+            text(x,y,'$m_r$','Interpreter', 'latex','FontSize',size_font)
+        end
+        
+        plot(rho,rv2(state,:),'b')
+        plot(rho_star(state),...
+            rho_star(state)*v_star(state),'ro', 'LineWidth', 2)
+    end
+    
+    % Adding middle state
+    if (example == 1)
+        syms x
+        eqn = x*(v_star(1) + p(rho_star(1)) - p(x)) ...
+            == (v_star(2)*x);
+
+        rho_m = vpasolve(eqn,[0.1;0.2]);
+        rhov_m = rho_m*v_star(2);
+        plot(rho_m, rhov_m,'ro', 'LineWidth', 2)
+        
+        x = double(rho_m) - 0.02;
[…]

### 2015-03-05 MatlabCode/MPlot.m

@@ -66,10 +66,9 @@ if (model == 1)
         
         hold on
         plot(rho,contourM,'m')
-        text(0.3,0.85,'\lambda_1 = 0','FontSize',size_font)
+        text(0.26,0.85,'\lambda_1 = 0','FontSize',size_font)
         text(1.8,0.85,'\lambda_1 = 0.3','FontSize',size_font)
         text(3.6,0.85,'\lambda_1 = 0.6','FontSize',size_font)
-        
         
         % Finding index for given point 
         switch_index = find(rho==rho_star(state));
@@ -95,9 +94,9 @@ if (model == 1)
             y = rho_star(state)*v_star(state)-0.005;
             
             if (k == 2)
-                text(x,y,'q_l','FontSize',size_font)
+                text(x,y,'$m_l$','Interpreter','latex','FontSize',size_font)
             else
-                text(x,y,'q_r','FontSize',size_font)
+                text(x,y,'$m_r$','Interpreter','latex','FontSize',size_font)
             end
             
             axis([0 .7 0 .1])
@@ -117,8 +116,8 @@ end
 % Formatting figures
 for k = 1:length(Mfigs)
     figure(Mfigs(k))
-    hXLabel = xlabel('\rho');
-    hYLabel = ylabel('m = \rho v');
+    hXLabel = xlabel('$\rho$','Interpreter','latex');
+    hYLabel = ylabel('$m = \rho v$','Interpreter','latex');
     set([hXLabel, hYLabel], 'FontSize', size_font);
     set(gca,'FontSize',size_font)
     set(gcf, 'Color', 'w');

### 2015-03-05 MatlabCode/UPlot.m

@@ -62,9 +62,9 @@ if (model == 1)
         
         hold on
         plot(rho,contourU,'m')
-        text(0.8,0.25,'\lambda_1 = 0','FontSize',size_font)
-        text(0.8,0.55,'\lambda_1 = 0.3','FontSize',size_font)
-        text(0.8,0.85,'\lambda_1 = 0.6','FontSize',size_font)
+        text(0.75,0.23,'\lambda_1 = 0','FontSize',size_font)
+        text(0.75,0.53,'\lambda_1 = 0.3','FontSize',size_font)
+        text(0.75,0.83,'\lambda_1 = 0.6','FontSize',size_font)
 
         % Finding index for given point 
         switch_index = find(rho==rho_star(state));
@@ -90,9 +90,9 @@ if (model == 1)
             y = v_star(state)+0.05;            
             
             if (k == 2)
-                text(x,y,'q_l','FontSize',size_font)
+                text(x,y,'$u_l$','Interpreter','latex','FontSize',size_font)
             else
-                text(x,y,'q_r','FontSize',size_font)
+                text(x,y,'$u_r$','Interpreter','latex','FontSize',size_font)
             end
             
             axis([0 1 0 1])
@@ -110,8 +110,8 @@ end
 % Formatting figures
 for k = 1:length(Ufigs)
     figure(Ufigs(k))
-    hXLabel = xlabel('\rho');
-    hYLabel = ylabel('v');
+    hXLabel = xlabel('$\rho$','Interpreter','latex');
+    hYLabel = ylabel('$v$','Interpreter','latex');
     set([hXLabel, hYLabel], 'FontSize', size_font);
     set(gca,'FontSize',size_font)
     set(gcf, 'Color', 'w');

### 2015-03-05 MatlabCode/YPlot.m

@@ -64,8 +64,8 @@ end
 % Formatting figures
 for k = 1:length(Yfigs)
     figure(Yfigs(k))
-    hXLabel = xlabel('\rho');
-    hYLabel = ylabel('y');
+    hXLabel = xlabel('$\rho$','Interpreter','latex');
+    hYLabel = ylabel('$y$','Interpreter','latex');
     set([hXLabel, hYLabel], 'FontSize', size_font);
     set(gca,'FontSize',size_font)
     set(gcf, 'Color', 'w');

### 2015-03-05 ClawpackCode/pw.f90
Added clawpack code for PW

@@ -1,0 +1,207 @@
+! =========================================================
+subroutine rp1(maxmx,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =========================================================
+
+! Solve Riemann problems for the 1D PW equations
+!   (rho)_t + (rho*v)_x = 0
+!   (rho*v)_t + (rho*v^2 + rho^gamma)_x = 0
+! using Roe's approximate Riemann solver with entropy fix for
+! transonic rarefactions.
+
+! waves: 2
+! equations: 2
+
+! Conserved quantities:
+!       1 density
+!       2 momentum
+
+! This function solves the Riemann problem at all interfaces in one call
+
+! On input, ql contains the state vector at the left edge of each cell
+!           qr contains the state vector at the right edge of each cell
+! On output, wave contains the waves,
+!            s the speeds,
+!            amdq the  left-going flux difference  A^- \Delta q
+!            apdq the right-going flux difference  A^+ \Delta q
+
+! Note that the i'th Riemann problem has left state qr(:,i-1)
+!                                    and right state ql(:,i)
+! From the basic clawpack routine step1, rp is called with ql = qr = q.
+
+
+    implicit double precision (a-h,o-z)
+
+    dimension   ql(meqn,1-mbc:maxmx+mbc)
+    dimension   qr(meqn,1-mbc:maxmx+mbc)
+    dimension    s(mwaves,1-mbc:maxmx+mbc)
+    dimension wave(meqn, mwaves,1-mbc:maxmx+mbc)
+    dimension amdq(meqn,1-mbc:maxmx+mbc)
+    dimension apdq(meqn,1-mbc:maxmx+mbc)
+
+!     # Local storage
+!     ---------------
+    dimension delta(2)
+    logical :: efix
+
+    data efix /.true./    !# use entropy fix for transonic rarefactions
+
+    gamma = 0.2d0
+
+!     # Main loop of the Riemann solver.
+    do 30 i=2-mbc,mx+mbc
+
+
+    !     # compute  Roe-averaged quantities:
+        rbar = 0.5d0*(qr(1,i-1) + ql(1,i))
+        cbar = dsqrt(gamma*rbar**(gamma-1.0d0))
+        if (rbar == 0) then
+            ubar = 0
+        else
[…]

### 2015-03-05 ClawpackCode/qinit.f90

@@ -1,0 +1,30 @@
+subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
+
+    ! Set initial conditions for the q array.
+    ! This default version prints an error message since it should
+    ! not be used directly.  Copy this to an application directory and
+    ! loop over all grid cells to set values of q(1:meqn, 1:mx).
+
+    implicit none
+    
+    integer, intent(in) :: meqn,mbc,mx,maux
+    real(kind=8), intent(in) :: xlower,dx
+    real(kind=8), intent(in) :: aux(maux,1-mbc:mx+mbc)
+    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc)
+
+    integer :: i
+    real(kind=8) :: xcell
+ 
+      do i=1,mx
+         xcell = xlower + (i-0.5d0)*dx
+        if (xcell.lt.0.0d0) then
+            q(1,i) = 0.01d0
+            q(2,i) = 0.01d0
+        else
+            q(1,i) = 0.4d0
+            q(2,i) = 0.12d0
+        endif
+      enddo
+
+end subroutine qinit
+

### 2015-03-05 ClawpackCode/setplot.py

@@ -1,0 +1,72 @@
+
+""" 
+Set up the plot figures, axes, and items to be done for each frame.
+
+This module is imported by the plotting routines and then the
+function setplot is called to set the plot parameters.
+    
+""" 
+
+#--------------------------
+def setplot(plotdata):
+#--------------------------
+    
+    """ 
+    Specify what is to be plotted at each frame.
+    Input:  plotdata, an instance of clawpack.visclaw.data.ClawPlotData.
+    Output: a modified version of plotdata.
+    
+    """ 
+
+
+    plotdata.clearfigures()  # clear any old figures,axes,items data
+
+    # Figure for q[0]
+    plotfigure = plotdata.new_plotfigure(name='density and momentum', figno=1)
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [-0.1,0.45]
+    plotaxes.title = 'density (rho)'
+
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 0
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+
+
+    # Figure for q[1]
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [-0.2,0.15]
+    plotaxes.title = 'momentum (rho*v)'
+
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 1
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+
+    # Parameters used only when creating html and/or latex hardcopy
+    # e.g., via clawpack.visclaw.frametools.printframes:
+
+    plotdata.printfigs = True                # print figures
+    plotdata.print_format = 'png'            # file format
+    plotdata.print_framenos = 'all'          # list of frames to print
+    plotdata.print_fignos = 'all'            # list of figures to print
+    plotdata.html = True                     # create html files of plots?
+    plotdata.html_homelink = '../README.html'
+    plotdata.latex = True                    # create latex file of plots?
[…]

### 2015-03-05 ClawpackCode/setrun.py

@@ -1,0 +1,217 @@
+""" 
+Module to set up run time parameters for Clawpack -- classic code.
+
+The values set in the function setrun are then written out to data files
+that will be read in by the Fortran code.
+    
+""" 
+
+import os
+import numpy as np
+
+#------------------------------
+def setrun(claw_pkg='classic'):
+#------------------------------
+    
+    """ 
+    Define the parameters used for running Clawpack.
+
+    INPUT:
+        claw_pkg expected to be "classic" for this setrun.
+
+    OUTPUT:
+        rundata - object of class ClawRunData 
+    
+    """ 
+    
+    from clawpack.clawutil import data 
+    
+    
+    assert claw_pkg.lower() == 'classic',  "Expected claw_pkg = 'classic'"
+
+    num_dim = 1
+    rundata = data.ClawRunData(claw_pkg, num_dim)
+
+
+    #------------------------------------------------------------------
+    # Standard Clawpack parameters to be written to claw.data:
+    #------------------------------------------------------------------
+
+    clawdata = rundata.clawdata  # initialized when rundata instantiated
+
+
+    # ---------------
+    # Spatial domain:
+    # ---------------
+
+    # Number of space dimensions:
+    clawdata.num_dim = num_dim
+    
+    # Lower and upper edge of computational domain:
+    clawdata.lower[0] = -1.000000e+00          # xlower
+    clawdata.upper[0] = 1.000000e+00          # xupper
+    
+    # Number of grid cells:
+    clawdata.num_cells[0] = 400      # mx
+    
+
+    # ---------------
+    # Size of system:
+    # ---------------
+
+    # Number of equations in the system:
+    clawdata.num_eqn = 2
+
+    # Number of auxiliary variables in the aux array (initialized in setaux)
+    clawdata.num_aux = 0
+    
+    # Index of aux array corresponding to capacity function, if there is one:
+    clawdata.capa_index = 0
+    
+    
+    # -------------
+    # Initial time:
+    # -------------
+
+    clawdata.t0 = 0.000000
+    
+
+    # Restart from checkpoint file of a previous run?
+    # Note: If restarting, you must also change the Makefile to set:
+    #    RESTART = True
+    # If restarting, t0 above should be from original run, and the
+    # restart_file 'fort.qNNNN' specified below should be in 
+    # the OUTDIR indicated in Makefile.
+
[…]

### 2015-03-05 FirstDraftMaterials/Draft.tex

@@
 \begin{align}
-&\partial_t\rho + \partial_x(\rho v) = 0, \label{AW:eq1}\\
-&\partial_t \left(v + p(\rho )\right) + v\partial_x \left( v + p(\rho )\right) = 0\label{AW:eq1.5}
+&\partial_t\rho + \partial_x(\rho v) = 0, \label{AR:eq1}\\
+&\partial_t \left(v + p(\rho )\right) + v\partial_x \left( v + p(\rho )\right) = 0\label{AR:eq1.5}
 \end{align}
@@
 Adding these two equations gives
-\begin{align}\label{AW:eq2}
+\begin{align}\label{AR:eq2}
 \partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
 \end{align}
-Then from (\ref{AW:eq1}) and (\ref{AW:eq2}) we have the system 
+Then from (\ref{AR:eq1}) and (\ref{AR:eq2}) we have the system 
 \begin{align*}
@@
 \begin{align}
-s(\rho_* - \rho) = v_*\rho_* - v\rho, \label{AW:eq3}\\
-s(y_* - y) = v_*y_* - vy.\label{AW:eq4}
+s(\rho_* - \rho) = v_*\rho_* - v\rho, \label{AR:eq3}\\
+s(y_* - y) = v_*y_* - vy.\label{AR:eq4}
 \end{align}
@@
 \begin{align}
-&\rho_*(s - v_*) = \rho (s - v), \label{AW:eq5}\\
-&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).\label{AW:eq6}
+&\rho_*(s - v_*) = \rho (s - v), \label{AR:eq5}\\
+&\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) = \rho(s - v)\left(v + p(\rho )\right).\label{AR:eq6}
 \end{align}
-From (\ref{AW:eq5}), we know that there are two cases:
+From (\ref{AR:eq5}), we know that there are two cases:
 
@@
 \end{align*}
-Since the velocity does not change, this is a contact discontinuity. Using the fact that $v = v_*$ in (\ref{AW:eq4}) gives
+Since the velocity does not change, this is a contact discontinuity. Using the fact that $v = v_*$ in (\ref{AR:eq4}) gives
 \begin{align*}
@@
 \end{align*}
-(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AW:eq6}) gives
+(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AR:eq6}) gives
 \begin{align*}
@@
 \end{align*}
-since $y = v + p(\rho )$ by definition. Note that solving (\ref{AW:eq3}) for $s$ gives
+since $y = v + p(\rho )$ by definition. Note that solving (\ref{AR:eq3}) for $s$ gives
 \begin{align*}
@@
    }
- \label{fig:AR_curves}
  \caption[Optional caption for list of figures]
@@
  Figures (c), (f), (i) show both curves for a given point in each plane.}
+  \label{fig:AR_curves}
 \end{figure}
@@
 \end{align*}
-where the second equation is found by multiplying (\ref{AW:eq1}) 
-by $p'(\rho)$ and subtracting it from (\ref{AW:eq1.5}). 
+where the second equation is found by multiplying (\ref{AR:eq1}) 
+by $p'(\rho)$ and subtracting it from (\ref{AR:eq1.5}). 
 This can be rewritten as $q_t + Aq_t = 0$ by defining
@@
    }
- \label{fig:AW_validity}
  \caption[Optional caption for list of figures]
@@
   waves in green. The contact discontinuity is shown in blue.}
+   \label{fig:AR_validity}
 \end{figure}
@@
 Therefore, it is important to take into account the contour lines of $\lambda$ when 
-looking at the loci and integral curves. Figures \ref{fig:AW_validity}(a) and 
-\ref{fig:AW_validity}(d) show the $\lambda_1$-curves in the $U$ and $M$ planes,
+looking at the loci and integral curves. Figures \ref{fig:AR_validity}(a) and 
+\ref{fig:AR_validity}(d) show the $\lambda_1$-curves in the $U$ and $M$ planes,
 with the contour lines of $\lambda_1$ shown in magenta. This information is used
@@
 \section{Examples}
-Brisa
+
+\begin{figure}[h!]
+ \centering
+ \subfigure[Solution from the AR model.]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
+   }
+% TODO: add subfigure for example1 in PW
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 1 from the AR and PW models.}
+  \label{fig:example1}
+\end{figure}
+
+\begin{figure}[h!]
+ \centering
+ \subfigure[Solution from the AR model.]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
+   }
+% TODO: add subfigure for example2 in PW
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 2 from the AR and PW models.}
+  \label{fig:example2}
+\end{figure}
+
+\begin{table}[t]
+\caption{Initial values used in examples 1 and 2.}
+\begin{center}
+\begin{tabular}{| c | c c |}
+\hline
+& Example 1 & Example 2\\
+\hline
+$v_l$ & 0.2 & 0 \\
+$\rho_l$ & 0.3 & 0\\
+$v_r$ & 0.3 & 0 .3\\
+$\rho_r$ & 0.4 & 0.4\\
+\hline
+\end{tabular}
+\end{center}
+ \label{table:1}
+\end{table}
+
+Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples.
+The variables used in these two examples are shown in Table \ref{table:1}.
+Note that in example 1, shown in Figure \ref{fig:example1}, the transition from $m_l = (\rho_l, v_l\rho_l)$ to $m_r = (\rho_r, v_r\rho_r)$ is first a shock and then a contact discontinuity.
+The middle state is denoted by $\tilde{m}$ in the figure.
+In example 2, show in Figure \ref{fig:example2}, the transition from $m_l = (\rho_l, v_l\rho_l)$ to $m_r = (\rho_r, v_r\rho_r)$ is a rarefaction wave, and there is no middle state.
 

### 2015-03-05 ClawpackCode/pw.f90
Fixed pw.f90

@@ -54,26 +54,16 @@ subroutine rp1(maxmx,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
     !     # compute  Roe-averaged quantities:
         rbar = 0.5d0*(qr(1,i-1) + ql(1,i))
         cbar = dsqrt(gamma*rbar**(gamma-1.0d0))
-        if (rbar == 0) then
-            ubar = 0
-        else
-            ubar = (qr(2,i-1)*dsqrt(qr(1,i-1)) + ql(2,i)*dsqrt(ql(1,i)))/ &
+        ubar = (qr(2,i-1)/dsqrt(qr(1,i-1)) + ql(2,i)/dsqrt(ql(1,i)))/ &
             ( dsqrt(qr(1,i-1)) + dsqrt(ql(1,i)) )
-        endif
-
 
     !     # delta(1)=rho(i)-rho(i-1) and  delta(2)=rho*v(i)-rho*v(i-1)
         delta(1) = ql(1,i) - qr(1,i-1)
         delta(2) = ql(2,i) - qr(2,i-1)
 
     !     # Compute coeffs in the evector expansion of delta(1),delta(2)
-        if (cbar == 0) then
-            a1 = 0
-            a2 = 0
-        else
-            a1 = 0.5d0*(-delta(2) + (ubar + cbar) * delta(1))/cbar
-            a2 = 0.5d0*( delta(2) - (ubar - cbar) * delta(1))/cbar
-        endif
+        a1 = 0.5d0*(-delta(2) + (ubar + cbar) * delta(1))/cbar
+        a2 = 0.5d0*( delta(2) - (ubar - cbar) * delta(1))/cbar
 
     !     # Finally, compute the waves.
         wave(1,1,i) = a1

### 2015-03-05 ClawpackCode/setplot.py

@@ -44,7 +44,7 @@ def setplot(plotdata):
     plotaxes = plotfigure.new_plotaxes()
     plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
     plotaxes.xlimits = 'auto'
-    plotaxes.ylimits = [-0.2,0.15]
+    plotaxes.ylimits = [-0.15,0.15]
     plotaxes.title = 'momentum (rho*v)'
 
     # Set up for item on these axes:

### 2015-03-05 MatlabCode/AR_examples.m
Fixing error that switched shocks and rarefactions in M plane.

@@ -41,14 +41,14 @@ for example = 1:2
         x = rho_star(state) + 0.009;
         y = rho_star(state)*v_star(state)-0.007;
         if (k == 1)
-            plot(rho(1:switch_index),rv1(state,1:switch_index),'r')
-            hold on
-            plot(rho(switch_index:end),rv1(state,switch_index:end),'g')
-            text(x,y,'$m_l$','Interpreter', 'latex','FontSize',size_font)
-        else
             plot(rho(1:switch_index),rv1(state,1:switch_index),'g')
             hold on
             plot(rho(switch_index:end),rv1(state,switch_index:end),'r')
+            text(x,y,'$m_l$','Interpreter', 'latex','FontSize',size_font)
+        else
+            plot(rho(1:switch_index),rv1(state,1:switch_index),'r')
+            hold on
+            plot(rho(switch_index:end),rv1(state,switch_index:end),'g')
             text(x,y,'$m_r$','Interpreter', 'latex','FontSize',size_font)
         end
         

### 2015-03-05 MatlabCode/MPlot.m

@@ -57,7 +57,7 @@ if (model == 1)
         % Creating contours for lambda1 = v - p'(rho)*rho
         % Solving for v gives v = lambda1 + p'(rho)*rho
         % So, rho*v = rho*(lambda1 + p'(rho)*rho)
-        lambda = [0 0.3 0.6];
+        lambda = [0 0.3 0.5];
         for h = 1:length(lambda)
             for k = 1:length(rho)
                 contourM(h,k) = rho(k)*(lambda(h) + pprime(rho(k))*rho(k));
@@ -66,24 +66,25 @@ if (model == 1)
         
         hold on
         plot(rho,contourM,'m')
-        text(0.26,0.85,'\lambda_1 = 0','FontSize',size_font)
+        contourM(1,:)
+        text(0.18,0.85,'\lambda_1 = 0.5','FontSize',size_font)
         text(1.8,0.85,'\lambda_1 = 0.3','FontSize',size_font)
-        text(3.6,0.85,'\lambda_1 = 0.6','FontSize',size_font)
+        text(3.6,0.85,'\lambda_1 = 0','FontSize',size_font)
         
         % Finding index for given point 
         switch_index = find(rho==rho_star(state));
         
         % Plotting valid states for ql
         Mfigs(2) = figure;
-        plot(rho(1:switch_index),rv1(state,1:switch_index),'r')
-        hold on
-        plot(rho(switch_index:end),rv1(state,switch_index:end),'g')
-        
-        % Plotting valid states for qr
-        Mfigs(3) = figure;
         plot(rho(1:switch_index),rv1(state,1:switch_index),'g')
         hold on
         plot(rho(switch_index:end),rv1(state,switch_index:end),'r')
+        
+        % Plotting valid states for qr
+        Mfigs(3) = figure;
+        plot(rho(1:switch_index),rv1(state,1:switch_index),'r')
+        hold on
+        plot(rho(switch_index:end),rv1(state,switch_index:end),'g')
         
         for k = 2:3
             figure(Mfigs(k))

### 2015-03-06 ClawpackCode/PW/pw.f90
Added updated PW code

@@ -1,0 +1,90 @@
+! =========================================================
+subroutine rp1(maxmx,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =========================================================
+
+! Solve Riemann problems for the 1D PW equations
+!   (rho)_t + (rho*v)_x = 0
+!   (rho*v)_t + (rho*v^2 + rho^gamma)_x = 0
+! using an approximate Riemann solver
+
+! waves: 2
+! equations: 2
+
+! Conserved quantities:
+!       1 density
+!       2 momentum
+
+! This function solves the Riemann problem at all interfaces in one call
+
+! On input, ql contains the state vector at the left edge of each cell
+!           qr contains the state vector at the right edge of each cell
+! On output, wave contains the waves,
+!            s the speeds,
+!            amdq the  left-going flux difference  A^- \Delta q
+!            apdq the right-going flux difference  A^+ \Delta q
+
+! Note that the i'th Riemann problem has left state qr(:,i-1)
+!                                    and right state ql(:,i)
+! From the basic clawpack routine step1, rp is called with ql = qr = q.
+
+    implicit double precision (a-h,o-z)
+
+    dimension   ql(meqn,1-mbc:maxmx+mbc)
+    dimension   qr(meqn,1-mbc:maxmx+mbc)
+    dimension    s(mwaves,1-mbc:maxmx+mbc)
+    dimension wave(meqn, mwaves,1-mbc:maxmx+mbc)
+    dimension amdq(meqn,1-mbc:maxmx+mbc)
+    dimension apdq(meqn,1-mbc:maxmx+mbc)
+
+!     # Local storage
+!     ---------------
+    dimension delta(2)
+    gamma = 0.2d0
+
+!     # Main loop of the Riemann solver.
+    do 30 i=2-mbc,mx+mbc
+
+    !     # compute  averaged quantities:
+        rbar = 0.5d0*(qr(1,i-1) + ql(1,i))
+        cbar = dsqrt(gamma*rbar**(gamma-1.0d0))
+        ubar = 0.5d0*(qr(2,i-1)/qr(1,i-1) + ql(2,i)/ql(1,i))
+
+    !     # delta(1)=rho(i)-rho(i-1) and  delta(2)=rho*v(i)-rho*v(i-1)
+        delta(1) = ql(1,i) - qr(1,i-1)
[…]

### 2015-03-06 ClawpackCode/PW/qinit.f90

@@ -1,0 +1,32 @@
+subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
+
+    ! Set initial conditions for the q array.
+    ! This default version prints an error message since it should
+    ! not be used directly.  Copy this to an application directory and
+    ! loop over all grid cells to set values of q(1:meqn, 1:mx).
+
+    implicit none
+    
+    integer, intent(in) :: meqn,mbc,mx,maux
+    real(kind=8), intent(in) :: xlower,dx
+    real(kind=8), intent(in) :: aux(maux,1-mbc:mx+mbc)
+    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc)
+
+    integer :: i
+    real(kind=8) :: xcell
+ 
+      do i=1,mx
+         xcell = xlower + (i-0.5d0)*dx
+        if (xcell.lt.0.0d0) then
+            q(1,i) = 0.3d0
+            q(2,i) = 0.06d0
+            ! q(1,i) = 0.001d0
+            ! q(2,i) = 0.0d0
+        else
+            q(1,i) = 0.4d0
+            q(2,i) = 0.12d0
+        endif
+      enddo
+
+end subroutine qinit
+

### 2015-03-06 ClawpackCode/PW/setplot.py

@@ -1,0 +1,79 @@
+
+""" 
+Set up the plot figures, axes, and items to be done for each frame.
+
+This module is imported by the plotting routines and then the
+function setplot is called to set the plot parameters.
+    
+""" 
+
+#--------------------------
+def setplot(plotdata):
+#--------------------------
+    
+    """ 
+    Specify what is to be plotted at each frame.
+    Input:  plotdata, an instance of clawpack.visclaw.data.ClawPlotData.
+    Output: a modified version of plotdata.
+    
+    """ 
+
+
+    plotdata.clearfigures()  # clear any old figures,axes,items data
+
+    # Density
+    plotfigure = plotdata.new_plotfigure(name='Density and Velocity', figno=1)
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [0.25,0.45]
+    # plotaxes.ylimits = [-0.1,0.5]
+    plotaxes.title = 'Density'
+
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 0
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+    
+    # Velocity
+    def velocity(current_data):
+        q = current_data.q  # solution when this function is called
+        p = q[0,:]          # density
+        m = q[1,:]          # momentum
+        v = m / p             # velocity
+        return v
+    
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [0.1,0.35]
+    # plotaxes.ylimits = [-9,2]
+    plotaxes.title = 'Velocity'
+    
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = velocity   # defined above
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+
+    # Parameters used only when creating html and/or latex hardcopy
+    # e.g., via clawpack.visclaw.frametools.printframes:
+
+    plotdata.printfigs = True                # print figures
+    plotdata.print_format = 'png'            # file format
[…]

### 2015-03-06 ClawpackCode/PW/setrun.py

@@ -1,0 +1,220 @@
+""" 
+Module to set up run time parameters for Clawpack -- classic code.
+
+The values set in the function setrun are then written out to data files
+that will be read in by the Fortran code.
+    
+""" 
+
+import os
+import numpy as np
+
+#------------------------------
+def setrun(claw_pkg='classic'):
+#------------------------------
+    
+    """ 
+    Define the parameters used for running Clawpack.
+
+    INPUT:
+        claw_pkg expected to be "classic" for this setrun.
+
+    OUTPUT:
+        rundata - object of class ClawRunData 
+    
+    """ 
+    
+    from clawpack.clawutil import data 
+    
+    
+    assert claw_pkg.lower() == 'classic',  "Expected claw_pkg = 'classic'"
+
+    num_dim = 1
+    rundata = data.ClawRunData(claw_pkg, num_dim)
+
+
+    #------------------------------------------------------------------
+    # Standard Clawpack parameters to be written to claw.data:
+    #------------------------------------------------------------------
+
+    clawdata = rundata.clawdata  # initialized when rundata instantiated
+
+
+    # ---------------
+    # Spatial domain:
+    # ---------------
+
+    # Number of space dimensions:
+    clawdata.num_dim = num_dim
+    
+    # Lower and upper edge of computational domain:
+    clawdata.lower[0] = -1.000000e+00
+    clawdata.upper[0] = 1.000000e+00
+    # clawdata.lower[0] = -0.500000e+00         # xlower
+    # clawdata.upper[0] = 0.500000e+00          # xupper
+    
+    # Number of grid cells:
+    clawdata.num_cells[0] = 400      # mx
+    
+
+    # ---------------
+    # Size of system:
+    # ---------------
+
+    # Number of equations in the system:
+    clawdata.num_eqn = 2
+
+    # Number of auxiliary variables in the aux array (initialized in setaux)
+    clawdata.num_aux = 0
+    
+    # Index of aux array corresponding to capacity function, if there is one:
+    clawdata.capa_index = 0
+    
+    
+    # -------------
+    # Initial time:
+    # -------------
+
+    clawdata.t0 = 0.000000
+    
+
+    # Restart from checkpoint file of a previous run?
+    # Note: If restarting, you must also change the Makefile to set:
+    #    RESTART = True
+    # If restarting, t0 above should be from original run, and the
+    # restart_file 'fort.qNNNN' specified below should be in 
[…]

### 2015-03-06 ClawpackCode/pw.f90
Removed old files

@@ -1,197 +1,0 @@
-! =========================================================
-subroutine rp1(maxmx,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-! =========================================================
-
-! Solve Riemann problems for the 1D PW equations
-!   (rho)_t + (rho*v)_x = 0
-!   (rho*v)_t + (rho*v^2 + rho^gamma)_x = 0
-! using Roe's approximate Riemann solver with entropy fix for
-! transonic rarefactions.
-
-! waves: 2
-! equations: 2
-
-! Conserved quantities:
-!       1 density
-!       2 momentum
-
-! This function solves the Riemann problem at all interfaces in one call
-
-! On input, ql contains the state vector at the left edge of each cell
-!           qr contains the state vector at the right edge of each cell
-! On output, wave contains the waves,
-!            s the speeds,
-!            amdq the  left-going flux difference  A^- \Delta q
-!            apdq the right-going flux difference  A^+ \Delta q
-
-! Note that the i'th Riemann problem has left state qr(:,i-1)
-!                                    and right state ql(:,i)
-! From the basic clawpack routine step1, rp is called with ql = qr = q.
-
-
-    implicit double precision (a-h,o-z)
-
-    dimension   ql(meqn,1-mbc:maxmx+mbc)
-    dimension   qr(meqn,1-mbc:maxmx+mbc)
-    dimension    s(mwaves,1-mbc:maxmx+mbc)
-    dimension wave(meqn, mwaves,1-mbc:maxmx+mbc)
-    dimension amdq(meqn,1-mbc:maxmx+mbc)
-    dimension apdq(meqn,1-mbc:maxmx+mbc)
-
-!     # Local storage
-!     ---------------
-    dimension delta(2)
-    logical :: efix
-
-    data efix /.true./    !# use entropy fix for transonic rarefactions
-
-    gamma = 0.2d0
-
-!     # Main loop of the Riemann solver.
-    do 30 i=2-mbc,mx+mbc
-
-
-    !     # compute  Roe-averaged quantities:
-        rbar = 0.5d0*(qr(1,i-1) + ql(1,i))
-        cbar = dsqrt(gamma*rbar**(gamma-1.0d0))
-        ubar = (qr(2,i-1)/dsqrt(qr(1,i-1)) + ql(2,i)/dsqrt(ql(1,i)))/ &
[…]

### 2015-03-06 ClawpackCode/qinit.f90

@@ -1,30 +1,0 @@
-subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
-
-    ! Set initial conditions for the q array.
-    ! This default version prints an error message since it should
-    ! not be used directly.  Copy this to an application directory and
-    ! loop over all grid cells to set values of q(1:meqn, 1:mx).
-
-    implicit none
-    
-    integer, intent(in) :: meqn,mbc,mx,maux
-    real(kind=8), intent(in) :: xlower,dx
-    real(kind=8), intent(in) :: aux(maux,1-mbc:mx+mbc)
-    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc)
-
-    integer :: i
-    real(kind=8) :: xcell
- 
-      do i=1,mx
-         xcell = xlower + (i-0.5d0)*dx
-        if (xcell.lt.0.0d0) then
-            q(1,i) = 0.01d0
-            q(2,i) = 0.01d0
-        else
-            q(1,i) = 0.4d0
-            q(2,i) = 0.12d0
-        endif
-      enddo
-
-end subroutine qinit
-

### 2015-03-06 ClawpackCode/setplot.py

@@ -1,72 +1,0 @@
-
-""" 
-Set up the plot figures, axes, and items to be done for each frame.
-
-This module is imported by the plotting routines and then the
-function setplot is called to set the plot parameters.
-    
-""" 
-
-#--------------------------
-def setplot(plotdata):
-#--------------------------
-    
-    """ 
-    Specify what is to be plotted at each frame.
-    Input:  plotdata, an instance of clawpack.visclaw.data.ClawPlotData.
-    Output: a modified version of plotdata.
-    
-    """ 
-
-
-    plotdata.clearfigures()  # clear any old figures,axes,items data
-
-    # Figure for q[0]
-    plotfigure = plotdata.new_plotfigure(name='density and momentum', figno=1)
-
-    # Set up for axes in this figure:
-    plotaxes = plotfigure.new_plotaxes()
-    plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
-    plotaxes.xlimits = 'auto'
-    plotaxes.ylimits = [-0.1,0.45]
-    plotaxes.title = 'density (rho)'
-
-    # Set up for item on these axes:
-    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
-    plotitem.plot_var = 0
-    plotitem.plotstyle = '-'
-    plotitem.color = 'b'
-
-
-    # Figure for q[1]
-
-    # Set up for axes in this figure:
-    plotaxes = plotfigure.new_plotaxes()
-    plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
-    plotaxes.xlimits = 'auto'
-    plotaxes.ylimits = [-0.15,0.15]
-    plotaxes.title = 'momentum (rho*v)'
-
-    # Set up for item on these axes:
-    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
-    plotitem.plot_var = 1
-    plotitem.plotstyle = '-'
-    plotitem.color = 'b'
-
-    # Parameters used only when creating html and/or latex hardcopy
-    # e.g., via clawpack.visclaw.frametools.printframes:
-
-    plotdata.printfigs = True                # print figures
-    plotdata.print_format = 'png'            # file format
-    plotdata.print_framenos = 'all'          # list of frames to print
-    plotdata.print_fignos = 'all'            # list of figures to print
-    plotdata.html = True                     # create html files of plots?
-    plotdata.html_homelink = '../README.html'
-    plotdata.latex = True                    # create latex file of plots?
[…]

### 2015-03-06 ClawpackCode/setrun.py

@@ -1,217 +1,0 @@
-""" 
-Module to set up run time parameters for Clawpack -- classic code.
-
-The values set in the function setrun are then written out to data files
-that will be read in by the Fortran code.
-    
-""" 
-
-import os
-import numpy as np
-
-#------------------------------
-def setrun(claw_pkg='classic'):
-#------------------------------
-    
-    """ 
-    Define the parameters used for running Clawpack.
-
-    INPUT:
-        claw_pkg expected to be "classic" for this setrun.
-
-    OUTPUT:
-        rundata - object of class ClawRunData 
-    
-    """ 
-    
-    from clawpack.clawutil import data 
-    
-    
-    assert claw_pkg.lower() == 'classic',  "Expected claw_pkg = 'classic'"
-
-    num_dim = 1
-    rundata = data.ClawRunData(claw_pkg, num_dim)
-
-
-    #------------------------------------------------------------------
-    # Standard Clawpack parameters to be written to claw.data:
-    #------------------------------------------------------------------
-
-    clawdata = rundata.clawdata  # initialized when rundata instantiated
-
-
-    # ---------------
-    # Spatial domain:
-    # ---------------
-
-    # Number of space dimensions:
-    clawdata.num_dim = num_dim
-    
-    # Lower and upper edge of computational domain:
-    clawdata.lower[0] = -1.000000e+00          # xlower
-    clawdata.upper[0] = 1.000000e+00          # xupper
-    
-    # Number of grid cells:
-    clawdata.num_cells[0] = 400      # mx
-    
-
-    # ---------------
-    # Size of system:
-    # ---------------
-
-    # Number of equations in the system:
-    clawdata.num_eqn = 2
-
-    # Number of auxiliary variables in the aux array (initialized in setaux)
-    clawdata.num_aux = 0
-    
-    # Index of aux array corresponding to capacity function, if there is one:
-    clawdata.capa_index = 0
-    
-    
-    # -------------
-    # Initial time:
-    # -------------
-
-    clawdata.t0 = 0.000000
-    
-
-    # Restart from checkpoint file of a previous run?
-    # Note: If restarting, you must also change the Makefile to set:
-    #    RESTART = True
-    # If restarting, t0 above should be from original run, and the
-    # restart_file 'fort.qNNNN' specified below should be in 
-    # the OUTDIR indicated in Makefile.
-
[…]

### 2015-03-06 ClawpackCode/AR/ar.f90
Added AR code files

@@ -1,0 +1,90 @@
+! =========================================================
+subroutine rp1(maxmx,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =========================================================
+
+! Solve Riemann problems for the 1D AR equations
+!   (rho)_t + (rho*v)_x = 0
+!   (v)_t + (v - rho * (rho^gamma)) * (v)_x = 0
+! using approximate Riemann solver
+
+! waves: 2
+! equations: 2
+
+! Conserved quantities:
+!       1 density
+!       2 "momentum" = rho * (v + rho*(rho^gamma))
+
+! This function solves the Riemann problem at all interfaces in one call
+
+! On input, ql contains the state vector at the left edge of each cell
+!           qr contains the state vector at the right edge of each cell
+! On output, wave contains the waves,
+!            s the speeds,
+!            amdq the  left-going flux difference  A^- \Delta q
+!            apdq the right-going flux difference  A^+ \Delta q
+
+! Note that the i'th Riemann problem has left state qr(:,i-1)
+!                                    and right state ql(:,i)
+! From the basic clawpack routine step1, rp is called with ql = qr = q.
+
+    implicit double precision (a-h,o-z)
+
+    dimension   ql(meqn,1-mbc:maxmx+mbc)
+    dimension   qr(meqn,1-mbc:maxmx+mbc)
+    dimension    s(mwaves,1-mbc:maxmx+mbc)
+    dimension wave(meqn, mwaves,1-mbc:maxmx+mbc)
+    dimension amdq(meqn,1-mbc:maxmx+mbc)
+    dimension apdq(meqn,1-mbc:maxmx+mbc)
+
+!     # Local storage
+!     ---------------
+    dimension delta(2)
+    gamma = 0.2d0
+
+!     # Main loop of the Riemann solver.
+    do 30 i=2-mbc,mx+mbc
+
+    !     # compute  averaged quantities:
+        rbar = 0.5d0*(qr(1,i-1) + ql(1,i))
+        pbar = gamma*rbar**(gamma-1.0d0)
+        ubar = 0.5d0*(qr(2,i-1) + ql(2,i))
+
+    !     # delta(1)=rho(i)-rho(i-1) and  delta(2)=rho*v(i)-rho*v(i-1)
+        delta(1) = ql(1,i) - qr(1,i-1)
[…]

### 2015-03-06 ClawpackCode/AR/qinit.f90

@@ -1,0 +1,32 @@
+subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
+
+    ! Set initial conditions for the q array.
+    ! This default version prints an error message since it should
+    ! not be used directly.  Copy this to an application directory and
+    ! loop over all grid cells to set values of q(1:meqn, 1:mx).
+
+    implicit none
+    
+    integer, intent(in) :: meqn,mbc,mx,maux
+    real(kind=8), intent(in) :: xlower,dx
+    real(kind=8), intent(in) :: aux(maux,1-mbc:mx+mbc)
+    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc)
+
+    integer :: i
+    real(kind=8) :: xcell
+ 
+      do i=1,mx
+         xcell = xlower + (i-0.5d0)*dx
+        if (xcell.lt.0.0d0) then
+            q(1,i) = 0.3d0
+            q(2,i) = 0.2d0
+            ! q(1,i) = 0.00001d0
+            ! q(2,i) = 0.0d0
+        else
+            q(1,i) = 0.4d0
+            q(2,i) = 0.3d0
+        endif
+      enddo
+
+end subroutine qinit
+

### 2015-03-06 ClawpackCode/AR/setplot.py

@@ -1,0 +1,73 @@
+
+""" 
+Set up the plot figures, axes, and items to be done for each frame.
+
+This module is imported by the plotting routines and then the
+function setplot is called to set the plot parameters.
+    
+""" 
+
+#--------------------------
+def setplot(plotdata):
+#--------------------------
+    
+    """ 
+    Specify what is to be plotted at each frame.
+    Input:  plotdata, an instance of clawpack.visclaw.data.ClawPlotData.
+    Output: a modified version of plotdata.
+    
+    """ 
+
+
+    plotdata.clearfigures()  # clear any old figures,axes,items data
+
+    # Density
+    plotfigure = plotdata.new_plotfigure(name='Density and Velocity', figno=1)
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [0.1,0.45]
+    # plotaxes.ylimits = [-0.1,0.5]
+    plotaxes.title = 'Density'
+
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 0
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+    
+    # Velocity
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [0.1,0.4]
+    # plotaxes.ylimits = [-0.1,0.4]
+    plotaxes.title = 'Velocity'
+    
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 1
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+
+    # Parameters used only when creating html and/or latex hardcopy
+    # e.g., via clawpack.visclaw.frametools.printframes:
+
+    plotdata.printfigs = True                # print figures
+    plotdata.print_format = 'png'            # file format
+    plotdata.print_framenos = 'all'          # list of frames to print
+    plotdata.print_fignos = 'all'            # list of figures to print
+    plotdata.html = True                     # create html files of plots?
+    plotdata.html_homelink = '../README.html'
[…]

### 2015-03-06 ClawpackCode/AR/setrun.py

@@ -1,0 +1,220 @@
+""" 
+Module to set up run time parameters for Clawpack -- classic code.
+
+The values set in the function setrun are then written out to data files
+that will be read in by the Fortran code.
+    
+""" 
+
+import os
+import numpy as np
+
+#------------------------------
+def setrun(claw_pkg='classic'):
+#------------------------------
+    
+    """ 
+    Define the parameters used for running Clawpack.
+
+    INPUT:
+        claw_pkg expected to be "classic" for this setrun.
+
+    OUTPUT:
+        rundata - object of class ClawRunData 
+    
+    """ 
+    
+    from clawpack.clawutil import data 
+    
+    
+    assert claw_pkg.lower() == 'classic',  "Expected claw_pkg = 'classic'"
+
+    num_dim = 1
+    rundata = data.ClawRunData(claw_pkg, num_dim)
+
+
+    #------------------------------------------------------------------
+    # Standard Clawpack parameters to be written to claw.data:
+    #------------------------------------------------------------------
+
+    clawdata = rundata.clawdata  # initialized when rundata instantiated
+
+
+    # ---------------
+    # Spatial domain:
+    # ---------------
+
+    # Number of space dimensions:
+    clawdata.num_dim = num_dim
+    
+    # Lower and upper edge of computational domain:
+    clawdata.lower[0] = -0.300000e+00
+    clawdata.upper[0] = 0.500000e+00
+    # clawdata.lower[0] = -0.500000e+00         # xlower
+    # clawdata.upper[0] = 0.500000e+00          # xupper
+    
+    # Number of grid cells:
+    clawdata.num_cells[0] = 400      # mx
+    
+
+    # ---------------
+    # Size of system:
+    # ---------------
+
+    # Number of equations in the system:
+    clawdata.num_eqn = 2
+
+    # Number of auxiliary variables in the aux array (initialized in setaux)
+    clawdata.num_aux = 0
+    
+    # Index of aux array corresponding to capacity function, if there is one:
+    clawdata.capa_index = 0
+    
+    
+    # -------------
+    # Initial time:
+    # -------------
+
+    clawdata.t0 = 0.000000
+    
+
+    # Restart from checkpoint file of a previous run?
+    # Note: If restarting, you must also change the Makefile to set:
+    #    RESTART = True
+    # If restarting, t0 above should be from original run, and the
+    # restart_file 'fort.qNNNN' specified below should be in 
[…]

### 2015-03-09 ClawpackCode/AR/Example1/qinit.f90
Organizing Clawpack code for the two examples

@@ -20,8 +20,6 @@ subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
         if (xcell.lt.0.0d0) then
             q(1,i) = 0.3d0
             q(2,i) = 0.2d0
-            ! q(1,i) = 0.00001d0
-            ! q(2,i) = 0.0d0
         else
             q(1,i) = 0.4d0
             q(2,i) = 0.3d0

### 2015-03-09 ClawpackCode/AR/Example1/setplot.py

@@ -29,7 +29,6 @@ def setplot(plotdata):
     plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
     plotaxes.xlimits = 'auto'
     plotaxes.ylimits = [0.1,0.45]
-    # plotaxes.ylimits = [-0.1,0.5]
     plotaxes.title = 'Density'
 
     # Set up for item on these axes:
@@ -45,7 +44,6 @@ def setplot(plotdata):
     plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
     plotaxes.xlimits = 'auto'
     plotaxes.ylimits = [0.1,0.4]
-    # plotaxes.ylimits = [-0.1,0.4]
     plotaxes.title = 'Velocity'
     
     # Set up for item on these axes:

### 2015-03-09 ClawpackCode/AR/Example1/setrun.py

@@ -1,0 +1,217 @@
+""" 
+Module to set up run time parameters for Clawpack -- classic code.
+
+The values set in the function setrun are then written out to data files
+that will be read in by the Fortran code.
+    
+""" 
+
+import os
+import numpy as np
+
+#------------------------------
+def setrun(claw_pkg='classic'):
+#------------------------------
+    
+    """ 
+    Define the parameters used for running Clawpack.
+
+    INPUT:
+        claw_pkg expected to be "classic" for this setrun.
+
+    OUTPUT:
+        rundata - object of class ClawRunData 
+    
+    """ 
+    
+    from clawpack.clawutil import data 
+    
+    
+    assert claw_pkg.lower() == 'classic',  "Expected claw_pkg = 'classic'"
+
+    num_dim = 1
+    rundata = data.ClawRunData(claw_pkg, num_dim)
+
+
+    #------------------------------------------------------------------
+    # Standard Clawpack parameters to be written to claw.data:
+    #------------------------------------------------------------------
+
+    clawdata = rundata.clawdata  # initialized when rundata instantiated
+
+
+    # ---------------
+    # Spatial domain:
+    # ---------------
+
+    # Number of space dimensions:
+    clawdata.num_dim = num_dim
+    
+    # Lower and upper edge of computational domain:
+    clawdata.lower[0] = -0.300000e+00         # xlower
+    clawdata.upper[0] = 0.500000e+00          # xupper
+    
+    # Number of grid cells:
+    clawdata.num_cells[0] = 400      # mx
+    
+
+    # ---------------
+    # Size of system:
+    # ---------------
+
+    # Number of equations in the system:
+    clawdata.num_eqn = 2
+
+    # Number of auxiliary variables in the aux array (initialized in setaux)
+    clawdata.num_aux = 0
+    
+    # Index of aux array corresponding to capacity function, if there is one:
+    clawdata.capa_index = 0
+    
+    
+    # -------------
+    # Initial time:
+    # -------------
+
+    clawdata.t0 = 0.000000
+    
+
+    # Restart from checkpoint file of a previous run?
+    # Note: If restarting, you must also change the Makefile to set:
+    #    RESTART = True
+    # If restarting, t0 above should be from original run, and the
+    # restart_file 'fort.qNNNN' specified below should be in 
+    # the OUTDIR indicated in Makefile.
+
[…]

### 2015-03-09 ClawpackCode/AR/Example2/qinit.f90

@@ -18,13 +18,11 @@ subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
       do i=1,mx
          xcell = xlower + (i-0.5d0)*dx
         if (xcell.lt.0.0d0) then
-            q(1,i) = 0.3d0
-            q(2,i) = 0.06d0
-            ! q(1,i) = 0.001d0
-            ! q(2,i) = 0.0d0
+            q(1,i) = 0.00001d0
+            q(2,i) = 0.0d0
         else
             q(1,i) = 0.4d0
-            q(2,i) = 0.12d0
+            q(2,i) = 0.3d0
         endif
       enddo
 

### 2015-03-09 ClawpackCode/AR/Example2/setplot.py

@@ -1,0 +1,71 @@
+
+""" 
+Set up the plot figures, axes, and items to be done for each frame.
+
+This module is imported by the plotting routines and then the
+function setplot is called to set the plot parameters.
+    
+""" 
+
+#--------------------------
+def setplot(plotdata):
+#--------------------------
+    
+    """ 
+    Specify what is to be plotted at each frame.
+    Input:  plotdata, an instance of clawpack.visclaw.data.ClawPlotData.
+    Output: a modified version of plotdata.
+    
+    """ 
+
+
+    plotdata.clearfigures()  # clear any old figures,axes,items data
+
+    # Density
+    plotfigure = plotdata.new_plotfigure(name='Density and Velocity', figno=1)
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [-0.1,0.5]
+    plotaxes.title = 'Density'
+
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 0
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+    
+    # Velocity
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [-0.1,0.4]
+    plotaxes.title = 'Velocity'
+    
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 1
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+
+    # Parameters used only when creating html and/or latex hardcopy
+    # e.g., via clawpack.visclaw.frametools.printframes:
+
+    plotdata.printfigs = True                # print figures
+    plotdata.print_format = 'png'            # file format
+    plotdata.print_framenos = 'all'          # list of frames to print
+    plotdata.print_fignos = 'all'            # list of figures to print
+    plotdata.html = True                     # create html files of plots?
+    plotdata.html_homelink = '../README.html'
+    plotdata.latex = True                    # create latex file of plots?
[…]

### 2015-03-09 ClawpackCode/AR/Example2/setrun.py

@@ -48,10 +48,8 @@ def setrun(claw_pkg='classic'):
     clawdata.num_dim = num_dim
     
     # Lower and upper edge of computational domain:
-    clawdata.lower[0] = -0.300000e+00
-    clawdata.upper[0] = 0.500000e+00
-    # clawdata.lower[0] = -0.500000e+00         # xlower
-    # clawdata.upper[0] = 0.500000e+00          # xupper
+    clawdata.lower[0] = -0.500000e+00         # xlower
+    clawdata.upper[0] = 0.500000e+00          # xupper
     
     # Number of grid cells:
     clawdata.num_cells[0] = 400      # mx
@@ -102,8 +100,7 @@ def setrun(claw_pkg='classic'):
         # Output ntimes frames at equally spaced times up to tfinal:
         # Can specify num_output_times = 0 for no output
         clawdata.num_output_times = 16
-        clawdata.tfinal = 1.000000
-        # clawdata.tfinal = 0.500000
+        clawdata.tfinal = 0.500000
         clawdata.output_t0 = True  # output at initial (or restart) time?
         
     elif clawdata.output_style == 2:

### 2015-03-09 ClawpackCode/PW/Example1/qinit.f90

@@ -1,0 +1,30 @@
+subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
+
+    ! Set initial conditions for the q array.
+    ! This default version prints an error message since it should
+    ! not be used directly.  Copy this to an application directory and
+    ! loop over all grid cells to set values of q(1:meqn, 1:mx).
+
+    implicit none
+    
+    integer, intent(in) :: meqn,mbc,mx,maux
+    real(kind=8), intent(in) :: xlower,dx
+    real(kind=8), intent(in) :: aux(maux,1-mbc:mx+mbc)
+    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc)
+
+    integer :: i
+    real(kind=8) :: xcell
+ 
+      do i=1,mx
+         xcell = xlower + (i-0.5d0)*dx
+        if (xcell.lt.0.0d0) then
+            q(1,i) = 0.3d0
+            q(2,i) = 0.06d0
+        else
+            q(1,i) = 0.4d0
+            q(2,i) = 0.12d0
+        endif
+      enddo
+
+end subroutine qinit
+

### 2015-03-09 ClawpackCode/PW/Example1/setplot.py

@@ -29,7 +29,6 @@ def setplot(plotdata):
     plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
     plotaxes.xlimits = 'auto'
     plotaxes.ylimits = [0.25,0.45]
-    # plotaxes.ylimits = [-0.1,0.5]
     plotaxes.title = 'Density'
 
     # Set up for item on these axes:
@@ -51,7 +50,6 @@ def setplot(plotdata):
     plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
     plotaxes.xlimits = 'auto'
     plotaxes.ylimits = [0.1,0.35]
-    # plotaxes.ylimits = [-9,2]
     plotaxes.title = 'Velocity'
     
     # Set up for item on these axes:

### 2015-03-09 ClawpackCode/PW/Example1/setrun.py

@@ -48,10 +48,8 @@ def setrun(claw_pkg='classic'):
     clawdata.num_dim = num_dim
     
     # Lower and upper edge of computational domain:
-    clawdata.lower[0] = -1.000000e+00
-    clawdata.upper[0] = 1.000000e+00
-    # clawdata.lower[0] = -0.500000e+00         # xlower
-    # clawdata.upper[0] = 0.500000e+00          # xupper
+    clawdata.lower[0] = -1.000000e+00         # xlower
+    clawdata.upper[0] = 1.000000e+00          # xupper
     
     # Number of grid cells:
     clawdata.num_cells[0] = 400      # mx
@@ -103,7 +101,6 @@ def setrun(claw_pkg='classic'):
         # Can specify num_output_times = 0 for no output
         clawdata.num_output_times = 16
         clawdata.tfinal = 1.000000
-        # clawdata.tfinal = 0.100000
         clawdata.output_t0 = True  # output at initial (or restart) time?
         
     elif clawdata.output_style == 2:

### 2015-03-09 ClawpackCode/PW/Example2/qinit.f90
Adding example that I missed

@@ -1,0 +1,30 @@
+subroutine qinit(meqn,mbc,mx,xlower,dx,q,maux,aux)
+
+    ! Set initial conditions for the q array.
+    ! This default version prints an error message since it should
+    ! not be used directly.  Copy this to an application directory and
+    ! loop over all grid cells to set values of q(1:meqn, 1:mx).
+
+    implicit none
+    
+    integer, intent(in) :: meqn,mbc,mx,maux
+    real(kind=8), intent(in) :: xlower,dx
+    real(kind=8), intent(in) :: aux(maux,1-mbc:mx+mbc)
+    real(kind=8), intent(inout) :: q(meqn,1-mbc:mx+mbc)
+
+    integer :: i
+    real(kind=8) :: xcell
+ 
+      do i=1,mx
+         xcell = xlower + (i-0.5d0)*dx
+        if (xcell.lt.0.0d0) then
+            q(1,i) = 0.001d0
+            q(2,i) = 0.0d0
+        else
+            q(1,i) = 0.4d0
+            q(2,i) = 0.12d0
+        endif
+      enddo
+
+end subroutine qinit
+

### 2015-03-09 ClawpackCode/PW/Example2/setplot.py

@@ -1,0 +1,77 @@
+
+""" 
+Set up the plot figures, axes, and items to be done for each frame.
+
+This module is imported by the plotting routines and then the
+function setplot is called to set the plot parameters.
+    
+""" 
+
+#--------------------------
+def setplot(plotdata):
+#--------------------------
+    
+    """ 
+    Specify what is to be plotted at each frame.
+    Input:  plotdata, an instance of clawpack.visclaw.data.ClawPlotData.
+    Output: a modified version of plotdata.
+    
+    """ 
+
+
+    plotdata.clearfigures()  # clear any old figures,axes,items data
+
+    # Density
+    plotfigure = plotdata.new_plotfigure(name='Density and Velocity', figno=1)
+
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,1)'   # top figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [-0.1,0.5]
+    plotaxes.title = 'Density'
+
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = 0
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+    
+    # Velocity
+    def velocity(current_data):
+        q = current_data.q  # solution when this function is called
+        p = q[0,:]          # density
+        m = q[1,:]          # momentum
+        v = m / p             # velocity
+        return v
+    
+    # Set up for axes in this figure:
+    plotaxes = plotfigure.new_plotaxes()
+    plotaxes.axescmd = 'subplot(2,1,2)'   # bottom figure
+    plotaxes.xlimits = 'auto'
+    plotaxes.ylimits = [-9,2]
+    plotaxes.title = 'Velocity'
+    
+    # Set up for item on these axes:
+    plotitem = plotaxes.new_plotitem(plot_type='1d_plot')
+    plotitem.plot_var = velocity   # defined above
+    plotitem.plotstyle = '-'
+    plotitem.color = 'b'
+
+    # Parameters used only when creating html and/or latex hardcopy
+    # e.g., via clawpack.visclaw.frametools.printframes:
+
+    plotdata.printfigs = True                # print figures
+    plotdata.print_format = 'png'            # file format
+    plotdata.print_framenos = 'all'          # list of frames to print
+    plotdata.print_fignos = 'all'            # list of figures to print
[…]

### 2015-03-09 ClawpackCode/PW/Example2/setrun.py

@@ -1,0 +1,217 @@
+""" 
+Module to set up run time parameters for Clawpack -- classic code.
+
+The values set in the function setrun are then written out to data files
+that will be read in by the Fortran code.
+    
+""" 
+
+import os
+import numpy as np
+
+#------------------------------
+def setrun(claw_pkg='classic'):
+#------------------------------
+    
+    """ 
+    Define the parameters used for running Clawpack.
+
+    INPUT:
+        claw_pkg expected to be "classic" for this setrun.
+
+    OUTPUT:
+        rundata - object of class ClawRunData 
+    
+    """ 
+    
+    from clawpack.clawutil import data 
+    
+    
+    assert claw_pkg.lower() == 'classic',  "Expected claw_pkg = 'classic'"
+
+    num_dim = 1
+    rundata = data.ClawRunData(claw_pkg, num_dim)
+
+
+    #------------------------------------------------------------------
+    # Standard Clawpack parameters to be written to claw.data:
+    #------------------------------------------------------------------
+
+    clawdata = rundata.clawdata  # initialized when rundata instantiated
+
+
+    # ---------------
+    # Spatial domain:
+    # ---------------
+
+    # Number of space dimensions:
+    clawdata.num_dim = num_dim
+    
+    # Lower and upper edge of computational domain:
+    clawdata.lower[0] = -0.500000e+00         # xlower
+    clawdata.upper[0] = 0.500000e+00          # xupper
+    
+    # Number of grid cells:
+    clawdata.num_cells[0] = 400      # mx
+    
+
+    # ---------------
+    # Size of system:
+    # ---------------
+
+    # Number of equations in the system:
+    clawdata.num_eqn = 2
+
+    # Number of auxiliary variables in the aux array (initialized in setaux)
+    clawdata.num_aux = 0
+    
+    # Index of aux array corresponding to capacity function, if there is one:
+    clawdata.capa_index = 0
+    
+    
+    # -------------
+    # Initial time:
+    # -------------
+
+    clawdata.t0 = 0.000000
+    
+
+    # Restart from checkpoint file of a previous run?
+    # Note: If restarting, you must also change the Makefile to set:
+    #    RESTART = True
+    # If restarting, t0 above should be from original run, and the
+    # restart_file 'fort.qNNNN' specified below should be in 
+    # the OUTDIR indicated in Makefile.
+
[…]

### 2015-03-09 Presentation/presentation.tex
Adding outline of presentation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Short title]{Full Title of the Talk} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Kelsey Maass and Brisa Davis} % Your name
\institute[UW] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
AMATH 574 \\ % Your institution for the title page \\
Group 6
\medskip
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\frametitle{Introduction} 
Kelsey
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

\begin{frame}
\frametitle{LWR}
Kelsey
\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{PW}
Intro, Kelsey. Eigenvectors and values
\end{frame}

%------------------------------------------------

%------------------------------------------------

\begin{frame}
\frametitle{PW}
Brisa, loci and integral curves. Areas of validity.
\end{frame}

%------------------------------------------------

%------------------------------------------------

\begin{frame}
\frametitle{AR}
Intro, Kelsey. Eigenvectors and values
\end{frame}

%------------------------------------------------

%------------------------------------------------

\begin{frame}
\frametitle{AR}
Brisa, loci and integral curves. Areas of validity.
\end{frame}

%------------------------------------------------

%------------------------------------------------

\begin{frame}
\frametitle{Examples}
Brisa
\end{frame}

%------------------------------------------------

\begin{frame}
\Huge{\centerline{Questions?}}
\end{frame}

%----------------------------------------------------------------------------------------

\end{document}

### 2015-03-10 MatlabCode/AR_LociAndCurves.m
Fixing plots, adding PW Matlab code.

@@ -16,7 +16,7 @@ validity = false;
 %% Setting up curves in the U = (rho, v) plane
 
 % Calling function to generate figures
-Ufigs = UPlot(1, validity);
+Ufigs = UPlot(validity);
 
 % Exporting figures
 figure(Ufigs(1))
@@ -82,7 +82,7 @@ cd ..
 %% Setting up curves in the Y = (rho,y) plane
 
 % Calling function to generate figures
-Yfigs = YPlot(1);
+Yfigs = YPlot();
 
 figure(Yfigs(1))
 cd export_fig_code

### 2015-03-10 MatlabCode/MPlot.m

@@ -10,7 +10,7 @@
 
 function Mfigs = MPlot(model,validity)
 
-global v_star rho_star p pprime rho size_font
+global v_star rho_star p pprime rho size_font gamma
 
 if (model == 1)
     
@@ -60,18 +60,18 @@ if (model == 1)
         lambda = [0 0.3 0.5];
         for h = 1:length(lambda)
             for k = 1:length(rho)
-                contourM(h,k) = rho(k)*(lambda(h) + pprime(rho(k))*rho(k));
+                contourM(h,k) = rho(k)*(lambda(h) ...
+                    + pprime(rho(k))*rho(k));
             end
         end
         
         hold on
-        plot(rho,contourM,'m')
-        contourM(1,:)
+        plot(rho,contourM,'c')
         text(0.18,0.85,'\lambda_1 = 0.5','FontSize',size_font)
         text(1.8,0.85,'\lambda_1 = 0.3','FontSize',size_font)
         text(3.6,0.85,'\lambda_1 = 0','FontSize',size_font)
         
-        % Finding index for given point 
+        % Finding index for given point
         switch_index = find(rho==rho_star(state));
         
         % Plotting valid states for ql
@@ -91,13 +91,15 @@ if (model == 1)
             plot(rho,rv2(state,:),'b')
             plot(rho_star(state),...
                 rho_star(state)*v_star(state),'ro', 'LineWidth', 2)
-            x = rho_star(state) + 0.02;
-            y = rho_star(state)*v_star(state)-0.005;
+            x = rho_star(state) + 0.01;
+            y = rho_star(state)*v_star(state)-0.004;
             
             if (k == 2)
-                text(x,y,'$m_l$','Interpreter','latex','FontSize',size_font)
+                text(x,y,'$q_l$','Interpreter',...
+                    'latex','FontSize',size_font)
             else
-                text(x,y,'$m_r$','Interpreter','latex','FontSize',size_font)
+                text(x,y,'$q_r$','Interpreter',...
+                    'latex','FontSize',size_font)
             end
             
             axis([0 .7 0 .1])
@@ -108,8 +110,214 @@ if (model == 1)
         end
     end
     
[…]

### 2015-03-10 MatlabCode/PW_LociAndCurves.m

@@ -1,6 +1,7 @@
 % This function generates plots for the Hugoniot loci and Integral Curves
-% for the Aw Rascle model, and overlays contour lines for lambda1 and 
-% lambda2.
+% for the PW model.
+% Note that the loci and the integral curves coincide for this model,
+% so only two equations are needed.
 
 clear all
 close all
@@ -12,49 +13,16 @@ run variableSetup
 size_font = 25;
 validity = true;
 
-%% Setting up curves in the U = (rho, v) plane
+%% Setting up curves in the M = (rho, rho*v) plane
 
 % Calling function to generate figures
-Ufigs = UPlot(validity);
+Mfigs = MPlot(2, true);
 
 % Exporting figures
-figure(Ufigs(1))
-cd export_fig_code
-export_fig( gcf, ...      % figure handle
-    '../Images/Validity_U',... % name of output file without extension
-    '-painters', ...      % renderer
-    '-jpg', ...           % file format
-    '-r72' );             % resolution in dpi
-cd ..
-
-figure(Ufigs(2))
-cd export_fig_code
-export_fig( gcf, ...      % figure handle
-    '../Images/Validity_U_ql',... % name of output file without extension
-    '-painters', ...      % renderer
-    '-jpg', ...           % file format
-    '-r72' );             % resolution in dpi
-cd ..
-
-figure(Ufigs(3))
-cd export_fig_code
-export_fig( gcf, ...      % figure handle
-    '../Images/Validity_U_qr',... % name of output file without extension
-    '-painters', ...      % renderer
-    '-jpg', ...           % file format
-    '-r72' );             % resolution in dpi
-cd ..
-
-%% Setting up curves in the M = (rho,rho*v) plane
-
-% Calling function to generate figures
-Mfigs = MPlot(1,validity);
-
-% Exporting figues
 figure(Mfigs(1))
 cd export_fig_code
 export_fig( gcf, ...      % figure handle
-    '../Images/Validity_M',... % name of output file without extension
+    '../Images/PW_Loci',... % name of output file without extension
     '-painters', ...      % renderer
     '-jpg', ...           % file format
     '-r72' );             % resolution in dpi
@@ -63,7 +31,7 @@ cd ..
 figure(Mfigs(2))
 cd export_fig_code
 export_fig( gcf, ...      % figure handle
-    '../Images/Validity_M_ql',... % name of output file without extension
+    '../Images/PW_IntegralCurves',... % name of output file without extension
[…]

### 2015-03-10 MatlabCode/PW_examples.m

@@ -1,0 +1,156 @@
+% Function to generate examples at the end of the Aw Rascle paper
+% for the PW model.
+
+clear all
+close all
+clc
+
+% Declaring various starting values
+gamma = 0.2;
+p = @(x) x^gamma;
+pprime = @(x) gamma*x^(gamma-1);
+rho = 0:0.001:5;
+size_font = 20;
+
+% Loop plots both examples
+for example = 1:2
+    figure
+    % Setting up data for the two points ql and qr
+    if (example == 1)
+        v_star = [0.2 .3];
+        rho_star = [0.3 .4];
+    else
+        v_star = [0 .3];
+        rho_star = [0 .4];
+    end
+    
+    % Setting up equations
+    for h = 1:length(v_star)
+        for k = 1:length(rho)
+            num = (rho(k) - rho_star(h))*(p(rho(k)) - p(rho_star(h)));
+            denom = (rho(k)*rho_star(h));
+            var = rho(k)*sqrt(num/denom);
+            
+            rv1_loci(h,k) = rho(k)*v_star(h) + var;
+            rv2_loci(h,k) = rho(k)*v_star(h) - var;
+            
+            rv1_int(h,k) = rho(k)*(v_star(h) ...
+                + (2/(gamma-1))*(sqrt(pprime(rho_star(h)))...
+                - sqrt(pprime(rho(k)))));
+            rv2_int(h,k) = rho(k)*(v_star(h) ...
+                + (2/(gamma-1))*(-sqrt(pprime(rho_star(h)))...
+                + sqrt(pprime(rho(k)))));
+        end
+    end
+    
+    % Loop plots each of the two points ql and qr
+    for k = 1:length(rho_star)
+        state = k;
+        
+        % Finding index for given point
+        switch_index = find(rho==rho_star(state));
+        
+        if (k == 1)
+            % lamdba1 curves
+            plot(rho(switch_index:end),...
+                rv2_loci(state,switch_index:end),'m')
+            hold on
+            plot(rho(1:switch_index),...
+                rv1_int(state,1:switch_index),'Color',[0,1,0])
+            
+            % lambda2 curves
+            plot(rho(1:switch_index),...
+                rv2_loci(state,1:switch_index),'r')
+            plot(rho(switch_index:end),...
+                rv2_int(state,switch_index:end),'Color',[0,0.3,0])
+            
+            if (example == 1)
+                x = rho_star(state) + 0.001;
+                y = rho_star(state)*v_star(state)-0.007;
+            else
[…]

### 2015-03-10 MatlabCode/UPlot.m

@@ -1,6 +1,5 @@
-% Function to generate plots in the U plane for the AR and PW model. 
-% Inputs: model - 1 means AR model, 2 means PW model.
-%         validity - boolean indicating whether or not validity is being
+% Function to generate plots in the U plane for the AR model.
+% Input: validity - boolean indicating whether or not validity is being
 %                    considered
 % Generates 3 plots: curves for lambda1, curves for lambda2, and both
 % curves for a given point if validity = false. curves for lambda1 with
@@ -8,103 +7,95 @@
 % if validity = true.
 % Note: expects that the function "variableSetup" has been run.
 
-function Ufigs = UPlot(model,validity)
+function Ufigs = UPlot(validity)
 
-global v_star rho_star p pprime rho size_font 
+global v_star rho_star p pprime rho size_font
 
-if (model == 1)
-    
-    % Setting up equations
-    for h = 1:length(v_star)
-        for k = 1:length(rho)
-            v1(h,k) = v_star(h) + p(rho_star(h)) - p(rho(k));
-            v2(h,k) = v_star(h);
-        end
+% Setting up equations
+for h = 1:length(v_star)
+    for k = 1:length(rho)
+        v1(h,k) = v_star(h) + p(rho_star(h)) - p(rho(k));
+        v2(h,k) = v_star(h);
     end
-    
-    % Plotting 1-curves
-    Ufigs(1) = figure;
-    plot(rho,v1,'k')
+end
+
+% Plotting 1-curves
+Ufigs(1) = figure;
+plot(rho,v1,'k')
+
+axis([0 1 0 1])
+set(gca,'XTick',[0 1])
+set(gca,'YTick',[0 1])
+state = 1;
+
+if (~validity)
+    % Plotting 2-curves
+    Ufigs(2) = figure;
+    plot(rho,v2,'b')
     
     axis([0 1 0 1])
     set(gca,'XTick',[0 1])
     set(gca,'YTick',[0 1])
-    state = 1;
-
-    if (~validity)
-        % Plotting 2-curves
-        Ufigs(2) = figure;
-        plot(rho,v2,'b')
-        
-        axis([0 1 0 1])
-        set(gca,'XTick',[0 1])
-        set(gca,'YTick',[0 1])
-        
-        % Plotting both curves for a given initial state
-        Ufigs(3) = figure;
-        plot(rho,v1(state,:),'k')
-        hold on
-        plot(rho,v2(state,:),'b')
[…]

### 2015-03-10 Presentation/presentation.tex
Adding updates to presentation.

@@
 \documentclass{beamer}
+\usepackage{subfigure}
+\usepackage{amsmath}
+\usepackage{mathtools}
+\usepackage{listings}
+\usepackage{multicol}
+\usepackage{media9}
+
+\makeatletter \@addtoreset{subfigure}{framenumber} \makeatother
 
@@
 \begin{frame}
-\frametitle{PW}
-Brisa, loci and integral curves. Areas of validity.
+\frametitle{PW: Hugoniot Loci}
+From this system, the Rankine-Hugoniot condition 
+\begin{align*}
+s(q_* - q ) = f(q_*)-f(q)
+\end{align*}
+gives us the two equations 
+\begin{align*}
+&s(\rho_* - \rho ) = \rho_*v_* - \rho v, \\
+&s (\rho_*v_* - \rho v ) = \rho_*v_*^2 + p(\rho) - \rho v^2 - p(\rho).
+\end{align*}
+From these two equations
+the equation for the Hugoniot loci can be found to be
+\begin{align*}
+\rho v = \rho v_* \pm \rho \sqrt{\left( \rho - \rho_*\right) \left( p(\rho ) - p(\rho_*)\right) / \rho_*\rho }.
+\end{align*}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{PW: Hugoniot Loci}
+Plotting this equation 
+\begin{align*}
+\rho v = \rho v_* \pm \rho \sqrt{\left( \rho - \rho_*\right) \left( p(\rho ) - p(\rho_*)\right) / \rho_*\rho }.
+\end{align*}
+gives
+\begin{figure}[h!]
+ \centering
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_Loci.jpg}
+\end{figure}
+Here, $\lambda_1$ loci are shown in magenta and $\lambda_2$ loci are shown in red.
+\end{frame}
+
+\begin{frame}
+\frametitle{PW: Integral Curves}
+Using $p(\rho) = \rho^{\gamma}$ for $\gamma \neq 1$, since this is the value used in the AR model we will examine later, the equations for the integral curves can be found to be
+\begin{align*}
+\rho v = \rho v_* + \frac{2 \rho}{\gamma - 1}\left( \sqrt{p'(\rho_*)} - \sqrt{p'(\rho)}\right)
+\end{align*}
+for $\lambda_1$, and 
+\begin{align*}
+\rho v = \rho v_* + \frac{2 \rho}{\gamma - 1}\left( \sqrt{p'(\rho)} - \sqrt{p'(\rho_*)} \right)
+\end{align*}
+for $\lambda_2$.
+\end{frame}
+
+\begin{frame}
+\frametitle{PW: Integral Curves}
+Plotting these curves gives
+\begin{figure}[h!]
+ \centering
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_IntegralCurves.jpg}
+\end{figure}
+Here, $\lambda_1$ integral curves are shown in light green and $\lambda_2$ integral curves are shown in dark green.
+\end{frame}
+
+\begin{frame}
+\frametitle{PW: Areas of Validity}
+Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point. \\
+\vspace{0.1in}
+Shocks:
+\begin{align*}
+q_l \rightarrow q_r \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must decrease}\\
+q_r \rightarrow q_l \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must increase}
+\end{align*}
+Rarefaction:
+\begin{align*}
+q_l \rightarrow q_r \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must increase}\\
+q_r \rightarrow q_l \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must decrease}
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{PW: Areas of Validity}
+Using this, we can determine the valid curves:
+\begin{figure}[h!]
+ \centering
+ \subfigure[Valid curves for a point $q_l$.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_validStateCurvesLeft.jpg}
+   }
+ \subfigure[Valid curves for a point $q_r$.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_validStateCurvesRight.jpg}
+   }
+  \label{fig:PW_curves}
+\end{figure}
+
+Here, shocks are shown in red or magenta and rarefaction waves are shown in either light or dark green.
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{AR}
-Brisa, loci and integral curves. Areas of validity.
+\frametitle{AR: Hugoniot Loci}
+In order to consider the Hugoniot loci for this system we must first write the system in conservation form. 
+With some manipulation, we can get the system
+\begin{align*}
+&\partial_t\rho + \partial_x(\rho v) = 0, \\
+&\partial_t \left(\rho\left(v + p(\rho )\right)\right) + \partial_x \left( \rho v\left(v + p(\rho )\right)\right) = 0.
+\end{align*}
+This can be rewritten as $q_t + f(q)_x = $ by defining
+\begin{align*}
+q = \left[ \begin{matrix}
+\rho \\ y
+\end{matrix}\right], \hspace{0.3in}
+f(q) = \left[ \begin{matrix}
+v\rho \\
+vy
+\end{matrix}\right].
+\end{align*}
+where $y = \rho\left(v + p(\rho )\right)$.
+\end{frame}
+
+\begin{frame}
+\frametitle{AR: Hugoniot Loci}
+From the Rankine-Hugoniot condition
+\begin{align*}
+s(q_*- q) = f(q_*) - f(q).
+\end{align*}
+we get the two equations
+\begin{align*}
+s(\rho_* - \rho) = v_*\rho_* - v\rho,\\
+s(y_* - y) = v_*y_* - vy.
+\end{align*}
+From these equations, the equations for the Hugoniot loci can be found to be
+\begin{align*}
+&\rho v = \rho \left( v_* + p(\rho_*)\right) - \rho p(\rho)),\\
+&\rho v = \rho_* v_*,
+\end{align*}
+corresponding to $\lambda_1$ and $\lambda_2$, respectively.
+\end{frame}
+
+\begin{frame}
+\frametitle{AR: Integral Curves}
+In order to consider the integral curves for this system \cite{AwRascle2000} 
+works from the system
+\begin{align*}
+&\partial_t \rho + \partial_x (\rho v) = 0, \\ 
+&\partial_t v + \left(v - p'(\rho)\rho\right)\partial_x v = 0.
+\end{align*}
+After working with these equations to find the integral curves, you find that they coincide with the Hugoniot loci.
+\end{frame}
+
+\begin{frame}
+\frametitle{AR: Curves}
+Plotting these curves gives:
+\begin{figure}[h!]
+    \subfigure[$\lambda_1$-curves in the M plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb1.jpg}
+   }
+ \subfigure[$\lambda_2$-curves in the M plane.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb2.jpg}
+   }
+     \label{fig:AR_curves}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{AR: Areas of Validity}
+Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point. \\
+\vspace{0.1in}
+Shocks:
+\begin{align*}
+q_l \rightarrow q_r \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must decrease}\\
+q_r \rightarrow q_l \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must increase}
+\end{align*}
+Rarefaction:
+\begin{align*}
+q_l \rightarrow q_r \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must increase}\\
+q_r \rightarrow q_l \hspace{0.2in}&: \hspace{0.2in} \lambda \textnormal{ must decrease}
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{AR: Areas of Validity}
+Using this, we can determine the valid curves:
+\begin{figure}[h!]
+ \subfigure[Valid waves for the point $q_l$.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_ql.jpg}
+   }
+ \subfigure[Valid waves for the point $q_r$.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_qr.jpg}
+   }
+\end{figure}
+Here, shocks are shown in red, rarefaction waves are shown in green, and the contact discontinuity is shown in blue.
 \end{frame}
@@
 \frametitle{Examples}
-Brisa
+
+To illustrate a couple of examples where the AR model preforms better than the PW model, consider two different examples: 
+\begin{table}[t]
+\caption{Initial values used in examples 1 and 2.}
+\begin{center}
+\begin{tabular}{| c | c c  c c|}
+\hline
+& $v_l$ & $\rho_l $ & $v_r$ & $\rho_r $\\
+\hline
+Example 1 & 0.2 & .03 & 0.3 & 0.4 \\
+Example 2 & 0 & 0 & 0.3 & 0.4\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+\end{frame}
+
+\begin{frame}
+\frametitle{Example 1}
+The solution to the Riemann problem is:
+\begin{figure}[h!]
+ \centering
+ \subfigure[Solution from the AR model.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/AR_example1.jpg}
+   }
+    \subfigure[Solution from the PW model.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_example1.jpg}
+   }
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Example 1}
+\begin{center}
+\begin{multicols}{2}%
+AR model:
+\includemedia[
+  width=0.9\linewidth,
+  height=0.7\linewidth,
+  activate=pageopen,
+  addresource=AR_Example1Movie.mp4,
+  flashvars={source=AR_Example1Movie.mp4}
+]{}{VPlayer.swf}
+\vfill
+\columnbreak
+PW Model:
+\includemedia[
+  width=0.9\linewidth,
+  height=0.7\linewidth,
+  activate=pageopen,
+  addresource=PW_Example1Movie.mp4,
+  flashvars={source=PW_Example1Movie.mp4}
+]{}{VPlayer.swf}
+\end{multicols}
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Example 2}
+The solution to the Riemann problem is:
+
+\begin{figure}[h!]
+ \centering
+ \subfigure[Solution from the AR model.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/AR_example2.jpg}
+   }
+ \subfigure[Solution from the PW model.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_example2.jpg}
+   }
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{Example 2}
+\begin{center}
+\begin{multicols}{2}%
+AR model:
+\includemedia[
+  width=0.9\linewidth,
+  height=0.7\linewidth,
+  activate=pageopen,
+  addresource=AR_Example2Movie.mp4,
+  flashvars={source=AR_Example2Movie.mp4}
+]{}{VPlayer.swf}
+\vfill
+\columnbreak
+PW Model:
+\includemedia[
+  width=0.9\linewidth,
+  height=0.7\linewidth,
+  activate=pageopen,
+  addresource=PW_Example2Movie.mp4,
+  flashvars={source=PW_Example2Movie.mp4}
+]{}{VPlayer.swf}
+\end{multicols}
+\end{center}
 \end{frame}
@@
 %----------------------------------------------------------------------------------------
-
+\bibliography{presentation}{}
+\bibliographystyle{plain}
 \end{document}

### 2015-03-10 Presentation/presentation.tex
Added Kelsey's slides

--- 2015-03-09 Presentation/presentation.tex
@@
 \usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
+\usepackage{amsmath}
+\usepackage{mathtools}
 
@@
 
-\title[Short title]{Full Title of the Talk} % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[Short title]{Beep Beep!} % The short title appears at the bottom of every slide, the full title is only on the title page
 
@@
 \titlepage % Print the title page as the first slide
+\includegraphics[width=4.75in]{mariokart.png}
 \end{frame}
@@
 \frametitle{Introduction} 
-Kelsey
+\begin{itemize}
+\item Are second-order models always better than first-order models?
+\item Not always! The PW model for traffic flow, meant to improve the LWR model's behavior near shocks, actually performs worse than the original in some cases.
+\item The PW model is based on compressible fluid dynamics, which ignores some fundamental differences between fluids and cars. Because of this, it can sometimes predict negative velocities.
+\item The AR model corrects the problem of negative velocities by utilizing a convective derivative.
+\end{itemize}
 \end{frame}
@@
 %----------------------------------------------------------------------------------------
-%	PRESENTATION SLIDES
+%	LWR Model
 %----------------------------------------------------------------------------------------
@@
 \begin{frame}
-\frametitle{LWR}
-Kelsey
+\frametitle{Lighthill-Whitham-Richards (LWR) Model}
+
+Proposed by Lighthill and Whitham (1955) and Richards (1956), the LWR model describes traffic flow on a single one-way road without entrances or exits:
+
+\[ \rho_t + \Big(\rho V(\rho)\Big)_x = 0 \]
+
+\begin{itemize}
+\item $\rho$ = density
+\item $V(\rho)$  = preferred velocity, a given nonincreasing function of $\rho$, nonnegative for $\rho$ between 0 and $\rho_m$ (the``jam" density)
+\item Predicts piece-wise smooth density, with transitions between regions approximated by shocks
+\item Problem: Doesn't adequately describe the motion of cars passing through shocks (cars change velocity instantaneously) 
+\end{itemize}
+
 \end{frame}
@@
 %------------------------------------------------
+%	PW Model Background
+%------------------------------------------------
 
 \begin{frame}
-\frametitle{PW}
-Intro, Kelsey. Eigenvectors and values
+\frametitle{Payne-Whitham (PW) Model}
+
+Viewing traffic as a compressible fluid, Payne (1971) and Whitham (1974) introduced a second equation analogous to the conservation of momentum, $\rho v$, in fluids:
+
+\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] (\rho v)_t + \Big(\rho v^2 + p(\rho)\Big)_x = 0 \end{matrix*} \right. \]
+
+\begin{itemize}
+\item $\rho$ = density
+\item $v$ = velocity
+\item $p(\rho)$ = ``anticipation factor", describes how a driver reacts to variations in density with respect to space
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Payne-Whitham (PW) Model}
+
+Rewriting the second equation in terms of velocity and adding relaxation and viscosity terms, we get the full PW Model:
+
+\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = \dfrac{V(\rho) - v}{\tau} + \mu v_{xx} \end{matrix*} \right. \]
+
+\begin{itemize}
+\item $\tau$ and $\mu$ are nonnegative constants
+\item Relaxation term: represents driver's reaction time
+\item Diffusion term: represents the driver's awareness of conditions ahead (and behind)
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Payne-Whitham (PW) Model}
+
+\hspace{2em} A fundamental difference between fluids and traffic: \\[3ex]
+
+\hspace{2em} ``A fluid particle responds to stimuli from the front and from \\ 
+\hspace{2em} behind, but a car is an anisotropic particle that mostly \\
+\hspace{2em} responds to frontal stimuli." CITATION
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Payne-Whitham (PW) Model}
+
+\hspace{5em} Homogeneous model: \hspace{7em} Eigenvalues:
+
+\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & v\end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = 0 \hspace{3em} \lambda = v \pm \sqrt{p'(\rho)} \] \\[5ex]
+
+From the eigenvalues we see that one characteristic speed is greater than the velocity, which means future conditions are partly determined by conditions behind.
+This produces an undesirable effect: negative velocities!
+
 \end{frame}
@@
 %------------------------------------------------
-
+%	AR Model Background
 %------------------------------------------------
@@
 \begin{frame}
-\frametitle{AR}
-Intro, Kelsey. Eigenvectors and values
+\frametitle{A thought experiment}
+
+Consider the situation where a driver is traveling with speed $v$.
+If the density ahead of him is increasing with respect to $x$ but decreasing with respect to $x - vt$, will he speed up or slow down?
+
+\begin{figure}
+\centering
+\includegraphics[width=2in]{experiment.png}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{A thought experiment}
+
+\hspace{2em} Since density ahead is increasing with respect to $x$, the PW \\
+\hspace{2em} model predicts that the driver will slow down (as we will see in \\
+\hspace{2em} our examples).  However, since the traffic ahead is traveling faster, \\
+\hspace{2em} most drivers would actually accelerate.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Aw-Rascle (AR) Model}
+
+Aw and Rascle (2000) claim that instead of depending upon the derivative of pressure with respect to $x$, the anticipation factor should involve the convective derivative:
+
+\[ \partial_t + v \, \partial_x \] \\[5ex]
+
+This change reflects the fact that the driver's perspective is a moving frame of reference.
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Aw-Rascle (AR) Model}
+
+Using the convective derivative, the proposed AR model is: \\[1ex]
+
+\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] (v + \rho)_t + v(v + \rho)_x = 0 \end{matrix*} \right. \]
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Aw-Rascle (AR) Model}
+
+AR model in conservation form: \\[3ex]
+
+\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] \Big(\rho (v + p(\rho)) \Big)_t + \Big(\rho v(v + p(\rho)) \Big)_x = 0 \end{matrix*} \right. \] \\[5ex]
+
+Conserved quantities:
+\begin{enumerate}
+\item Mass (density): $\rho$
+\item ``Momentum": $y = v + p(\rho)$
+\end{enumerate}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Aw-Rascle (AR) Model}
+
+AR model rewritten in terms of density and velocity: \\[1ex]
+
+\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] v_t + \Big( v - \rho p'(\rho) \Big)_x = 0 \end{matrix*} \right. \] \\[3ex]
+
+\hspace{5em} Linearized system: \hspace{7em} Eigenvalues:
+\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] 0 & v - \rho p'(\rho)  \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = 0 \hspace{5ex} \begin{matrix*}[l] \lambda_1 = v - \rho p'(\rho) \\[1ex] \lambda_2 = v \end{matrix*} \] \\[1ex]
+
+Note that neither characteristic speed is greater than the velocity!
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Aw-Rascle (AR) Model}
+
+Aw and Rascle (2000) propose some conditions for traffic models:
+\begin{enumerate}
+\item The system must be hyperbolic.
+\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity must remain nonnegative and bounded from above.
+\item When solving the Riemann problem with data $Q_{\pm} = (\rho_{\pm}, v_{\pm})$, all waves connecting any state $Q = (\rho, v)$ to its left (behind it) must have a propagation speed (eigenvalue or shock speed) at most equal to the velocity $v$.
+\item The solution to the Riemann problem must agree with the qualitative properties that each driver practically observes every day.
+\item Near the vacuum, the solution to the Riemann problem must be very sensitive to the data (no continuous dependence with respect to initial data at $\rho = 0$). 
+\end{enumerate}
+
 \end{frame}

### 2015-03-10 FirstDraftMaterials/Draft.tex

@@
 \subsection{Hugoniot Loci and Integral Curves}
-Brisa
+In \cite{AwRascle2000}, they consider the simplified PW model
+\begin{align*}
+&\rho_t + \left( \rho v\right)_x = 0 \\
+&\left( \rho v\right)_t + \left( \rho v^2 + p(\rho )\right)_x = 0,
+\end{align*}
+which is equivalent to the system shown above if we let $1/\tau = 0$ and $\mu = 0$.
+From this system, the equation for the Hugoniot loci can be found to be
+\begin{align*}
+\rho v = \rho v_* \pm \rho \sqrt{\left( \rho - \rho_*\right) \left( p(\rho ) - p(\rho_*)\right) / \rho_*\rho }.
+\end{align*}
+Using $p(\rho) = \rho^{\gamma}$ for $\gamma \neq 1$, since this is the value used in the AR model we will examine later, the equations for the integral curves can be found to be
+\begin{align*}
+\rho v = \rho v_* + \frac{2 \rho}{\gamma - 1}\left( \sqrt{p'(\rho_*)} - \sqrt{p'(\rho)}\right)
+\end{align*}
+for $\lambda_1$, and 
+\begin{align*}
+\rho v = \rho v_* + \frac{2 \rho}{\gamma - 1}\left( \sqrt{p'(\rho)} - \sqrt{p'(\rho_*)} \right)
+\end{align*}
+for $\lambda_2$.
+
+\begin{figure}[h!]
+ \centering
+ \subfigure[Hugonio loci for $\lambda_1$ and $\lambda_2$ shown in magenta and red, respectively.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_Loci.jpg}
+   }
+ \subfigure[Integral curves for $\lambda_1$, $\lambda_2$ shown in light and dark green, respectively.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_IntegralCurves.jpg}
+   }
+ \subfigure[$\lambda_1$ loci and integral curve passing through
+  a point $(\rho_*, v_*)$, with $\lambda_1$ contour lines shown in cyan.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_Lambda1Curves.jpg}
+   }
+    \subfigure[$\lambda_2$ loci and integral curve passing through
+  a point $(\rho_*, v_*)$, with $\lambda_2$ contour lines shown in cyan.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_Lambda2Curves.jpg}
+   }
+ \subfigure[Valid curves for a point $q_l$.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_validStateCurvesLeft.jpg}
+   }
+ \subfigure[Valid curves for a point $q_r$.]{
+  \includegraphics[width=45mm]{../MatlabCode/Images/PW_validStateCurvesRight.jpg}
+   }
+ \caption[Optional caption for list of figures]
+ {Figures of the Hugonio loci and integral curves for $\lambda_1$ and $\lambda_2$ in the $M = (\rho, \rho v)$ plane.
+Note that loci are shown in red and magenta and integral curves are shown in green.}
+  \label{fig:PW_curves}
+\end{figure}
+
+Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point.
+When traveling across a shock from a left state, $q_l$, to a right state $q_r$, $\lambda$ must decrease.
+When traveling across a rarefaction wave from a left state, $q_l$, to a right state $q_r$, $\lambda$ must increase.
+In Figure \ref{fig:PW_curves} (c) and (d) we see the Hugoniot loci and integral curves for this system with the contour lines for $\lambda_1$ and $\lambda_2$.
+Using these plots, we can determine which areas of the curves are valid shocks and which are valid rarefaction waves.
+Figure \ref{fig:PW_curves} (e) and (d) shows the points $q_l$ and $q_r$, respectively, with only the valid waves displayed.
+Here, shocks are shown in red or magenta and rarefaction waves are shown in either light or dark green.
 
@@
 \subsubsection{Hugoniot Loci}
-In order to consider the Hugoniot loci for this system we must first write the system in conversation form. 
+In order to consider the Hugoniot loci for this system we must first write the system in conservation form. 
 To accomplish this, note that multiplying the first equation by $(v + p(\rho ))$ and the second equation 
@@
 \end{align*}
-except if one of the two densities, $\rho_*$, $\rho$ is zero. WHY DOES THIS NOT MATTER? Therefore, we can say that
+except if one of the two densities, $\rho_*$, $\rho$ is zero. Therefore, we can say that
 \begin{align*}
@@
 \end{align*}
-(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AR:eq6}) gives
-\begin{align*}
-\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
-\rho(s - v)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
-v_* + p(\rho_* )&= v + p(\rho ).
-\end{align*}
-Therefore, the Hugoniot loci for this case in the Y plane are given by
-\begin{align*}
-y = v_* + p(\rho_* ),
-\end{align*}
-since $y = v + p(\rho )$ by definition. Note that solving (\ref{AR:eq3}) for $s$ gives
-\begin{align*}
-s = \frac{\rho_*v_* - \rho v}{\rho_* - \rho}.
-\end{align*}
-Using the fact that $v_* = v + p(\rho) - p(\rho_*)$, we get that
-\begin{align*}
-s = \frac{\rho_*\left( v + p(\rho) - p(\rho_*)\right)- \rho v}{\rho_* - \rho}
-= v - \rho_*\frac{p(\rho ) - p(\rho_*)}{\rho - \rho_*}
-\approx v - \rho p'(\rho)
-\end{align*}
-for $\rho_* \approx \rho$.
-
-In the $U = (\rho, v)$ plane these loci become (a) $v = v_*$, $\rho$ free, and (b) $v = v_* + p(\rho_* ) - p(\rho )$.
- In the $M = (\rho, \rho v)$ plane these loci become (a) $\rho v = \rho_* v_*$, $\rho$ free, 
- and (b) $\rho v = \rho \left( v_* + p(\rho_*)\right) - \rho p(\rho))$. Figure \ref{fig:AR_curves} shows the Hugoniot loci 
- in the different planes. TODO: ADD PARAMETERS USED FOR FIGURES? OR, WILL INCLUDING THE CODE MAKE THAT REDUNDANT? 
- IF WE ADD PARAMETERS, MAYBE THE BEST PRESENTATION WOULD BE IN A TABLE?
 
@@
 \end{figure}
+
+(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AR:eq6}) gives
+\begin{align*}
+\rho_*(s - v_*)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
+\rho(s - v)\left(v_* + p(\rho_* )\right) &= \rho(s - v)\left(v + p(\rho )\right)\\
+v_* + p(\rho_* )&= v + p(\rho ).
+\end{align*}
+Therefore, the Hugoniot loci for this case in the Y plane are given by
+\begin{align*}
+y = v_* + p(\rho_* ),
+\end{align*}
+since $y = v + p(\rho )$ by definition. Note that solving (\ref{AR:eq3}) for $s$ gives
+\begin{align*}
+s = \frac{\rho_*v_* - \rho v}{\rho_* - \rho}.
+\end{align*}
+Using the fact that $v_* = v + p(\rho) - p(\rho_*)$, we get that
+\begin{align*}
+s = \frac{\rho_*\left( v + p(\rho) - p(\rho_*)\right)- \rho v}{\rho_* - \rho}
+= v - \rho_*\frac{p(\rho ) - p(\rho_*)}{\rho - \rho_*}
+\approx v - \rho p'(\rho)
+\end{align*}
+for $\rho_* \approx \rho$.
+
+In the $U = (\rho, v)$ plane these loci become (a) $v = v_*$, $\rho$ free, and (b) $v = v_* + p(\rho_* ) - p(\rho )$.
+ In the $M = (\rho, \rho v)$ plane these loci become (a) $\rho v = \rho_* v_*$, $\rho$ free, 
+ and (b) $\rho v = \rho \left( v_* + p(\rho_*)\right) - \rho p(\rho))$. Figure \ref{fig:AR_curves} shows the Hugoniot loci 
+ in the different planes. TODO: ADD PARAMETERS USED FOR FIGURES? OR, WILL INCLUDING THE CODE MAKE THAT REDUNDANT? 
+ IF WE ADD PARAMETERS, MAYBE THE BEST PRESENTATION WOULD BE IN A TABLE?
 
@@
 
-\subsubsection{Regions of Validity}
-Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point.
-When traveling across a shock from a left state, $q_l$, to a right state $q_r$, $\lambda$ must decrease.
-When traveling across a rarefaction wave from a left state, $q_l$, to a right state $q_r$, $\lambda$ must increase.
-
 \begin{figure}[h!]
@@
   and (d) show the $\lambda_1$ curves in black with the $\lambda_1$ contours 
-  in magenta. The other figures show valid shocks in red and valid rarefaction 
+  in cyan. The other figures show valid shocks in red and valid rarefaction 
   waves in green. The contact discontinuity is shown in blue.}
@@
 \end{figure}
+
+\subsubsection{Regions of Validity}
+Restrictions on the way that $\lambda$ must vary across a wave allow us to identify the regions of the Hugoniot loci and integral curves that are valid for any given point.
+When traveling across a shock from a left state, $q_l$, to a right state $q_r$, $\lambda$ must decrease.
+When traveling across a rarefaction wave from a left state, $q_l$, to a right state $q_r$, $\lambda$ must increase.
 
@@
    }
-% TODO: add subfigure for example1 in PW
+    \subfigure[Solution from the PW model.]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example1.jpg}
+   }
  \caption[Optional caption for list of figures]
@@
    }
-% TODO: add subfigure for example2 in PW
+ \subfigure[Solution from the PW model.]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example2.jpg}
+   }
  \caption[Optional caption for list of figures]
@@
 \begin{center}
-\begin{tabular}{| c | c c |}
+\begin{tabular}{| c | c c  c c|}
 \hline
-& Example 1 & Example 2\\
+& $v_l$ & $\rho_l $ & $v_r$ & $\rho_r $\\
 \hline
-$v_l$ & 0.2 & 0 \\
-$\rho_l$ & 0.3 & 0\\
-$v_r$ & 0.3 & 0 .3\\
-$\rho_r$ & 0.4 & 0.4\\
+Example 1 & 0.2 & .03 & 0.3 & 0.4 \\
+Example 2 & 0 & 0 & 0.3 & 0.4\\
 \hline
@@
 The variables used in these two examples are shown in Table \ref{table:1}.
-Note that in example 1, shown in Figure \ref{fig:example1}, the transition from $m_l = (\rho_l, v_l\rho_l)$ to $m_r = (\rho_r, v_r\rho_r)$ is first a shock and then a contact discontinuity.
-The middle state is denoted by $\tilde{m}$ in the figure.
-In example 2, show in Figure \ref{fig:example2}, the transition from $m_l = (\rho_l, v_l\rho_l)$ to $m_r = (\rho_r, v_r\rho_r)$ is a rarefaction wave, and there is no middle state.
+Note that in the solution from the AR model for example 1, shown in Figure \ref{fig:example1} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is first a shock and then a contact discontinuity.
+The middle state is denoted by $q_m$ in the figure.
+In the solution given by the AR model for example 2, show in Figure \ref{fig:example2} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is a shock wave, and there is no middle state.
 

### 2015-03-10 Presentation/presentation.tex
Updated slides again
Adding edits to presentation.

@@
 \usepackage{media9}
+\usepackage{graphicx}
+\usepackage{mathtools}
 
@@
 
-\title[Short title]{Beep Beep!} % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[Short title]{Comparison of Two Second Order Traffic Flow Models} % The short title appears at the bottom of every slide, the full title is only on the title page
 
@@
 \end{align*}
-This can be rewritten as $q_t + f(q)_x = $ by defining
+This can be rewritten as $q_t + f(q)_x = 0$ by defining
 \begin{align*}
@@
 \hline
-Example 1 & 0.2 & .03 & 0.3 & 0.4 \\
+Example 1 & 0.2 & 0.3 & 0.3 & 0.4 \\
 Example 2 & 0 & 0 & 0.3 & 0.4\\
@@
 \frametitle{Example 1}
+Initial Conditions:
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{| c c  c c|}
+\hline
+$v_l$ & $\rho_l $ & $v_r$ & $\rho_r $\\
+\hline
+0.2 & 0.3 & 0.3 & 0.4 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 The solution to the Riemann problem is:
@@
 \frametitle{Example 1}
+Initial Conditions:
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{| c c  c c|}
+\hline
+$v_l$ & $\rho_l $ & $v_r$ & $\rho_r $\\
+\hline
+0.2 & 0.3 & 0.3 & 0.4 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \begin{center}
@@
 \frametitle{Example 2}
+Initial Conditions:
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{| c c  c c|}
+\hline
+$v_l$ & $\rho_l $ & $v_r$ & $\rho_r $\\
+\hline
+0 & 0 & 0.3 & 0.4 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 The solution to the Riemann problem is:
@@
 \frametitle{Example 2}
+Initial Conditions:
+\begin{table}[t]
+\begin{center}
+\begin{tabular}{| c c  c c|}
+\hline
+$v_l$ & $\rho_l $ & $v_r$ & $\rho_r $\\
+\hline
+0 & 0 & 0.3 & 0.4 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \begin{center}
@@
 %----------------------------------------------------------------------------------------
+\begin{frame}
+\frametitle{Bibliography}
 \bibliography{presentation}{}
 \bibliographystyle{plain}
+\end{frame}
+
 \end{document}

### 2015-03-11 Presentation/presentation.tex
More edits to presentation

@@
 \frametitle{Introduction}
+\hspace{4.75em} First and second order traffic flow models: \\[3ex]
+
+\[ \begin{matrix*}[l]
+\text{LWR: } & \rho_t + \Big(\rho V(\rho)\Big)_x = 0 \\[4ex]
+
+\text{PW: } & \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[1ex] v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = \dfrac{V(\rho) - v}{\tau} + \mu v_{xx} \end{matrix*} \right. \\[6ex]
+
+\text{AR: } &  \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[1ex] v_t + \Big( v - \rho p'(\rho) \Big)_x = 0 \end{matrix*} \right.
+\end{matrix*} \]
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Introduction} 
 \begin{itemize}
@@
 \hspace{2em} behind, but a car is an anisotropic particle that mostly \\
-\hspace{2em} responds to frontal stimuli." CITATION
+\hspace{2em} responds to frontal stimuli." \cite{Daganzo1995}
 
@@
 
-\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] (v + \rho)_t + v(v + \rho)_x = 0 \end{matrix*} \right. \]
+\[ \left\{ \begin{matrix*}[l] \rho_t + (\rho v)_x = 0 \\[2ex] \Big(v + p(\rho)\Big)_t + v\Big(v + p(\rho)\Big)_x = 0 \end{matrix*} \right. \]
 
@@
 \item Mass (density): $\rho$
-\item ``Momentum": $y = v + p(\rho)$
+\item ``Momentum": $y = \rho(v + p(\rho))$
 \end{enumerate}
@@
 
-Aw and Rascle (2000) propose some conditions for traffic models:
+Aw and Rascle propose some conditions for traffic models \cite{AwRascle2000}:
 \begin{enumerate}
-\item The system must be hyperbolic.
-\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity must remain nonnegative and bounded from above.
-\item When solving the Riemann problem with data $Q_{\pm} = (\rho_{\pm}, v_{\pm})$, all waves connecting any state $Q = (\rho, v)$ to its left (behind it) must have a propagation speed (eigenvalue or shock speed) at most equal to the velocity $v$.
+\item When solving the Riemann problem, the density and velocity must remain nonnegative and bounded from above.
+\item When solving the Riemann problem, all waves connecting any state $Q = (\rho, v)$ to its left (behind it) must have a propagation speed (eigenvalue or shock speed) at most equal to the velocity $v$.
 \item The solution to the Riemann problem must agree with the qualitative properties that each driver practically observes every day.
-\item Near the vacuum, the solution to the Riemann problem must be very sensitive to the data (no continuous dependence with respect to initial data at $\rho = 0$). 
 \end{enumerate}

### 2015-03-13 FirstDraftMaterials/Draft.tex
Updated intro and PW section

@@
 \usepackage{enumerate}
-\setlength\parindent{0pt}
 
@@
 
+\title{A comparison of two second order traffic flow models}
+\author{Kelsey Maass and Brisa Davis}
 \begin{document}
+\maketitle
 
@@
 When a model is called ``higher order", we usually associate it with being more accurate than a first order model.
-This is not always the case.
-Adding complexity to a model does not necessarily improve it, as is demonstrated in the class of higher order traffic models developed with the intention of improving the first order traffic model proposed by Lighthill and Whitham (1955) and Richards (1956).
-As Daganzo observes in his paper, \\
-
-``A new theory (e.g.
-relativity) is deemed successful if it explains previously unexplained phenomena, plus everything else that was correctly explained with the theory it intends to replace (e.g.
-Newtonian physics).
-In the case of traffic flow, the new theories (high-order fluid models) fail to explain the behavior of traffic at the end of a queue (unlike the simpler, older LWR model; which explains it perfectly)." \\
-
-In this paper we examine the second order traffic model developed by Payne (1971) and Whitham (1974) which was meant to improve how the LWR model explains traffic behavior near shocks.
-Their model takes its form from equations used to describe compressible fluid dynamics.
-Unfortunately, cars and fluid particles have some fundamental differences which make the model actually perform worse than the LWR model in certain cases.
-For example, by smoothing out discontinuities, the PW model predicts negative velocities, which is clearly a problem.
+However, this is not always the case.
+Adding complexity to a model does not necessarily improve it, as is demonstrated in the class of higher order traffic flow models developed with the intention of improving the first order traffic flow model proposed by Lighthill, Whitham and Richards.
+As Carlos F.
+Daganzo observes in his paper \cite{Daganzo1995}, ``A new theory ...
+is deemed successful if it explains previously unexplained phenomena, plus everything else that was correctly explained with the theory it intends to replace."  Unfortunately, some of these higher order models are not successful because they actually behave worse than the original.
 \\
 
-In addition to considering the PW model, we also take a look a more successful model proposed by Aw and Rascle (2000).
+In this paper we first examine the second order traffic model developed by Payne and Whitham which was meant to improve how the LWR model explains traffic behavior near shocks.
+Viewing traffic as a compressible fluid, the model borrows equations used to describe compressible fluid dynamics.
+Unfortunately, due to the anisotropic nature of cars, the model actually performs worse than the LWR model in certain cases.
+For example, by smoothing out discontinuities, the PW model sometimes predicts negative velocities.
+\\
+
+In addition to considering the PW model, we also take at a look a more successful model proposed by Aw and Rascle \cite{AwRascle2000}.
 These authors challenge some of the assumptions of the PW model and suggest the use of a convective derivative to fix the problem of negative velocities.
@@
 \section{LWR Model}
-We can start by considering a first order (one equation) model of traffic flow on a one-way highway without any exits, entrances, or passing.
+We start by considering a first order (one equation) model of traffic flow on a one-way road without any exits, entrances, or passing.
 In this case, we can expect that the number of cars will be conserved, so we can use the equation for conservation of mass (with density $\rho$ and velocity $v$):
@@
 \item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho \, dt = -\Big[ f( \rho ) \Big]_{x_1}^{x_2}$
-\item Differential form of equation: $\rho_t + f( \rho )_x = 0$
-\begin{enumerate}[i]
-\item Light traffic: $\rho_t + (\rho v)_x = 0$
-\item Congested traffic: $\rho_t + \Big( \rho V(\rho) \Big)_x = 0$ (LWR model)
-\end{enumerate}
+\item Differential form of equation: $\rho_t + f( \rho )_x = \rho_t + \Big( \rho V(\rho) \Big)_x = 0$
 \end{enumerate}
 
-The LWR behaves well macroscopically, but it does have some flaws.
-For example, it assumes cars change their velocity instantaneously as they travel through shocks.
-One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big( \rho V(\rho) \Big)_x = \mu \rho_{xx}$.
-This term is meant to account for the drivers awareness of the road ahead.
-However, this model can be shown to predict negative velocities in some instances, which is obviously wrong.
+The LWR model uses the flux function $f(\rho) = \rho V(\rho)$, where $V(\rho)$ represents the preferred, or equilibrium, velocity, a given nonincreasing function of $\rho$, nonnegative for $\rho$ between 0 and $\rho_m$, the ``jam" density.
+The LWR model behaves well macroscopically, predicting piece-wise smooth density with transitions between regions approximated by shocks.
+However, it assumes that cars change their velocities instantaneously as they travel through shocks.
+One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big(\rho V(\rho))_x = \mu \rho_{xx}$.
+This additional term is meant to account for the driver's awareness of the road ahead and behind.
+However, this model can be shown to predict negative velocities in some instances.
 
@@
 Another approach to improving the LWR model would be to use a second order (two equation) model.
-Since cars can be thought of as a compressible gas, we can use equations for conservation of mass (above) and conservation of momentum, $\rho v$, used to model compressible gas flow:
+Since cars can be thought of as a compressible fluid, we can use equations for conservation of mass (above) and conservation of momentum, $\rho v$, used to model compressible fluid flow:
 
@@
 \item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho v \, dt = -\Big[ f( \rho v ) \Big]_{x_1}^{x_2}$
-\item Differential form of equation: $(\rho v)_t + f( \rho v )_x = (\rho v)_t + \Big( \rho v^2 + \rho \Big)_x = 0$
+\item Differential form of equation: $(\rho v)_t + f( \rho v )_x = (\rho v)_t + \Big( \rho v^2 + p(\rho) \Big)_x = 0$
 \end{enumerate}
 
-We can rewrite the second equation in terms of velocity instead of momentum: $v_t + v v_x + \dfrac{p'(\rho)}{\rho} \rho_x = 0$. \\
-
-The PW model is a modification of these equations for conservation of mass and conservation of momentum in gas dynamics, with added relaxation and diffusion terms:
+Here the flux function contains a contribution from the momentum, $(\rho v)v = \rho v^2$, with an additional pressure term.
+Since traffic doesn't really have pressure, this term is thought of as an ``anticipation factor", describing how a driver reacts to variations in density with respect to space.
+Rewriting the system in terms of velocity and adding relaxation and viscosity terms, we get the full PW model:
 
@@
 
+The first term on the right hand side is a relaxation term meant to mimic the driver's tendency to drive at the preferred velocity, where $\tau$ represents the driver's reaction time.
+The second term, like the diffusion term above, represents the driver's awareness of nearby conditions.
+Both $\tau$ and $\mu$ are nonnegative constants.
+
 At first glance, this seems like a reasonable model.
-However, there a couple of fundamental differences between gas particles and cars that could be problematic.
-For example, as pointed out by Daganzo, ``a fluid particle responds to stimuli from the front and from behind, but a car is an anisotropic particle that mostly responds to frontal stimuli."  If we consider the eigenvalues of the Jacobian matrix of the linearized system, we see that one of the eigenvalues, $\lambda_2 = v + \sqrt{p'(\rho)}$, represents information traveling faster than the velocity of cars in the system.
-This implies that traffic conditions are partially determined by what is going on behind.
+One fundamental difference between cars and fluids causes this model to predict nonphysical situations.
+As pointed out by Daganzo, 	``a fluid particle responds to stimuli from the front and from behind, but a car is an anisotropic particle that mostly responds to frontal stimuli" \cite{Daganzo1995}.
+If we consider the eigenvalues of the linearized system,
+
+\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & v \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] 0 \end{pmatrix}, \]
+
+we see that the largest eigenvalue, $\lambda_2 = v + \sqrt{p'(\rho})$, represents information traveling faster than the velocity of cars in the system.
+This implies that traffic conditions are partially determined by what is going on behind the drivers.
+This contradicts the anisotropic nature of cars.
 

### 2015-03-13 FirstDraftMaterials/Draft.tex
Edited AR section

@@
 However, this is not always the case.
-Adding complexity to a model does not necessarily improve it, as is demonstrated in the class of higher order traffic flow models developed with the intention of improving the first order traffic flow model proposed by Lighthill, Whitham and Richards.
+Adding complexity to a model does not necessarily improve it, as is demonstrated in the class of higher order traffic flow models developed with the intention of improving the first order traffic flow model proposed by Lighthill and Whitham \cite{Lighthill1955}, and Richards \cite{Richards1956}.
 As Carlos F.
@@
 
-In this paper we first examine the second order traffic model developed by Payne and Whitham which was meant to improve how the LWR model explains traffic behavior near shocks.
+In this paper we first examine the second order traffic model developed by Payne \cite{Payne1971} and Whitham \cite{Whitham1974} which was meant to improve how the LWR model explains traffic behavior near shocks.
 Viewing traffic as a compressible fluid, the model borrows equations used to describe compressible fluid dynamics.
@@
 \subsection{Theory}
-Another issues with the PW model, as Aw and Rascle point out in their paper, is that there is no conservation of momentum in car traffic.
-Instead, the term involving pressure is mean to represent an anticipation factor, or how a driver would react to a variation in the concentration of cars with respect to space.
-However, what we should really be concerned with is the perspective of the driver, which is relative to a moving timeframe.
-In this case, we should be using a convective derivative, $\partial_t + v \partial_x$, of the pressure.
-The new model becomes:
+To illustrate the PW model's problematic representation of cars as a fluid, Aw and Rascle \cite{AwRascle2000} consider the following situation.
+Assume you are in your car, traveling with velocity $v$.
+If the density ahead of you is increasing with respect to $x$ but decreasing with respect to $x - vt$, would you speed up or slow down?
+Although the cars ahead are more densely spaced, they are traveling faster than you, so you would most likely accelerate.
+However, since the density ahead is increasing with respect to $x$, the PW model predicts that you would slow down.
 
-\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho v)_x = 0 \\[1ex] & (v + p)_t + v (v + p)_x = 0 \end{matrix*} \right. \]
+According to Aw and Rascle, the model should be more concerned with the perspective of the driver, which is relative to a moving frame of reference.
+In this case, we should involve the convective derivative of the anticipation factor, $\partial_t + v \partial_x$, instead of the space derivative.
+They propose the model:
 
-The second equation can also be rewritten as an equation for velocity: $v_t + \Big(v - \rho p'(\rho)\Big)v_x = 0$. \\
+\[ \left\{ \begin{matrix*}[l] & \rho_t + (\rho v)_x = 0 \\[1ex] & \Big(v + p(\rho)\Big)_t + v \Big(v + p(\rho)\Big)_x = 0 \end{matrix*} \right. \]
 
-This updated model no longer conserves momentum.
-Instead, it conserves the quantity $y = v + p(\rho)$, which we can see if we rearrange the second equation to a conservation form: $(\rho (v + p(\rho)))_t + (\rho v(v + p(\rho)))_x = 0$.
-While the authors admit that there is no obvious physical interpretation of this quantity, they claim that it is the most natural pair of conservative variables.
-\\
+In conservative form, the second equation becomes $\Big( \rho (v + p(\rho)) \Big)_t + \Big( \rho v (v + p(\rho)) \Big)_x = 0$.
+Aw and Rascle call this conserved quantity, $y = \rho v + \rho p(\rho)$, "momentum", although they admit that $y$ has no obvious physical interpretation.
+Furthermore, the second equation can also be rewritten as an equation for velocity: $v_t + \Big(v - \rho p'(\rho)\Big)v_x = 0$.
+If we consider the linearized system,
 
-The authors also claim that their improved model satisfies the following important conditions:
+\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] 0 & v - \rho p'(\rho) \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] 0 \end{pmatrix}, \]
+
+we see that one issues is resolved.
+In this model, the largest eigenvalue $\lambda_2 = v$.
+Therefore no characteristic speed is greater than the velocity, so that drivers are no longer influenced by conditions behind them.
+To validate their model, Aw and Rascle claim that their improved model satisfies the following important conditions:
 \begin{enumerate}
 \item The system is hyperbolic.
-\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity must remain nonnegative and bounded from above.
-\item When solving the Riemann problem, no waves connecting any state to its left (behind it) can have a propagation speed greater than the velocity $v$.
-\item The solution to the Riemann problem must agree with qualitative properties that drivers actually experience: braking produces shock waves and acceleration produces rarefaction waves.
-\item Near a vacuum, the solution to the Riemann problem must be sensitive to the data.
+\item When solving the Riemann problem with bounded nonnegative data $(\rho, v)$, the density and velocity remain nonnegative and bounded from above.
+\item When solving the Riemann problem, no waves connecting any state to its left (behind it) have a propagation speed greater than the velocity $v$.
+\item The solution to the Riemann problem agrees with qualitative properties that drivers actually experience: braking produces shock waves and acceleration produces rarefaction waves.
 \end{enumerate}

### 2015-03-13 FirstDraftMaterials/Draft.tex
Added clawpack figures

@@
 \usepackage{enumerate}
+\usepackage{float}
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
  \centering
@@
 
+\begin{figure}[H]
+ \centering
+ \subfigure[Solution from the AR model.]{
+  \includegraphics[width=70mm]{../ClawpackCode/AR/ARexample1.png}
+   }
+ \subfigure[Solution from the PW model.]{
+  \includegraphics[width=70mm]{../ClawpackCode/PW/PWexample1.png}
+   }
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 1 computed in Clawpack.}
+  \label{fig:example2}
+\end{figure}
+
+\begin{figure}[H]
+ \centering
+ \subfigure[Solution from the AR model.]{
+  \includegraphics[width=70mm]{../ClawpackCode/AR/ARexample2.png}
+   }
+ \subfigure[Solution from the PW model.]{
+  \includegraphics[width=70mm]{../ClawpackCode/PW/PWexample2.png}
+   }
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 2 computed in Clawpack.}
+  \label{fig:example2}
+\end{figure}
+
 \bibliography{Draft}{}

### 2015-03-13 FirstDraftMaterials/Draft.tex
Added conclusion, wrote more in examples

@@
 is deemed successful if it explains previously unexplained phenomena, plus everything else that was correctly explained with the theory it intends to replace."  Unfortunately, some of these higher order models are not successful because they actually behave worse than the original.
-\\
 
@@
 For example, by smoothing out discontinuities, the PW model sometimes predicts negative velocities.
-\\
 
@@
 
+In both examples the density and velocity are greater in the right state than in the left state.
+This corresponds to the situation considered by Aw and Rascle that motivated the use of the convective derivative.
+In the first example, we see that the AR model predicts that the density of the middle state is less than both left or right states, but that the velocity increases steadily from left state to right.
+This is what we would expect, as drivers spread out as the accelerate to catch up with the cars ahead.
+On the other hand, the PW model predicts that the middle state will have a greater density than the left state but a smaller velocity.
+This is opposite of what we would expect.
+
+In the second example, we consider the case where there are no cars on the left.
+Once again the AR model predicts what we would expect.
+Here the cars continue to move to the right at a constant velocity, while the state on the left remains zero.
+Note that in Clawpack solutions the density and velocity are undefined in the middle, but that this is a result of numerical issues.
+In reality they should both be zero.
+The PW method again fails to produce reasonable results.
+Not only does it predict negative velocities, but we can see from the density plot that some cars have in fact moved backwards into the left region.
+While it would make sense for fluids to act this way, we do not expect drivers to start moving in reverse when they see that nobody is behind them!
+
 \begin{figure}[H]
@@
 
+\section{Conclusion}
+By studying the examples of the LWR, PW and AR models for traffic flow, we see that higher order models are not always more accurate than first order models.
+Additionally, the assumptions made in developing higher order relations must conform to the phenomena they are designed to model.
+For example, using equations for isotropic fluids to model anisotropic cars in the PW model produces nonphysical results.
+The AR model corrects this problem by using a convective derivative rather than a space derivative for the anticipation factor.
+Therefore, whenever we introduce higher order models, we should follow the example of Aw and Rascle by validating the model's results in a systematic way.
+
 \bibliography{Draft}{}

### 2015-03-13 FirstDraftMaterials/Draft.tex
Taking out some images

--- 2015-03-13 FirstDraftMaterials/Draft.tex
@@
  \centering
- \subfigure[$\lambda_1$-curves in the U plane.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U_lamb1.jpg}
-   }
- \subfigure[$\lambda_2$-curves in the U plane.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U_lamb2.jpg}
-   }
- \subfigure[$\lambda_1$-curve and $\lambda_2$-curve in the U plane passing through
-  the point $(\rho_*, v_*)$ denoted by the red point.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_U.jpg}
-   }
-    \subfigure[$\lambda_1$-curves in the M plane.]{
+    \subfigure[$\lambda_1$-curves.]{
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb1.jpg}
    }
- \subfigure[$\lambda_2$-curves in the M plane.]{
+ \subfigure[$\lambda_2$-curves.]{
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M_lamb2.jpg}
    }
- \subfigure[$\lambda_1$-curve and $\lambda_2$-curve in the M plane passing through 
- the point $(\rho_*,m_*)$ denoted by the red point.]{
+ \subfigure[$\lambda_1$-curve and $\lambda_2$-curve passing through 
+ the point $(\rho_*,m_*)$.]{
   \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_M.jpg}
    }
-       \subfigure[$\lambda_1$-curves in the Y plane.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y_lamb1.jpg}
-   }
- \subfigure[$\lambda_2$-curves in the Y plane.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y_lamb2.jpg}
-   }
- \subfigure[$\lambda_1$-curve and $\lambda_2$-curve in the Y plane passing through 
- the point $(\rho_*, y_*)$ denoted by the red point.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/HLIC_Y.jpg}
-   }
  \caption[Optional caption for list of figures]
- {Figures of the Hugonio loci and integral curves for $\lambda_1$ and $\lambda_2$ 
- in each of the different planes. Since the loci and integral curves coincide, they are 
+ {Figures of the Hugonio loci and integral curves for $\lambda_1$ and $\lambda_2$. 
+ Since the loci and integral curves coincide, they are 
  both simply called the ``curves'' in the captions of the above figures. 
- Figures (c), (f), (i) show both curves for a given point in each plane.}
+ Figure (c) shows both curves for a given point.}
   \label{fig:AR_curves}
@@
  \centering
- \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the U plane.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U.jpg}
-   }
-    \subfigure[Valid waves for the point $q_l$ in the U plane.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_ql.jpg}
-   }
-    \subfigure[Valid waves for the point $q_r$ in the U plane.]{
-  \includegraphics[width=45mm]{../MatlabCode/Images/Validity_U_qr.jpg}
-   }
-    \subfigure[$\lambda_1$-curves with contours of $\lambda_1$ in the M plane.]{
+    \subfigure[$\lambda_1$-curves with contours of $\lambda_1$.]{
   \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M.jpg}
    }
- \subfigure[Valid waves for the point $q_l$ in the M plane.]{
+ \subfigure[Valid waves for the point $q_l$.]{
   \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_ql.jpg}
    }
- \subfigure[Valid waves for the point $q_r$ in the M plane.]{
+ \subfigure[Valid waves for the point $q_r$.]{
   \includegraphics[width=45mm]{../MatlabCode/Images/Validity_M_qr.jpg}

### 2015-03-13 FirstDraftMaterials/Draft.tex
Adding code to show approximating examples, and general edits to draft.

@@
 However, it assumes that cars change their velocities instantaneously as they travel through shocks.
-One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big(\rho V(\rho))_x = \mu \rho_{xx}$.
+One way to eliminate shocks would be to add a diffusion term: $\rho_t + \Big(\rho V(\rho)\Big)_x = \mu \rho_{xx}$.
 This additional term is meant to account for the driver's awareness of the road ahead and behind.
@@
 Figure \ref{fig:PW_curves} (e) and (d) shows the points $q_l$ and $q_r$, respectively, with only the valid waves displayed.
+For these plots,
+$q_* = (\rho_*,\rho_*v_*) = (0.2, 0.04)$ and $\gamma = 0.2$. 
 Here, shocks are shown in red or magenta and rarefaction waves are shown in either light or dark green.
@@
 In conservative form, the second equation becomes $\Big( \rho (v + p(\rho)) \Big)_t + \Big( \rho v (v + p(\rho)) \Big)_x = 0$.
-Aw and Rascle call this conserved quantity, $y = \rho v + \rho p(\rho)$, "momentum", although they admit that $y$ has no obvious physical interpretation.
+Aw and Rascle call this conserved quantity, $y = \rho v + \rho p(\rho)$,``momentum", although they admit that $y$ has no obvious physical interpretation.
 Furthermore, the second equation can also be rewritten as an equation for velocity: $v_t + \Big(v - \rho p'(\rho)\Big)v_x = 0$.
@@
 \end{align}
-From (\ref{AR:eq5}), we know that there are two cases:
+From equation (\ref{AR:eq5}), we know that there are two cases:
 
@@
 describe a wave. Therefore, $s = v$. Note that we have no restrictions on $y$ other than its definition. 
-Therefore this loci in the $Y$ plane is given by 
+Therefore this loci in the $Y = (\rho, y)$ plane is given by 
 \begin{align*}
@@
 
-(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into (\ref{AR:eq6}) gives
+(b) $\rho_*(s - v_*) = \rho (s - v) \neq 0$. In this case, plugging this into equation (\ref{AR:eq6}) gives
 \begin{align*}
@@
 \end{align*}
-Therefore, the Hugoniot loci for this case in the Y plane are given by
+Therefore, the Hugoniot loci for this case in the $Y = (\rho, y)$ plane are given by
 \begin{align*}
@@
 \end{align*}
-since $y = v + p(\rho )$ by definition. Note that solving (\ref{AR:eq3}) for $s$ gives
+since $y = v + p(\rho )$ by definition. Note that solving equation (\ref{AR:eq3}) for $s$ gives
 \begin{align*}
@@
  and (b) $\rho v = \rho \left( v_* + p(\rho_*)\right) - \rho p(\rho))$. Figure \ref{fig:AR_curves} shows the Hugoniot loci 
- in the different planes. TODO: ADD PARAMETERS USED FOR FIGURES? OR, WILL INCLUDING THE CODE MAKE THAT REDUNDANT? 
- IF WE ADD PARAMETERS, MAYBE THE BEST PRESENTATION WOULD BE IN A TABLE?
+ in the M plane. Note that for these figures we used $\gamma = 0.2$, and in Figure \ref{fig:AR_curves} (c) we used $\rho_* = 0.2$ and $m_* = 0.04$.
 
@@
 \end{align*}
-where the second equation is found by multiplying (\ref{AR:eq1}) 
-by $p'(\rho)$ and subtracting it from (\ref{AR:eq1.5}). 
+where the second equation is found by multiplying equation (\ref{AR:eq1}) 
+by $p'(\rho)$ and subtracting it from equation (\ref{AR:eq1.5}). 
 This can be rewritten as $q_t + Aq_t = 0$ by defining
@@
  \caption[Optional caption for list of figures]
- {Figures showing the calculation of the valid waves for a given point. Figures (a)
-  and (d) show the $\lambda_1$ curves in black with the $\lambda_1$ contours 
+ {Figures showing the calculation of the valid waves for a given point. Figure (a)
+  shows the $\lambda_1$ curves in black with the $\lambda_1$ contours 
   in cyan. The other figures show valid shocks in red and valid rarefaction 
@@
 Therefore, it is important to take into account the contour lines of $\lambda$ when 
-looking at the loci and integral curves. Figures \ref{fig:AR_validity}(a) and 
-\ref{fig:AR_validity}(d) show the $\lambda_1$-curves in the $U$ and $M$ planes,
+looking at the loci and integral curves. Figure \ref{fig:AR_validity}(a) 
+shows the $\lambda_1$-curves in the $M$ plane,
 with the contour lines of $\lambda_1$ shown in magenta. This information is used
@@
 specific points $q_*$, depending on whether $q_* = q_l$ or $q_* = q_r$. For these plots,
-$q_* = (\rho_*,v_*) = (0.2, 0.2)$. Note that the contact discontinuity is valid
-anywhere in the plane.
+$q_* = (\rho_*,\rho_*v_*) = (0.2, 0.04)$ and $\gamma = 0.2$. Note that the contact discontinuity is valid
+anywhere in the plane since $\lambda_2$ is linearly degenerate.
 
@@
 \section{Examples}
-
-\begin{figure}[H]
- \centering
- \subfigure[Solution from the AR model.]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
-   }
-    \subfigure[Solution from the PW model.]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example1.jpg}
-   }
- \caption[Optional caption for list of figures]
- {Solutions to the Riemann problem for example 1 from the AR and PW models.}
-  \label{fig:example1}
-\end{figure}
-
-\begin{figure}[H]
- \centering
- \subfigure[Solution from the AR model.]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
-   }
- \subfigure[Solution from the PW model.]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example2.jpg}
-   }
- \caption[Optional caption for list of figures]
- {Solutions to the Riemann problem for example 2 from the AR and PW models.}
-  \label{fig:example2}
-\end{figure}
 
@@
 
-Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples.
-The variables used in these two examples are shown in Table \ref{table:1}.
-Note that in the solution from the AR model for example 1, shown in Figure \ref{fig:example1} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is first a shock and then a contact discontinuity.
-The middle state is denoted by $q_m$ in the figure.
-In the solution given by the AR model for example 2, show in Figure \ref{fig:example2} (a), the transition from $q_l = (\rho_l, v_l\rho_l)$ to $q_r = (\rho_r, v_r\rho_r)$ is a shock wave, and there is no middle state.
+Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples. The variables used in these two examples are shown in Table \ref{table:1}.
 
@@
 This corresponds to the situation considered by Aw and Rascle that motivated the use of the convective derivative.
-In the first example, we see that the AR model predicts that the density of the middle state is less than both left or right states, but that the velocity increases steadily from left state to right.
-This is what we would expect, as drivers spread out as the accelerate to catch up with the cars ahead.
+In the first example, we see that the AR model predicts that the density of the middle state is less than both left or right states, but that the velocity increases steadily from left state to the middle state and then remains constant while the density changes due to the contact discontinuity.
+This is what we would expect, as drivers spread out as they accelerate to catch up with the cars ahead.
 On the other hand, the PW model predicts that the middle state will have a greater density than the left state but a smaller velocity.
@@
 While it would make sense for fluids to act this way, we do not expect drivers to start moving in reverse when they see that nobody is behind them!
+
+\begin{figure}[H]
+ \centering
+ \subfigure[Solution from the AR model. The transition from $q_l$ to $q_r$ is first a shock and then a contact discontinuity.]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
+   }
+    \subfigure[Solution from the PW model.  The transition from $q_l$ to $q_r$ is first a shock and then a rarefaction wave. ]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example1.jpg}
+   }
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 1 from the AR and PW models. The middle state is denoted by $q_m$.}
+  \label{fig:example1}
+\end{figure}
+
+\begin{figure}[H]
+ \centering
+ \subfigure[Solution from the AR model. The transition from $q_l$ to $q_r$ is a contact discontinuity,  shown in blue, there is no middle state.]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
+   }
+ \subfigure[Solution from the PW model. The transition from $q_l$ to $q_r$ is a rarefaction wave, shown in dark green, there is no middle state.]{
+  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example2.jpg}
+   }
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 2 from the AR and PW models.}
+  \label{fig:example2}
+\end{figure}
 

### 2015-03-13 MatlabCode/AR_examples.m

@@ -12,12 +12,15 @@ rho = 0:0.001:5;
 size_font = 20;
 
 % Loop plots both examples
-for example = 1:2
+for example = 1:3
     figure
     % Setting up data for the two points ql and qr
     if (example == 1)
         v_star = [0.2 .3];
         rho_star = [0.3 .4];
+    elseif (example == 2)
+        v_star = [0.1 .3];
+        rho_star = [0.1 .4];
     else
         v_star = [0 .3];
         rho_star = [0 .4];
@@ -97,6 +100,14 @@ for example = 1:2
         cd export_fig_code
         export_fig( gcf, ...      % figure handle
             '../Images/AR_example1',... % name of output file
+            '-painters', ...      % renderer
+            '-jpg', ...           % file format
+            '-r72' );             % resolution in dpi
+        cd ..
+    elseif (example == 2)
+        cd export_fig_code
+        export_fig( gcf, ...      % figure handle
+            '../Images/AR_example2_smallql',... % name of output file
             '-painters', ...      % renderer
             '-jpg', ...           % file format
             '-r72' );             % resolution in dpi

### 2015-03-13 MatlabCode/MPlot.m

@@ -123,7 +123,7 @@ elseif (model == 2)
             
             rv1_loci(h,k) = rho(k)*v_star(h) + var;
             rv2_loci(h,k) = rho(k)*v_star(h) - var;
-            
+
             rv1_int(h,k) = rho(k)*(v_star(h) ...
                 + (2/(gamma-1))*(sqrt(pprime(rho_star(h)))...
                 - sqrt(pprime(rho(k)))));

### 2015-03-13 MatlabCode/PW_examples.m

@@ -13,12 +13,15 @@ rho = 0:0.001:5;
 size_font = 20;
 
 % Loop plots both examples
-for example = 1:2
+for example = 1:3
     figure
     % Setting up data for the two points ql and qr
     if (example == 1)
         v_star = [0.2 .3];
         rho_star = [0.3 .4];
+    elseif (example == 2)
+        v_star = [0.05 .3];
+        rho_star = [0.05 .4];
     else
         v_star = [0 .3];
         rho_star = [0 .4];
@@ -140,6 +143,14 @@ for example = 1:2
         cd export_fig_code
         export_fig( gcf, ...      % figure handle
             '../Images/PW_example1',... % name of output file
+            '-painters', ...      % renderer
+            '-jpg', ...           % file format
+            '-r72' );             % resolution in dpi
+        cd ..
+    elseif (example == 2)
+        cd export_fig_code
+        export_fig( gcf, ...      % figure handle
+            '../Images/PW_example2_smallql',... % name of output file
             '-painters', ...      % renderer
             '-jpg', ...           % file format
             '-r72' );             % resolution in dpi

### 2015-03-13 FirstDraftMaterials/Draft.tex
Modified conclusion

@@
 At first glance, this seems like a reasonable model.
-One fundamental difference between cars and fluids causes this model to predict nonphysical situations.
+However, one fundamental difference between cars and fluids causes this model to predict nonphysical situations.
 As pointed out by Daganzo, 	``a fluid particle responds to stimuli from the front and from behind, but a car is an anisotropic particle that mostly responds to frontal stimuli" \cite{Daganzo1995}.
@@
 
-\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & v \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] 0 \end{pmatrix}, \]
+\[ \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_t + \begin{pmatrix} v & \rho \\[1ex] \dfrac{p'(\rho)}{\rho} & v \end{pmatrix} \begin{pmatrix} \rho \\[1ex] v \end{pmatrix}_x = \begin{pmatrix} 0 \\[1ex] \dfrac{V(\rho) - v}{\tau} + \mu v_{xx} \end{pmatrix}, \]
 
@@
 If the density ahead of you is increasing with respect to $x$ but decreasing with respect to $x - vt$, would you speed up or slow down?
-Although the cars ahead are more densely spaced, they are traveling faster than you, so you would most likely accelerate.
+Although the cars ahead are more densely packed, they are traveling faster than you, so you would most likely accelerate.
 However, since the density ahead is increasing with respect to $x$, the PW model predicts that you would slow down.
@@
 we see that one issues is resolved.
-In this model, the largest eigenvalue $\lambda_2 = v$.
+In this model, the largest eigenvalue is $\lambda_2 = v$.
 Therefore no characteristic speed is greater than the velocity, so that drivers are no longer influenced by conditions behind them.
@@
 The AR model corrects this problem by using a convective derivative rather than a space derivative for the anticipation factor.
+Furthermore, the authors verified that their model behaved reasonably by checking it against a list of important conditions.
 Therefore, whenever we introduce higher order models, we should follow the example of Aw and Rascle by validating the model's results in a systematic way.

### 2015-03-13 FirstDraftMaterials/Draft.tex
Reordered example section

@@
 
-\begin{table}[t]
+Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples. The variables used in these two examples are shown in Table \ref{table:1}.
+
+\vspace{-2ex}
+\begin{table}[H]
 \caption{Initial values used in examples 1 and 2.}
@@
 
-Reproducing the Riemann problems found in \cite{AwRascle2000}, we consider two different examples. The variables used in these two examples are shown in Table \ref{table:1}.
-
+\vspace{-4ex}
 In both examples the density and velocity are greater in the right state than in the left state.
@@
 This is opposite of what we would expect.
+
+\begin{figure}[H]
+ \centering
+ \subfigure[Solution from the AR model. The transition from $q_l$ to $q_r$ is first a shock and then a contact discontinuity.]{
+  \includegraphics[width=60mm]{../MatlabCode/Images/AR_example1.jpg}
+   }
+    \subfigure[Solution from the PW model.  The transition from $q_l$ to $q_r$ is first a shock and then a rarefaction wave. ]{
+  \includegraphics[width=60mm]{../MatlabCode/Images/PW_example1.jpg}
+   }
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 1 from the AR and PW models. The middle state is denoted by $q_m$.}
+  \label{fig:example1}
+\end{figure}
+
+\begin{figure}[H]
+ \centering
+ \subfigure[Solution from the AR model.]{
+  \includegraphics[width=65mm]{../ClawpackCode/AR/ARexample1.png}
+   }
+ \subfigure[Solution from the PW model.]{
+  \includegraphics[width=65mm]{../ClawpackCode/PW/PWexample1.png}
+   }
+ \caption[Optional caption for list of figures]
+ {Solutions to the Riemann problem for example 1 computed in Clawpack.}
+  \label{fig:example1_PW}
+\end{figure}
 
@@
  \centering
- \subfigure[Solution from the AR model. The transition from $q_l$ to $q_r$ is first a shock and then a contact discontinuity.]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example1.jpg}
-   }
-    \subfigure[Solution from the PW model.  The transition from $q_l$ to $q_r$ is first a shock and then a rarefaction wave. ]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example1.jpg}
-   }
- \caption[Optional caption for list of figures]
- {Solutions to the Riemann problem for example 1 from the AR and PW models. The middle state is denoted by $q_m$.}
-  \label{fig:example1}
-\end{figure}
-
-\begin{figure}[H]
- \centering
  \subfigure[Solution from the AR model. The transition from $q_l$ to $q_r$ is a contact discontinuity,  shown in blue, there is no middle state.]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/AR_example2.jpg}
+  \includegraphics[width=60mm]{../MatlabCode/Images/AR_example2.jpg}
    }
  \subfigure[Solution from the PW model. The transition from $q_l$ to $q_r$ is a rarefaction wave, shown in dark green, there is no middle state.]{
-  \includegraphics[width=70mm]{../MatlabCode/Images/PW_example2.jpg}
+  \includegraphics[width=60mm]{../MatlabCode/Images/PW_example2.jpg}
    }
@@
  \subfigure[Solution from the AR model.]{
-  \includegraphics[width=70mm]{../ClawpackCode/AR/ARexample1.png}
+  \includegraphics[width=65mm]{../ClawpackCode/AR/ARexample2.png}
    }
  \subfigure[Solution from the PW model.]{
-  \includegraphics[width=70mm]{../ClawpackCode/PW/PWexample1.png}
-   }
- \caption[Optional caption for list of figures]
- {Solutions to the Riemann problem for example 1 computed in Clawpack.}
-  \label{fig:example2}
-\end{figure}
-
-\begin{figure}[H]
- \centering
- \subfigure[Solution from the AR model.]{
-  \includegraphics[width=70mm]{../ClawpackCode/AR/ARexample2.png}
-   }
- \subfigure[Solution from the PW model.]{
-  \includegraphics[width=70mm]{../ClawpackCode/PW/PWexample2.png}
+  \includegraphics[width=65mm]{../ClawpackCode/PW/PWexample2.png}
    }
@@
  {Solutions to the Riemann problem for example 2 computed in Clawpack.}
-  \label{fig:example2}
+  \label{fig:example2_PW}
 \end{figure}

### 2015-03-13 FirstDraftMaterials/Draft.tex
Added sentence in section 3.2

@@
 \subsection{Hugoniot Loci and Integral Curves}
-In \cite{AwRascle2000}, they consider the simplified PW model
+In \cite{AwRascle2000}, they consider the homogeneous PW model (arguing that the diffusion term actually makes the model worse, while the relaxation term neither corrects nor adds to the model's problems):
 \begin{align*}

### 2015-03-13 ClawpackCode/AR/ar.f90
Fixing error in Clawpack code.

@@ -45,7 +45,18 @@ subroutine rp1(maxmx,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
     do 30 i=2-mbc,mx+mbc
 
     !     # compute  averaged quantities:
-        rbar = 0.5d0*(qr(1,i-1) + ql(1,i))
+        if (0.0d0 < qr(1,i-1)) then
+            rl = qr(1,i-1)
+        else
+            rl = 0.0d0
+        endif
+        if (0.0d0 < ql(1,i)) then
+            rr = ql(1,i)
+        else
+            rr = 0.0d0
+        endif
+
+        rbar = 0.5d0*(rl + rr)
         pbar = gamma*rbar**(gamma-1.0d0)
         ubar = 0.5d0*(qr(2,i-1) + ql(2,i))
 

### 2015-03-13 FirstDraftMaterials/Draft.tex
Adding edits.

@@
 \begin{enumerate}[(a)]
-\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho \, dt = -\Big[ f( \rho ) \Big]_{x_1}^{x_2}$
+\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho \, dx = -\Big[ f( \rho ) \Big]_{x_1}^{x_2}$
 \item Differential form of equation: $\rho_t + f( \rho )_x = \rho_t + \Big( \rho V(\rho) \Big)_x = 0$
@@
 \begin{enumerate}[(a)]
-\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho v \, dt = -\Big[ f( \rho v ) \Big]_{x_1}^{x_2}$
+\item Integral form of equation: $\displaystyle \frac{d}{dt} \int_{x_1}^{x_2} \rho v \, dx = -\Big[ f( \rho v ) \Big]_{x_1}^{x_2}$
 \item Differential form of equation: $(\rho v)_t + f( \rho v )_x = (\rho v)_t + \Big( \rho v^2 + p(\rho) \Big)_x = 0$
@@
 Using these plots, we can determine which areas of the curves are valid shocks and which are valid rarefaction waves.
-Figure \ref{fig:PW_curves} (e) and (d) shows the points $q_l$ and $q_r$, respectively, with only the valid waves displayed.
+Figure \ref{fig:PW_curves} (e) and (f) shows the points $q_l$ and $q_r$, respectively, with only the valid waves displayed.
 For these plots,
@@
 shows the $\lambda_1$-curves in the $M$ plane,
-with the contour lines of $\lambda_1$ shown in magenta. This information is used
+with the contour lines of $\lambda_1$ shown in cyan. This information is used
 to determine which areas of the $\lambda_1$ curves are valid shock waves and which
@@
 In the second example, we consider the case where there are no cars on the left.
+In figure \ref{fig:example2}, note that there are two valid curves for both models.
+In order to determine which curve is the correct one to follow, we considered a case when we had a very small density in the left state (rather than a zero density) and then considered the limit as the density on the left went to zero.
+This analysis made it clear that for the AR model the correct curve to follow is the contact discontinuity between $q_l$ and $q_r$, and for the PW model the correct curve to follow is the integral curve between $q_l$ and $q_r$.
+
 Once again the AR model predicts what we would expect.
 Here the cars continue to move to the right at a constant velocity, while the state on the left remains zero.
-Note that in Clawpack solutions the density and velocity are undefined in the middle, but that this is a result of numerical issues.
-In reality they should both be zero.
 The PW method again fails to produce reasonable results.

### 2015-03-13 FirstDraftMaterials/Draft.tex
Edited draft

@@
 The AR model corrects this problem by using a convective derivative rather than a space derivative for the anticipation factor.
-Furthermore, the authors verified that their model behaved reasonably by checking it against a list of important conditions.
-Therefore, whenever we introduce higher order models, we should follow the example of Aw and Rascle by validating the model's results in a systematic way.
+Furthermore, the authors verified that their model behaved reasonably by checking it against a list of important conditions for traffic flow.
+As illustrated in the analysis of each model and the results that they produce, we see that that the AR model is more successful in modeling realistic traffic flow than the PW model.
+Therefore, whenever we introduce higher order relations, we should follow the example of Aw and Rascle by validating the model's results in a systematic way.
 
